---
title: "Classification, UPF, GEF, Kappa"
author: "Torsten Sällström"
format: html
editor: visual
---

This script starts with an Excel-file named Food_itmes, containing variables: food_ID, rater1, rater2, maingroup, subgroup, keyhole, keyhole_ratio, ratio_upf_low, ratio, upf_mid and ratio_upf_up.

```{r}
#| label: setup
#| include: false

project_root <- normalizePath(file.path(getwd(), ".."))

dir_data <- file.path(project_root, "data")
dir_output <- file.path(project_root, "output")

dir.create(dir_output, showWarnings = FALSE, recursive = TRUE)

```

analysis code starts here

```{r, warning=FALSE, message=FALSE}

library(openxlsx)
library(writexl)
library(dplyr)
library(ggplot2)
library(psych)
library(tidyr)

```

Import data

```{r}

data2 <- read.xlsx(file.path(dir_data, "Food_items.xlsx"))

```

Convert variable classes

```{r}
data2$food_ID <- as.character(data2$food_ID)
data2$keyhole <- as.character(data2$keyhole)
data2$keyhole_ratio <- as.factor(data2$keyhole_ratio)
data2$rater1 <- as.factor(data2$rater1)
data2$rater2 <- as.factor(data2$rater2)

```

Count number of unique maingroups and subgroups

```{r}
n_maingroup <- n_distinct(data2$maingroup)
n_subgroup <- n_distinct(data2$subgroup)
n_maingroup
n_subgroup
```

Count the number and percent of upf

```{r}
data2 <- data2 %>% 
  mutate(upf_category = case_when(
    ratio_upf_mid_vux == 0 ~"non-upf",
    ratio_upf_mid_vux >0 & ratio_upf_mid_vux <1 ~"partly upf",
    ratio_upf_mid_vux == 1 ~"upf",
  ))

#Count percentages
summary_table <- data2 %>% 
  count(upf_category) %>% 
  mutate(percent = round(100*n/sum(n), 1))

print(summary_table)

```

Count by subgroup

```{r}
summary_by_subgroup <- data2 %>% 
  group_by(subgroup, upf_category) %>% 
  summarise(count = n(), .groups = "drop") %>% 
  group_by(subgroup) %>% 
  mutate(
    total_in_subgroup = sum(count),
    percent = round(100*count/total_in_subgroup, 1)
  ) %>% 
  ungroup()

print(summary_by_subgroup)

```

Pivot to wide format

```{r}
wide_summary <- summary_by_subgroup %>% 
  dplyr::select(subgroup, upf_category, count, percent, total_in_subgroup) %>% 
  pivot_wider(
    names_from = upf_category,
    values_from = c(count, percent),
    values_fill = 0
  )

print(wide_summary)

write.xlsx(
  x = wide_summary,
  file = file.path(dir_output, "Foodgroup_summary.xlsx")
)


```

What share of food items were classified according to alternative scenarios?

```{r}

alternative <- sum(!(data2$ratio_upf_low == data2$ratio_upf_up &
           data2$ratio_upf_up == data2$ratio_upf_mid_vux)) 

total <- nrow(data2)

alternative
round(100 * alternative / total, 1)

```

KAPPA-calculation

Select columns to keep

```{r}
data1 <- data2 %>% 
  dplyr::select(food_ID, rater1, rater2, subgroup)

```

```{r}

overall_table <- table(data1$rater1, data1$rater2)

overall_kappa <- cohen.kappa(overall_table, alpha = 0.05)

print(overall_kappa)
```

```{r}
overall_n <- nrow(data1)
overall_agreement <- sum(diag(overall_table)) / overall_n * 100
overall_n
overall_agreement

```

```{r}
overall_row <- data.frame(
  category = "Overall",
  n = overall_n,
  percent_agreement = overall_agreement,
  kappa = overall_kappa$weighted.kappa,  # Use weighted kappa here
  lower = overall_kappa$confid["weighted kappa", "lower"],
  upper = overall_kappa$confid["weighted kappa", "upper"]
)

print(overall_row)
```

```{r, message=FALSE, warning=FALSE}

calculate_kappa_by_category <- function(df) {
  # Create the agreement table for this subset of data
  agreement_table <- table(df$rater1, df$rater2)
  
  # Calculate Cohen's kappa for this subset
  kappa <- cohen.kappa(agreement_table, alpha = 0.05)
  
  # Count the number of items in this subset
  n_items <- nrow(df)
  
  # Calculate correct agreement (diagonal of the table)
  correct_agreement <- sum(diag(agreement_table))
  
  # Calculate percent agreement
  percent_agreement <- correct_agreement / n_items * 100
  
  # Return a data frame with the results for this category
return(data.frame(
  n = n_items,
  percent_agreement = percent_agreement,
  kappa = kappa$weighted.kappa,
  lower = kappa$confid["weighted kappa", "lower"],
  upper = kappa$confid["weighted kappa", "upper"]
))
}

# Apply the function to each category
kappa_by_category <- data1 |> 
  group_by(subgroup) |> 
  group_modify(~ calculate_kappa_by_category(.x)) |> 
  ungroup()

# Print the results
print(kappa_by_category)
```

KEYHOLE

Total counts and percentage of keyhole == "yes"

```{r}
summary_total <- data2 %>% 
  summarise(
    upf_category = "Total",
    n_cases = n(),
    n_keyhole_yes = sum(keyhole == "yes", na.rm = TRUE),
    pct_keyhole_yes = round(100 * n_keyhole_yes / n_cases, 1),
    n_keyhole_x = sum(keyhole == "x", na.rm = TRUE),
    pct_keyhole_x = round(100 * n_keyhole_x / n_cases, 1),
    n_keyhole_no = sum(keyhole == "no", na.rm = TRUE),
    pct_keyhole_no = round(100 * n_keyhole_no / n_cases, 1)
  )

summary_total
```

Counts and percentage of keyhole == "yes" per upf_category

value "x" in the keyhole column indicates that the food item can not be categorized according to these criteria.

```{r}
summary_by_category <- data2 %>% 
  group_by(upf_category) %>% 
  summarise(
    n_cases = n(),
    n_keyhole_yes = sum(keyhole == "yes", na.rm = TRUE),
    pct_keyhole_yes = round(100 * n_keyhole_yes / n_cases, 1),
    n_keyhole_x = sum(keyhole == "x", na.rm = TRUE),
    pct_keyhole_x = round(100 * n_keyhole_x / n_cases, 1),
    n_keyhole_no = sum(keyhole == "no", na.rm = TRUE),
    pct_keyhole_no = round(100 * n_keyhole_no / n_cases, 1)
  )

summary_by_category

```

Combine total and per-category summaries into one dataframe

```{r}
summary_final <- bind_rows(summary_total, summary_by_category)

print(summary_final)
```

STACKED BARPLOT

1)  Recreate dataset with both yes and no fractions

```{r}
plot_data <- data2 %>% 
  mutate(upf_category = factor(
    upf_category, 
    levels = c("Total", "non-upf", "partly upf", "upf")
  )) %>% 
  #Add total category
  bind_rows(data2 %>% mutate(upf_category = "Total")) %>% 
  group_by(upf_category, keyhole) %>% 
  summarise(n = n(), .groups = "drop") %>% 
  group_by(upf_category) %>% 
  mutate(perc = 100 * n / sum(n))

```

2)  Rename levels

```{r}
plot_data$upf_category <- factor(
  plot_data$upf_category, 
  levels = c("Total", "non-upf", "partly upf", "upf"),
  labels = c("Total Food Items", "Non-UPF", "Partly UPF", "UPF")
)
```

Create a small data frame with categories and labels

```{r}
label_df <- data.frame(
  upf_category = c("Total Food Items", "Non-UPF", "Partly UPF", "UPF"),
  perc_top = c(100, 100, 100, 100),
  label = c(
    paste0("n = ", summary_final$n_cases[summary_final$upf_category == "Total"]),
    paste0("n = ", summary_final$n_cases[summary_final$upf_category == "non-upf"]),
    paste0("n = ", summary_final$n_cases[summary_final$upf_category == "partly upf"]),
    paste0("n = ", summary_final$n_cases[summary_final$upf_category == "upf"])
 )
)
```

Create another data frame with labels for percentage points

```{r}
right_label_df <- data.frame(
  upf_category = c("Total Food Items", "Non-UPF", "Partly UPF", "UPF"),
  perc_point = summary_final$pct_keyhole_yes,
  label = c(
    paste0("n = ", summary_final$pct_keyhole_yes[summary_final$upf_category == "Total"]),
    paste0("n = ", summary_final$pct_keyhole_yes[summary_final$upf_category == "non-upf"]),
    paste0("n = ", summary_final$pct_keyhole_yes[summary_final$upf_category == "partly upf"]),
    paste0("n = ", summary_final$pct_keyhole_yes[summary_final$upf_category == "upf"])
 )
)
```

Order the keyhole categories by defining factor levels

```{r}
plot_data$keyhole <- factor(plot_data$keyhole, levels = c("x", "no", "yes"))
```

3)  Create the stacked bar chart

```{r}
p1 <- ggplot(plot_data, aes(x = upf_category, y = perc, fill = keyhole)) +
  geom_bar(stat = "identity", width = 0.35, color = "white") +
  scale_y_continuous(expand = expansion(mult = c(0, 0.1))
  ) +
  
  scale_fill_manual(
    name = "GEF",
    values = c("yes" = "#6B8E7F", "no" = "#9C6B6B", "x" = "snow4"),
    labels = c("yes" = "Yes", "no" = "No", "x" = "Not classified")
  ) +
  labs(
    x = "",
    y = "Percentage (%)"
  ) + 
  
  geom_text(
    data = label_df,
    aes(x = upf_category, y = perc_top, label = label),
    inherit.aes = FALSE,
    vjust = -0.6,
    size = 4
  ) +
  
  geom_text(
    data = right_label_df,
    aes(x = upf_category, y = perc_point, label = label),
    inherit.aes = FALSE,
    hjust = -0.5,
    size = 3.5
  ) +
  
  theme_minimal(base_size = 12)+
  theme(
    plot.background = element_rect(fill = "white", color = NA),
    panel.background = element_rect(fill = "white", color = NA),
    axis.line = element_line(color = "black", linewidth = 0.1),
    legend.position = "right",
    legend.text = element_text(size = 10),
    panel.grid= element_blank(),
    axis.text.x = element_text(angle = 45, hjust = 1, size = 10)
  )


print(p1)
```

Save as image

```{r}
ggsave(
  filename = file.path(dir_output, "Stacked_barplot.png"),
  plot = p1,
  width = 16,
  height = 12,
  unit = "cm",
  dpi = 300
)

```


```{r}
ggsave(
  filename = file.path(dir_output, "Stacked_barplot.pdf"),
  plot = p1,
  width = 16,
  height = 12,
  unit = "cm",
  dpi = 300
)

```

Investigate which are the five top overlapping food groups

```{r}
upf_gef <- data2 %>% 
  filter(upf_category == "upf", 
         keyhole == "yes")

```

```{r}
top_foodgroups <- upf_gef %>% 
  count(subgroup, sort = TRUE)

```

```{r}
top5_foodgroups <- top_foodgroups %>% 
  slice_head(n =5)

top5_foodgroups
```

Export outputs to Excel

```{r}
write_xlsx(
  list(
    food_items = summary_final,
    Kappa = kappa_by_category
  ),
  path = file.path(dir_output, "Classification_kappa.xlsx")
)
```
