---
title: "Cleaning of raw background data"
author: "Torsten Sällström"
format: html
editor: visual
---

Starting files (Excel format):

RM_vuxna_background.xlsx = Demographic and anthropometric variables for RM adults RM_ungdom_background.xlsx = Demographic and anthropometric variables for RM ungdom RM_barn_background.xlsx = Demographic and anthropometric variables for RM barn

RM_ungdom_questionnaire.xlsx = Answer to questionnaire regarding physical activity in adolescents.

```{r}
#| label: setup
#| include: false

project_root <- normalizePath(file.path(getwd(), ".."))

dir_data <- file.path(project_root, "data")
dir_output <- file.path(project_root, "output")

dir.create(dir_output, showWarnings = FALSE, recursive = TRUE)

```

analysis code starts here

```{r message=FALSE, warning=FALSE}

library(openxlsx)
library(writexl)
library(dplyr)
library (tidyr)
library(stringr)
library(purrr)

library(childsds)

```

Import data

```{r}

vux_bg <- read.xlsx(file.path(dir_data,"RM_vuxna_background.xlsx"))
ungdom_bg <- read.xlsx(file.path(dir_data,"RM_ungdom_background.xlsx"))
barn_bg <- read.xlsx(file.path(dir_data,"RM_barn_background.xlsx"))

ungdom_pa <- read.xlsx(file.path(dir_data,"RM_ungdom_questionnaire_PA.xlsx"))

```

Standardize variables in background datasets

Adults

```{r}

vux_bg <- vux_bg %>% 
  mutate(
   person_ID = paste0("V", as.character(PersonId)),
   
   RM = "Vuxna",
                      
   sex = case_when(
   `Kön` == "Man" ~"M",
   `Kön`== "Kvinna" ~"F",
    TRUE ~NA_character_
    ),
   
  age_grp = case_when(
    `Ålder` >= 18 & `Ålder`<= 39 ~"18-39",
    `Ålder` >= 40 & `Ålder`<= 59 ~"40-59",
    `Ålder` >= 60 ~"60-80",
    TRUE ~NA_character_
  )
)


```

Calculate BMI

```{r}

vux_bg <- vux_bg %>% 
  mutate(
    
#Replace 0 with NA
`Längd,.cm`= na_if(`Längd,.cm`, 0),
`Vikt,.kg`= na_if(`Vikt,.kg`, 0),
`Vikt`= na_if(`Vikt`, 0),

#Height in meters
height_m = `Längd,.cm`/ 100, 
height_cm = `Längd,.cm`,

#Create one unified weight variable. When both weight measurement points exist, use the mean value. If only one measurement exist, this should be used. 
weight = case_when(
  !is.na(`Vikt,.kg`) & !is.na(`Vikt`) ~
    (`Vikt,.kg`+`Vikt`) / 2,
  !is.na(`Vikt,.kg`) ~`Vikt,.kg`,
  !is.na(`Vikt`) ~`Vikt`,
  TRUE ~NA_real_
),

#Calculate BMI
bmi = weight / (height_m ^2)
) %>% 
  rename(age = `Ålder`)

```

```{r}
summary(vux_bg$bmi)
```

Create a BMI class variable for adults

```{r}
vux_bg <- vux_bg %>% 
  mutate(bmi_class = case_when(
    bmi < 18.5 ~"undervikt",
    bmi >= 18.5 & bmi <=25.0 ~"normalvikt",
    bmi >25.0 & bmi <= 30.0 ~"övervikt",
    bmi > 30.0 ~"fetma"
  )) %>% 
  mutate(overweight = case_when(
    bmi_class %in% c("övervikt", "fetma") ~"yes",
    is.na(bmi_class) ~NA_character_,
    TRUE ~"no"
  )
)

```

```{r}
vux_bg %>% 
  count(bmi_class, sort = T)
```

Recode variable for leisure activity

```{r}
vux_bg %>% 
  count(Fritid, sort = T)
```

```{r}
vux_bg <- vux_bg %>% 
  mutate(
    leisure = case_when(
      Fritid == "Stilla" ~"sedentary",
      Fritid == "Måttlig" ~"light",
      Fritid == "Måttlig/Regelbunden" ~"moderate",
      Fritid == "Regelbunden" ~"hard",
      TRUE ~NA_character_
    )
  )
```

Recode smoking variable

```{r}
vux_bg %>% 
  count(Röker, sort = T)
```

```{r}
vux_bg <- vux_bg %>% 
  mutate(
    smoke = case_when(
      `Röker` == "Aldrig rökt" ~"never",
      `Röker` == "Har slutat" ~"former",
      `Röker` == "Vid enstaka tillfällen" ~"occasional",
      `Röker` == "Dagligen" ~"daily",
      TRUE ~NA_character_
    )
  )
```

Recode education variable

```{r}
vux_bg %>% 
  count(Högsta.utbildning, sort = T)
```

Recode into two categories

```{r}
vux_bg <- vux_bg %>% 
  mutate(
    edu_combined = case_when(
      str_detect(`Högsta.utbildning`, "Uppgift saknas") ~NA_character_,
      #<= 12 years
      str_detect(`Högsta.utbildning`, "\\bFörgymnasial\\b") ~"lower",
      str_detect(`Högsta.utbildning`, "\\bGymnasial\\b") ~"lower",
      
      #> 12 years
      str_detect(`Högsta.utbildning`, "Eftergymnasial") ~"higher",
      str_detect(`Högsta.utbildning`, "Forskarutbildning") ~"higher",
      
      TRUE ~NA_character_
    )
  )

```

```{r}
vux_bg %>% 
  count(edu_combined, sort = T)
```

Adolescents

```{r}
ungdom_bg <- ungdom_bg %>% 
  mutate(
    person_ID = paste0("U", as.character(Person_ID)),
    
    RM = "Ungdom",
                       
    sex = case_when(
    kön %in% c("M", "F") ~kön, 
    TRUE ~NA_character_
),

    age_grp = case_when(
     `Årskurs`== "Åk 5" ~"Ak5",
     `Årskurs`== "Åk 8" ~"Ak8",
     `Årskurs`== "Åk 2 gymn" ~"Ak2_g",
     TRUE ~NA_character_
),
    #Rename anthropometrics
    age = na_if(Age, 0),
    weight = na_if(Vikt, 0),
    height_m = na_if(Längd, 0),
    height_cm = height_m * 100,

    #BMI
    bmi = weight / (height_m ^2)
) 
```

Recode education variable

```{r}
ungdom_bg <- ungdom_bg %>% 
  mutate(
    edu_combined = case_when(
      Högsta.utb.i.hushållet.två.kategorier == ">12" ~"higher",
      Högsta.utb.i.hushållet.två.kategorier == "<=12" ~"lower",
      TRUE ~NA_character_
    )
  )
```

Calculate BMIz-score for individuals under 18

```{r}

#Create temporary sex-variable
ungdom_bg <- ungdom_bg %>% 
  mutate(sex_temp = case_when(
    sex == "M" ~"male",
    sex == "F" ~"female"
  ))

```

Filter adolescents to include only cases younger than 18 years

```{r}
ungdom_bmi <- ungdom_bg %>% 
  filter(age < 18) %>% 
  mutate(bmi_z = sds(age = age,
                     sex = sex_temp,
                     item = "bmi",
                     value = bmi,
                     ref = cole_lobstein.ref)) %>% 
  dplyr::select(person_ID, bmi_z)

```

Join back bmi_z to original data

```{r}
ungdom_bg <- ungdom_bg %>% 
  left_join(ungdom_bmi, by = "person_ID")

```

```{r}
ungdom_bg <- ungdom_bg %>% 
  mutate(bmi_class = case_when(
    
    age >= 2 & age < 18 & is.na(bmi_z) ~NA_character_,
    
    age < 18 & sex == "M" & bmi_z >=2.288 ~"fetma",
    age < 18 & sex == "M" & bmi_z <2.288 & bmi_z >= 1.310 ~"övervikt",
    age < 18 & sex == "M" & bmi_z <1.310 & bmi_z >= -1.014 ~"normalvikt",
    age < 18 & sex == "M" & bmi_z < -1.014 ~"undervikt",
    
    age < 18 & sex == "F" & bmi_z >=2.192 ~"fetma",
    age < 18 & sex == "F" & bmi_z <2.192 & bmi_z >= 1.244 ~"övervikt",
    age < 18 & sex == "F" & bmi_z <1.244 & bmi_z >= -0.975 ~"normalvikt",
    age < 18 & sex == "F" & bmi_z < -0.975 ~"undervikt",
    
    age >= 18 & is.na(bmi) ~NA_character_,
    
    age >= 18 & bmi < 18.5 ~"undervikt",
    age >= 18 & bmi >= 18.5 & bmi <=25.0 ~"normalvikt",
    age >= 18 & bmi >25.0 & bmi <= 30.0 ~"övervikt",
    age >= 18 & bmi > 30.0 ~"fetma"
  )
)

```

```{r}
ungdom_bg %>% 
  group_by(age_grp) %>% 
  count(bmi_class, sort = T)
```

Create overweight variable

```{r}
ungdom_bg <- ungdom_bg %>% 
  mutate(
    overweight = if_else(
      bmi_class %in% c("övervikt", "fetma"),
      "yes", 
      if_else(is.na(bmi_class), NA_character_, "no")
    )
  )
```

Create and join physical activity variables for adolescents

```{r}

ungdom_pa <- ungdom_pa %>% 
  mutate(
    person_ID = paste0("U", as.character(PersonId)), 
  )%>% 
    rename_with(
      ~"leisure",
      starts_with("Vilken.av.följande.")
    )
  
```

```{r}
ungdom_pa %>% 
  count(leisure, sort = T)
```

Recode variable leisure

```{r}
ungdom_pa <- ungdom_pa %>% 
  mutate(
    leisure = case_when(
      str_detect(leisure, "Läsa") ~"sedentary",
      str_detect(leisure, "lätt aktivitet") ~"light",
      str_detect(leisure, "motionsidrott") ~"moderate",
      str_detect(leisure, "tävlingsidrott") ~"hard",
      TRUE ~NA_character_
    )
  )
```

Join back to ungdom_bg

```{r}
ungdom_bg <- ungdom_bg %>% 
  left_join(
    ungdom_pa %>% dplyr::select(person_ID, leisure),
    by = "person_ID"
  )
```

Children

```{r}
sum(is.na(barn_bg$`ålder_startdatum`))
sum(is.na(barn_bg$`ålder_mätdatum_år`))
```

Are there any cases that miss values for ålder_mätdatum_år but has values for ålder_startdatum?

```{r}
sum(
  is.na(barn_bg$ålder_mätdatum_år) &
  !is.na(barn_bg$ålder_startdatum)
)

```

Create two age variables:

age = ålder_startdatum. This is used as age variable throughout. age_bmiz = "ålder_mätdatum_år" when it exists. Otherwise, use values from "ålder_startdatum". This is used only to calculate bmi_z score.

```{r}
barn_bg <- barn_bg %>% 
  mutate(
    age_bmiz = coalesce(ålder_mätdatum_år, ålder_startdatum),
    age = coalesce(ålder_startdatum, ålder_mätdatum_år)
  )
```

```{r}
barn_bg <- barn_bg %>% 
  mutate(
    person_ID = paste0("B", as.character(personid)),
    sex = case_when(
      `Kön.(från.pnr)`== "Flicka" ~"F",
      `Kön.(från.pnr)`== "Pojke" ~"M",
      TRUE ~NA_character_
    ),
    
    RM = "Barn",
    
    age_grp = case_when(
      `åldersgrupp`== "18 månader" ~"18_manad",
      `åldersgrupp`== "4 år" ~"4_ar",
      TRUE ~NA_character_
    ),
    
    #Height in centimeters
    height_cm = längd_cm,
    
    #Height in meters
    height_m = längd_cm / 100,

    #Weight
    weight = as.numeric(gsub(",", ".", vikt_kg)),
    
    #0 -> NA
    weight = if_else(weight == 0, NA_real_, weight),
    
    #BMI calculated
    bmi = weight / (height_m ^2)
  ) 

```

Create temporary sex variable for calculation of bmi_z

```{r}
barn_bg <- barn_bg %>% 
  mutate(sex_temp = case_when(
    sex == "M" ~"male",
    sex == "F" ~"female"
  ))
```

```{r}
barn_bmi <- barn_bg %>% 
  filter(age_grp == "4_ar") %>% 
  mutate(bmi_z = sds(age = age_bmiz,
                     sex = sex_temp,
                     item = "bmi",
                     value = bmi,
                     ref = cole_lobstein.ref)) %>% 
  dplyr::select(person_ID, bmi_z)

```

Join back

```{r}
barn_bg <- barn_bg %>% 
  left_join(
    barn_bmi, by = "person_ID"
  )

```

Check bmi_z calculation:

```{r}
problems <- barn_bg %>% 
  filter(
    age_grp == "4_ar",
    !is.na(bmi),
    !is.na(height_m),
    !is.na(age_bmiz),
    is.na(bmi_z)
  ) %>% 
  dplyr::select(
    age_grp,
    age,
    age_bmiz,
    bmi,
    height_m,
    bmi_z
  )

problems
```

These cases are kept as NA, as the measurements were when they were much younger.

Create bmi class for 4 year olds only

```{r}
barn_bg <- barn_bg %>% 
  mutate(bmi_class = case_when(
    
    age_grp == "4_ar" & is.na(bmi_z) ~NA_character_,
    
    age_grp == "4_ar" & sex == "M" & bmi_z >=2.288 ~"fetma",
    age_grp == "4_ar" & sex == "M" & bmi_z <2.288 & bmi_z >= 1.310 ~"övervikt",
    age_grp == "4_ar" & sex == "M" & bmi_z <1.310 & bmi_z >= -1.014 ~"normalvikt",
    age_grp == "4_ar" & sex == "M" & bmi_z < -1.014 ~"undervikt",
    
    age_grp == "4_ar" & sex == "F" & bmi_z >=2.192 ~"fetma",
    age_grp == "4_ar" & sex == "F" & bmi_z <2.192 & bmi_z >= 1.244 ~"övervikt",
    age_grp == "4_ar" & sex == "F" & bmi_z <1.244 & bmi_z >= -0.975 ~"normalvikt",
    age_grp == "4_ar" & sex == "F" & bmi_z < -0.975 ~"undervikt"
  )
)

```

Crate overweight variable

```{r}
barn_bg <- barn_bg %>% 
  mutate(
    overweight = if_else(
      bmi_class %in% c("övervikt", "fetma"),
      "yes", 
      if_else(is.na(bmi_class), NA_character_, "no")
    )
  )
```

Trim

```{r}
ungdom_bg$bmi_class <- trimws(ungdom_bg$bmi_class)
barn_bg$bmi_class <- trimws(barn_bg$bmi_class)

```

Check

```{r}
barn_bg %>% 
  count(bmi_class, sort = TRUE)
```

Inspect screentime variables

```{r}
barn_bg %>% 
  count(skärm_vardag, sort = TRUE)
```

```{r}
barn_bg %>% 
  count(skärm_helgdag, sort = TRUE)
```

Recode to minutes (numeric)

```{r}
#Helper function
recode_screen_minutes <- function(x) {
  case_when(
    is.na(x) ~NA_real_,
    x == "Vet ej" ~NA_real_,
    x == "ingen enkät" ~NA_real_,
    x == ".a" ~NA_real_,
    x == "Ingen tid" ~0,
    x == "15 minuter" ~15,
    x == "30 minuter" ~30,
    x == "45 minuter" ~45,
    x == "1 timme" ~60,
    x == "1,5 timmar" ~90,
    x == "2 timmar" ~120,
    x == "2,5 timmar" ~150,
    x == "3 timmar" ~180,
    x == "Mer än 3,5 timmar" ~210
  )
}
```

Apply function to screentime variables

```{r}
barn_bg <- barn_bg %>% 
  mutate(
    screentime_weekday = recode_screen_minutes(skärm_vardag),
    screentime_weekend = recode_screen_minutes(skärm_helgdag)
  )

```

```{r}
table(barn_bg$screentime_weekday, useNA = "ifany")
```

Calculate a weighted average

```{r}
barn_bg <- barn_bg %>% 
  mutate(
    screentime_combined_min = case_when(
      !is.na(screentime_weekday) & !is.na(screentime_weekend) ~
        (5 * screentime_weekday + 2 * screentime_weekend) / 7,
         
      !is.na(screentime_weekday) & is.na(screentime_weekend) ~
        screentime_weekday,
      
      is.na(screentime_weekday) & !is.na(screentime_weekend) ~
        screentime_weekend, 
      
      TRUE ~NA_real_
    )
  )

```

Convert to hours

```{r}
barn_bg <- barn_bg %>% 
    mutate(
        screentime_combined = screentime_combined_min / 60
    )
```

Rename education variable

```{r}
sort(unique(barn_bg$`Högsta.utb.hushållet`))
```

```{r}
barn_bg <- barn_bg %>% 
  rename(
    edu_combined = `Högsta.utb.hushållet`
  ) %>% 
  mutate(edu_combined = case_when(
    edu_combined == ">12 år" ~"higher",
    edu_combined == "≤ 12 år" ~"lower",
    TRUE ~edu_combined
  )
)
```

Define variables to keep

```{r}
select_safe <- function(df, vars) {
  df %>% dplyr::select(any_of(vars))
}

vars_keep_vux <- c(
  "person_ID", "sex", "age", "RM", "age_grp", "weight", "height_m", "height_cm", "bmi", "bmi_class", "overweight", "edu_combined", "leisure", "smoke"
)

vars_keep_ungdom <- c(
  "person_ID", "sex", "age", "RM", "age_grp", "weight", "height_m", "height_cm", "bmi","bmi_z",  "bmi_class", "overweight", "edu_combined", "leisure"
)

vars_keep_barn <- c(
  "person_ID", "sex", "age", "RM", "age_grp", "weight", "height_m", "height_cm", "bmi","bmi_z",  "bmi_class", "overweight", "edu_combined", "screentime_combined"
)

```

Select

```{r}
vux_bg_clean <- select_safe(vux_bg, vars_keep_vux)
ungdom_bg_clean <- select_safe(ungdom_bg, vars_keep_ungdom)
barn_bg_clean <- select_safe(barn_bg, vars_keep_barn)

```

Save cleaned intake datasets as rds files

```{r}
saveRDS(vux_bg_clean, file.path(dir_output, "RM_vux_bg_clean.rds"))
saveRDS(ungdom_bg_clean, file.path(dir_output, "RM_ungdom_bg_clean.rds"))
saveRDS(barn_bg_clean, file.path(dir_output, "RM_barn_bg_clean.rds"))
```
