---
title: "Cleaning of raw intake data"
author: "Torsten Sällström"
format: html
editor: source
---

Starting files:

RM_vuxna_raw.xlsx = Riksmaten adults raw intake data RM_ungdom_raw.xlsx = Riksmaten adolescents raw intake data RM_barn_raw.xlsx = Riksmaten young children raw intake data

RM_vuxna_bg_clean.rds = Demographic and anthropometric variables for RM adults RM_ungdom_bg_clean.rds = Demographic and anthropometric variables for RM ungdom RM_barn_bg_clean.rds = Demographic and anthropometric variables for RM barn

Food_items.xlsx = List of all food items with their classification according to UPF and GEF 
Salt_sugar.xlsx = List of all food items and their content of salt and added sugar per 100 g

RDS files are created in cleaning step 02.1.


```{r}
#| label: setup
#| include: false

project_root <- normalizePath(file.path(getwd(), ".."))

dir_data <- file.path(project_root, "data")
dir_output <- file.path(project_root, "output")

dir.create(dir_output, showWarnings = F, recursive = T)

```

analysis code starts here


```{r message=FALSE, warning=FALSE}

library(openxlsx)
library(writexl)
library(dplyr)
library (tidyr)
library(stringr)
library(purrr)

```

Import data

```{r}
vux1 <- read.xlsx(file.path(dir_data,"RM_vuxna_raw.xlsx"))
ungdom1 <- read.xlsx(file.path(dir_data,"RM_ungdom_raw.xlsx"))
barn1 <- read.xlsx(file.path(dir_data,"RM_barn_raw.xlsx"))

vux_bg <- readRDS(file.path(dir_data,"RM_vux_bg_clean.rds"))
ungdom_bg <- readRDS(file.path(dir_data,"RM_ungdom_bg_clean.rds"))
barn_bg <- readRDS(file.path(dir_data,"RM_barn_bg_clean.rds"))

food_items <- read.xlsx(file.path(dir_data,"Food_items.xlsx"))
salt_sugar <- read.xlsx(file.path(dir_data,"Salt_sugar.xlsx"))

```

Rename selected variables

Adults

```{r}
vux <- vux1 %>% 
  rename(
    person_ID = PersonId,
    food_ID = LivsmedelsId,
    food = Livsmedelsnamn,
    gram = Livsmedel.mängd,
    day_nr = Dag,
    kcal = `Ener.(kcal)`,
    carbs = `Kolh.(g)`,
    fat = `Fett.(g)`,
    protein = `Prot.(g)`,
    fiber = `Fibe.(g)`,
    wholegrain = `Fullk/tot.(g)`,
    SFA = `Mfet.(g)`,
    vitA = `VitA.(RE/µg)`,
    vitD = `VitD.(µg)`,
    vitE = `VitE.(mg)`,
    folate = `Folat.(µg)`,
    vitB12 = `VitB12.(µg)`,
    vitC = `VitC.(mg)`,
    p = `P.(mg)`,
    fe = `Fe.(mg)`,
    ca = `Ca.(mg)`,
    k = `K.(mg)`,
    mg = `Mg.(mg)`,
    se = `Se.(µg)`,
    zn = `Zn.(mg)`
  )
```

Adolescents

Convert Energy variable from kJ to kcal

```{r}
ungdom1 <- ungdom1 %>% 
  mutate(
    kcal = `Ener.(kJ)` / 4.184
  )
```

```{r}
ungdom <- ungdom1 %>% 
  rename(
    person_ID = PersonId,
    food_ID = LivsmedelsId,
    food = Livsmedelsnamn,
    gram = Livsmedel.mängd,
    day_nr = Dag,
    carbs = `Kolh.(g)`,
    fat = `Fett.(g)`,
    protein = `Prot.(g)`,
    fiber = `Fibe.(g)`,
    wholegrain = `Fullk/tot.(g)`,
    SFA = `Mfet.(g)`,
    naCl = `NaCl.(g)`,
    vitA = `VitA.(RE/µg)`,
    vitD = `VitD.(µg)`,
    vitE = `VitE.(mg)`,
    folate = `Folat.(µg)`,
    vitB12 = `VitB12.(µg)`,
    vitC = `VitC.(mg)`,
    p = `P.(mg)`,
    fe = `Fe.(mg)`,
    ca = `Ca.(mg)`,
    k = `K.(mg)`,
    mg = `Mg.(mg)`,
    se = `Se.(µg)`,
    zn = `Zn.(mg)`
  )
```

Young children

```{r}

barn <- barn1 %>% 
  rename(
    person_ID = PersonId,
    food_ID = LivsmedelsId,
    food = Livsmedelsnamn,
    gram = Livsmedel.mängd,
    day_nr = Dag,
    kcal = `Ener.(kcal)`,
    carbs = `Kolh.(g)`,
    fat = `Fett.(g)`,
    protein = `Prot.(g)`,
    fiber = `Fibe.(g)`,
    sugar_added = `Socker/t.(g)`,
    wholegrain = `Fullk/tot.(g)`,
    SFA = `Mfet.(g)`,
    naCl = `NaCl.(g)`,
    vitA = `VitA.(RE/µg)`,
    vitD = `VitD.(µg)`,
    vitE = `VitE.(mg)`,
    folate = `Folat.(µg)`,
    vitB12 = `VitB12.(µg)`,
    vitC = `VitC.(mg)`,
    p = `P.(mg)`,
    fe = `Fe.(mg)`,
    ca = `Ca.(mg)`,
    k = `K.(mg)`,
    mg = `Mg.(mg)`,
    se = `Se.(µg)`,
    zn = `Zn.(mg)`
  )
```

Convert food_ID to character

```{r}
vux$food_ID <- as.character(vux$food_ID)
ungdom$food_ID <- as.character(ungdom$food_ID)
barn$food_ID <- as.character(barn$food_ID)
food_items$food_ID <- as.character(food_items$food_ID)
salt_sugar$food_ID <- as.character(salt_sugar$food_ID)

```

Join values for NaCl and Added sugars (for Adults) and Added sugars (for Adolescents). Values are expressed as g/100 g. Calculate intake.

```{r}

#Create a key
salt_sugar_key <- salt_sugar %>% 
  dplyr::select(food_ID, salt_per100, sugar_added_per100) %>% 
  distinct(food_ID, .keep_all = T)

#Check for NA
salt_sugar_key %>% 
  summarise(
    any_na_salt = any(is.na(salt_per100)),
    any_na_sugar = any(is.na(sugar_added_per100))
  )

```

```{r}
#For adults

vux <- vux %>% 
  left_join(salt_sugar_key, by = "food_ID") %>% 
  mutate(
    naCl = gram * salt_per100 / 100,
    sugar_added = gram * sugar_added_per100 / 100
  )

#For adolescents
ungdom <- ungdom %>% 
  left_join(salt_sugar_key, by = "food_ID") %>% 
  mutate(
    sugar_added = gram * sugar_added_per100 / 100
  )

```

Initial cleaning

```{r}
#1 Prepare the food_items dataframe

food_lookup <- food_items %>% 
  dplyr::select(
    food_ID, subgroup, bev,
    keyhole, keyhole_ratio,
    ratio_upf_low, ratio_upf_mid_vux, ratio_upf_mid_barn, ratio_upf_mid_ungdom, ratio_upf_up
  ) %>% 
  distinct(food_ID, .keep_all = TRUE)

```

```{r}
#2. Put datasets in one list
datasets <- list(
  vux = vux,
  ungdom = ungdom,
  barn = barn
)

prefixes <- c(
  vux = "V",
  ungdom = "U",
  barn = "B"
)

rm_labels <- c(
  vux = "Vuxna",
  ungdom = "Ungdom",
  barn = "Barn"
)
```


```{r}
#3. Function to apply to vux / ungdom / barn
clean_and_join <- function(df, prefix, rm_label, food_lookup) {
  
  mid_ratio_col <- switch(
    rm_label,
    "Vuxna" = "ratio_upf_mid_vux",
    "Barn" = "ratio_upf_mid_barn",
    "Ungdom" = "ratio_upf_mid_ungdom",
    stop("Unknown rm_label: ", rm_label)
  )
  
  df %>% 
    mutate(
      person_ID = paste0(prefix, as.character(person_ID)),
      RM = rm_label
    ) %>% 
    left_join(food_lookup, by = "food_ID") %>% 
    group_by(person_ID) %>% 
    mutate(
      day_count = n_distinct(day_nr)
    ) %>% 
    ungroup() %>% 
    mutate(
      #GEF kcal
      keyhole_kcal = kcal * keyhole_ratio,
      
      #UPF kcal by scenario
      upf_low_kcal = kcal * ratio_upf_low,
      upf_mid_kcal = kcal * .data[[mid_ratio_col]],
      upf_up_kcal = kcal * ratio_upf_up
    )
}
```

```{r}
#4. Apply function
datasets_clean <- imap(
  datasets, 
  ~clean_and_join(
    df = .x, 
    prefix = prefixes[[.y]], 
    rm_label = rm_labels[[.y]],
    food_lookup = food_lookup
  )
)
```

```{r}
#5. Write back to objects

vux <- datasets_clean$vux
ungdom <- datasets_clean$ungdom
barn<- datasets_clean$barn
```

Spot-check, using food_ID 241 (Råstekt potatis)

```{r}
#inspect in food_items

food_items %>% 
  filter(food_ID == 241) %>% 
  select(
    food_ID,
    food,
    ratio_upf_mid_vux,
    ratio_upf_mid_barn
  )
```
```{r}
#Inspect the first five matching rows in vux
vux %>% 
  filter(food_ID == 241) %>% 
  select(
    food_ID,
    food,
    subgroup,
    kcal,
    ratio_upf_mid_vux,
    upf_mid_kcal
  ) %>% 
  slice_head(n = 5)

```

```{r}
#Inspect the first five matching rows in barn
barn %>% 
  filter(food_ID == 6405) %>% 
  select(
    food_ID,
    food,
    subgroup,
    kcal,
    ratio_upf_mid_barn,
    upf_mid_kcal
  ) %>% 
  slice_head(n = 5)

```


Sanity check

```{r}
vux %>% distinct(person_ID, day_count) %>% 
  count(day_count)
```

Select vars to add from background datasets

```{r}
vars_keep <- c(
  "person_ID",
  "sex",
  "age_grp"
)
```

```{r}
#Helper function
add_bg_vars <- function(intake_df, bg_df, vars_keep) {
  bg_key <- bg_df %>% 
    dplyr::select(any_of(c("person_ID", vars_keep))) %>% 
    distinct(person_ID, .keep_all = T)
  
  intake_df %>% 
    left_join(bg_key, by = "person_ID")
}

```

Apply to each intake dataset

```{r}
vux_final <- add_bg_vars(vux, vux_bg, vars_keep)
ungdom_final <- add_bg_vars(ungdom, ungdom_bg, vars_keep)
barn_final <- add_bg_vars(barn, barn_bg, vars_keep)

```

Select variables to keep in the clean dataset going forward

```{r}
vars_final <- c(
  "person_ID",
  "sex",
  "RM",
  "age_grp",
  
  "food_ID",
  "food",
  "bev",
  "subgroup",
  "day_nr",
  "day_count",
  
  "gram",
  "kcal",
  "keyhole",
  "keyhole_ratio",
  "keyhole_kcal",
  "upf_low_kcal",
  "upf_mid_kcal",
  "upf_up_kcal",
  
  "carbs",
  "fat",
  "protein",
  "fiber",
  "wholegrain",
  "sugar_added",
  "SFA",
  "vitA",
  "vitD",
  "vitE",
  "vitC",
  "vitB12",
  "folate",
  "fe",
  "ca",
  "k",
  "mg",
  "se",
  "zn",
  "naCl"
  
)
```

Helper function to keep variables

```{r}
keep_vars <- function(df, vars_final) {
  df %>% dplyr::select(any_of(vars_final))
}

```

Apply

```{r}
vux_final <- keep_vars(vux_final, vars_final)
ungdom_final <- keep_vars(ungdom_final, vars_final)
barn_final <- keep_vars(barn_final, vars_final)

```

Save cleaned intake datasets as rds files

```{r}
saveRDS(vux_final, file.path(dir_data, "RM_vux_clean.rds"))
saveRDS(ungdom_final, file.path(dir_data, "RM_ungdom_clean.rds"))
saveRDS(barn_final, file.path(dir_data, "RM_barn_clean.rds"))
```
