---
title: "UPF intake by food group, UPF E% and GEF E%"
author: "Torsten Sällström"
format: html
editor: visual
---

In this script we: 1) summarize the foodgroups that contribute most kcal from UPF on population level. 2) calculate the dietary share of food items classified as GEF 3) collapses the dataset to one row per observatio and calculates E% UPF according to three classification scenarios.

Starting with the clean intake datasets created in step 02.1

```{r}
#| label: setup
#| include: false

project_root <- normalizePath(file.path(getwd(), ".."))

dir_data <- file.path(project_root, "data")
dir_output <- file.path(project_root, "output")

dir.create(dir_output, showWarnings = FALSE, recursive = TRUE)

```

analysis code starts here

```{r message=FALSE, warning=FALSE}

library(openxlsx)
library(writexl)
library(dplyr)
library (tidyr)
library(stringr)
library(purrr)

```

Import data (choose dataset)

```{r}
vux <- readRDS(file.path(dir_data,"RM_vux_clean.rds"))
ungdom <- readRDS(file.path(dir_data, "RM_ungdom_clean.rds"))
barn <- readRDS(file.path(dir_data, "RM_barn_clean.rds"))

```

Create three separate lists of unique food_ID

```{r}
vux_foods <- vux %>% 
  select(food_ID, RM, food) %>% 
  distinct() %>% 
  arrange(food_ID)

ungdom_foods <- ungdom %>% 
  select(food_ID, RM, food) %>% 
  distinct() %>% 
  arrange(food_ID)

barn_foods <- barn %>% 
  select(food_ID, RM, food) %>% 
  distinct() %>% 
  arrange(food_ID)
```

Compare side by side

```{r}
food_compare <- full_join(
  vux_foods %>% 
    rename(
      RM_vux = RM,
      food_vux = food
    ),
  ungdom_foods %>% 
    rename(
      RM_ungdom = RM,
      food_ungdom = food
    ),
  by = "food_ID"
) %>% 
  full_join(
    barn_foods %>% 
      rename(
        RM_barn = RM,
        food_barn = food
      ),
    by = "food_ID"
  ) %>% 
  arrange(food_ID)

```

Inspect

```{r}
View(food_compare)
```

Combine datasets

```{r}
# Combine datasets
data <- bind_rows(vux, barn, ungdom) 

```

Create list of unique food items reported

```{r}
food_lookup <- data %>%
  group_by(food_ID) %>% 
  summarise(food = first(food), .groups = "drop")

```

Check for cells with NA for food intake data per dataset

```{r}

check_cols <- setdiff(names(data), c("RM", "food_ID", "food"))

na_summary <- data %>% 
    group_by(RM, food_ID, food) %>% 
    summarise(
      across(all_of(check_cols), ~sum(is.na(.)), .names = "na_{col}"),
      .groups = "drop"
    ) %>% 
    pivot_longer(
      cols = starts_with("na_"),
      names_to = "variable",
      values_to = "n_NA"
    ) %>% 
    mutate(variable = sub("^na_", "", variable)) %>% 
    filter(n_NA > 0) %>% 
    arrange(RM, food_ID)
  
na_summary
```

Clean subgroups to prevent duplicates

```{r}

data <- data %>%
 mutate( 
   subgroup = str_trim(str_to_title(subgroup))) # Trim whitespace and standardize case

summary(data$subgroup)

n_distinct(data$subgroup)
unique(data$subgroup)

```

MIDDLE-BOUND

Population-level E% UPF by food group, stratified by RM

```{r}
upf_by_foodgroup <- data %>% 
  filter(!is.na(subgroup)) %>% 
  group_by(RM) %>% 
  mutate(
    total_kcal_RM = sum(kcal, na.rm = TRUE)
  ) %>% 
  group_by(RM, subgroup) %>% 
  summarise(
    upf_mid_kcal_subgroup = sum(upf_mid_kcal, na.rm = TRUE),
    total_kcal_RM = first(total_kcal_RM),
    .groups = "drop"
  ) %>% 
  mutate(
    upf_mid_E_pct = 100 * upf_mid_kcal_subgroup / total_kcal_RM
  )

```

Population-level E% UPF by food group, stratified by RM and age-groups

```{r}
upf_by_foodgroup_and_agegroup <- data %>% 
  filter(!is.na(subgroup), !is.na(age_grp)) %>% 
  group_by(RM, age_grp) %>% 
  mutate(
    total_kcal_agegroup = sum(kcal, na.rm = TRUE)
  ) %>% 
  group_by(RM, age_grp, subgroup) %>% 
  summarise(
    upf_mid_kcal_subgroup = sum(upf_mid_kcal, na.rm = TRUE),
    total_kcal_agegroup = first(total_kcal_agegroup),
    .groups = "drop"
  ) %>% 
  mutate(
    upf_mid_E_pct = 100 * upf_mid_kcal_subgroup / total_kcal_agegroup
  )

```

Population-level E% UPF by subgroup, stratified by RM x sex

Denominator: total kcal within each RM x sex

```{r}
upf_by_foodgroup_and_sex <- data %>% 
  filter(
    !is.na(subgroup),
    !is.na(sex)
  ) %>% 
  group_by(RM, sex) %>% 
  mutate(
    total_kcal_grp = sum(kcal, na.rm = TRUE)
  ) %>% 
  group_by(RM, sex, subgroup) %>% 
  summarise(
    upf_mid_kcal_subgroup = sum(upf_mid_kcal, na.rm = TRUE),
    total_kcal_grp = first(total_kcal_grp),
    .groups = "drop"
  ) %>% 
  mutate(
    upf_mid_E_pct = 100 * upf_mid_kcal_subgroup / total_kcal_grp
  )
  
```

Population-level E% UPF by subgroup, stratified by RM x age_grp x sex

```{r}
upf_by_foodgroup_age_and_sex <- data %>% 
  filter(
    !is.na(subgroup),
    !is.na(age_grp),
    !is.na(sex)
  ) %>% 
  group_by(RM, age_grp, sex) %>% 
  mutate(
    total_kcal_grp = sum(kcal, na.rm = TRUE)
  ) %>% 
  group_by(RM, age_grp, sex, subgroup) %>% 
  summarise(
    upf_mid_kcal_subgroup = sum(upf_mid_kcal, na.rm = TRUE),
    total_kcal_grp = first(total_kcal_grp),
    .groups = "drop"
  ) %>% 
  mutate(
    upf_mid_E_pct = 100 * upf_mid_kcal_subgroup / total_kcal_grp
  )

```

Validation step

```{r}
upf_by_foodgroup_age_and_sex %>% 
  group_by(RM, age_grp, sex) %>% 
  summarise(total_E_pct = sum(upf_mid_E_pct),
   .groups = "drop")

upf_by_foodgroup_age_and_sex
```

Export foodgroup-results to one excelfile

```{r}

wb_groups <- createWorkbook()

addWorksheet(wb_groups, "by RM only")
writeData(wb_groups, "by RM only", upf_by_foodgroup)

addWorksheet(wb_groups, "by agegroup")
writeData(wb_groups, "by agegroup", upf_by_foodgroup_and_agegroup)

addWorksheet(wb_groups, "by_sex")
writeData(wb_groups, "by_sex", upf_by_foodgroup_and_sex)

addWorksheet(wb_groups, "by age and sex")
writeData(wb_groups, "by age and sex", upf_by_foodgroup_age_and_sex)

saveWorkbook(wb_groups, 
             file = file.path(dir_output, "Foodgroups_results_mid.xlsx"),
             overwrite = TRUE
          )

```

LOWER BOUND

Population-level E% UPF by foodgroup, stratified by RM

```{r}
upf_by_foodgroup <- data %>% 
  filter(!is.na(subgroup)) %>% 
  group_by(RM) %>% 
  mutate(
    total_kcal_RM = sum(kcal, na.rm = TRUE)
  ) %>% 
  group_by(RM, subgroup) %>% 
  summarise(
    upf_low_kcal_subgroup = sum(upf_low_kcal, na.rm = TRUE),
    total_kcal_RM = first(total_kcal_RM),
    .groups = "drop"
  ) %>% 
  mutate(
    upf_low_E_pct = 100 * upf_low_kcal_subgroup / total_kcal_RM
  )

```

Population-level E% UPF by foodgroup, stratified by RM and age-groups

```{r}
upf_by_foodgroup_and_agegroup <- data %>% 
  filter(!is.na(subgroup), !is.na(age_grp)) %>% 
  group_by(RM, age_grp) %>% 
  mutate(
    total_kcal_agegroup = sum(kcal, na.rm = TRUE)
  ) %>% 
  group_by(RM, age_grp, subgroup) %>% 
  summarise(
    upf_low_kcal_subgroup = sum(upf_low_kcal, na.rm = TRUE),
    total_kcal_agegroup = first(total_kcal_agegroup),
    .groups = "drop"
  ) %>% 
  mutate(
    upf_low_E_pct = 100 * upf_low_kcal_subgroup / total_kcal_agegroup
  )

```

Population-level E% UPF by subgroup, stratified by RM x sex

Denominator: total kcal within each RM x sex

```{r}
upf_by_foodgroup_and_sex <- data %>% 
  filter(
    !is.na(subgroup),
    !is.na(sex)
  ) %>% 
  group_by(RM, sex) %>% 
  mutate(
    total_kcal_grp = sum(kcal, na.rm = TRUE)
  ) %>% 
  group_by(RM, sex, subgroup) %>% 
  summarise(
    upf_low_kcal_subgroup = sum(upf_low_kcal, na.rm = TRUE),
    total_kcal_grp = first(total_kcal_grp),
    .groups = "drop"
  ) %>% 
  mutate(
    upf_low_E_pct = 100 * upf_low_kcal_subgroup / total_kcal_grp
  )
  
```

Population-level E% UPF by subgroup, stratified by RM x age_grp x sex

```{r}
upf_by_foodgroup_age_and_sex <- data %>% 
  filter(
    !is.na(subgroup),
    !is.na(age_grp),
    !is.na(sex)
  ) %>% 
  group_by(RM, age_grp, sex) %>% 
  mutate(
    total_kcal_grp = sum(kcal, na.rm = TRUE)
  ) %>% 
  group_by(RM, age_grp, sex, subgroup) %>% 
  summarise(
    upf_low_kcal_subgroup = sum(upf_low_kcal, na.rm = TRUE),
    total_kcal_grp = first(total_kcal_grp),
    .groups = "drop"
  ) %>% 
  mutate(
    upf_low_E_pct = 100 * upf_low_kcal_subgroup / total_kcal_grp
  )

```

Validation step

```{r}
upf_by_foodgroup_age_and_sex %>% 
  group_by(RM, age_grp, sex) %>% 
  summarise(total_E_pct = sum(upf_low_E_pct),
            .groups = "drop")

upf_by_foodgroup_age_and_sex
```

Export foodgroup-results to one excelfile

```{r}

wb_groups <- createWorkbook()

addWorksheet(wb_groups, "by RM only")
writeData(wb_groups, "by RM only", upf_by_foodgroup)

addWorksheet(wb_groups, "by agegroup")
writeData(wb_groups, "by agegroup", upf_by_foodgroup_and_agegroup)

addWorksheet(wb_groups, "by_sex")
writeData(wb_groups, "by_sex", upf_by_foodgroup_and_sex)

addWorksheet(wb_groups, "by age and sex")
writeData(wb_groups, "by age and sex", upf_by_foodgroup_age_and_sex)

saveWorkbook(wb_groups, 
             file = file.path(dir_output, "Foodgroups_results_low.xlsx"),
             overwrite = TRUE
          )

```

UPPER-BOUND

Population-level E% UPF by foodgroup, stratified by RM

```{r}
upf_by_foodgroup <- data %>% 
  filter(!is.na(subgroup)) %>% 
  group_by(RM) %>% 
  mutate(
    total_kcal_RM = sum(kcal, na.rm = TRUE)
  ) %>% 
  group_by(RM, subgroup) %>% 
  summarise(
    upf_up_kcal_subgroup = sum(upf_up_kcal, na.rm = TRUE),
    total_kcal_RM = first(total_kcal_RM),
    .groups = "drop"
  ) %>% 
  mutate(
    upf_up_E_pct = 100 * upf_up_kcal_subgroup / total_kcal_RM
  )

```

Population-level E% UPF by foodgroup, stratified by RM and age-groups

```{r}
upf_by_foodgroup_and_agegroup <- data %>% 
  filter(!is.na(subgroup), !is.na(age_grp)) %>% 
  group_by(RM, age_grp) %>% 
  mutate(
    total_kcal_agegroup = sum(kcal, na.rm = TRUE)
  ) %>% 
  group_by(RM, age_grp, subgroup) %>% 
  summarise(
    upf_up_kcal_subgroup = sum(upf_up_kcal, na.rm = TRUE),
    total_kcal_agegroup = first(total_kcal_agegroup),
    .groups = "drop"
  ) %>% 
  mutate(
    upf_up_E_pct = 100 * upf_up_kcal_subgroup / total_kcal_agegroup
  )

```

Population-level E% UPF by subgroup, stratified by RM x sex

Denominator: total kcal within each RM x sex

```{r}
upf_by_foodgroup_and_sex <- data %>% 
  filter(
    !is.na(subgroup),
    !is.na(sex)
  ) %>% 
  group_by(RM, sex) %>% 
  mutate(
    total_kcal_grp = sum(kcal, na.rm = TRUE)
  ) %>% 
  group_by(RM, sex, subgroup) %>% 
  summarise(
    upf_up_kcal_subgroup = sum(upf_up_kcal, na.rm = TRUE),
    total_kcal_grp = first(total_kcal_grp),
    .groups = "drop"
  ) %>% 
  mutate(
    upf_up_E_pct = 100 * upf_up_kcal_subgroup / total_kcal_grp
  )
  
```

Population-level E% UPF by subgroup, stratified by RM x age_grp x sex

```{r}
upf_by_foodgroup_age_and_sex <- data %>% 
  filter(
    !is.na(subgroup),
    !is.na(age_grp),
    !is.na(sex)
  ) %>% 
  group_by(RM, age_grp, sex) %>% 
  mutate(
    total_kcal_grp = sum(kcal, na.rm = TRUE)
  ) %>% 
  group_by(RM, age_grp, sex, subgroup) %>% 
  summarise(
    upf_up_kcal_subgroup = sum(upf_up_kcal, na.rm = TRUE),
    total_kcal_grp = first(total_kcal_grp),
    .groups = "drop"
  ) %>% 
  mutate(
    upf_up_E_pct = 100 * upf_up_kcal_subgroup / total_kcal_grp
  )

```

Validation step

```{r}
upf_by_foodgroup_age_and_sex %>% 
  group_by(RM, age_grp, sex) %>% 
  summarise(total_E_pct = sum(upf_up_E_pct),
            .groups = "drop")

upf_by_foodgroup_age_and_sex
```

Export foodgroup-results to one excelfile

```{r}

wb_groups <- createWorkbook()

addWorksheet(wb_groups, "by RM only")
writeData(wb_groups, "by RM only", upf_by_foodgroup)

addWorksheet(wb_groups, "by agegroup")
writeData(wb_groups, "by agegroup", upf_by_foodgroup_and_agegroup)

addWorksheet(wb_groups, "by_sex")
writeData(wb_groups, "by_sex", upf_by_foodgroup_and_sex)

addWorksheet(wb_groups, "by age and sex")
writeData(wb_groups, "by age and sex", upf_by_foodgroup_age_and_sex)

saveWorkbook(wb_groups, 
             file = file.path(dir_output, "Foodgroups_results_up.xlsx"),
             overwrite = TRUE
          )

```

Calculate person-level total calories from GEF (=keyhole variable) and UPF

1.  Calculate per person totals needed

```{r}
id_col <- "person_ID"

totals <- data %>% 
  group_by(across(all_of(id_col))) %>% 
  summarise(
    total_kcal = sum(kcal, na.rm = TRUE),
    RM = first(RM),
    .groups = "drop"
  )
  
```

2.  GEF totals (only rows where keyhole == "yes")

```{r}
keyhole_totals <- data %>% 
  filter(keyhole == "yes") %>% 
  group_by(across(all_of(id_col))) %>% 
  summarise(
    keyhole_kcal = sum(kcal, na.rm = TRUE),
    keyhole_upf_kcal = sum(upf_mid_kcal, na.rm = TRUE),
    .groups = "drop"
  )

```

Step 3 Combine and compute percentages GEF and UPF within GEF per person

```{r}
person_results <- totals %>% 
  left_join(keyhole_totals, by = id_col) %>% 
  mutate(
    keyhole_kcal = coalesce(keyhole_kcal, 0),
    keyhole_upf_kcal = coalesce(keyhole_upf_kcal, 0),

# E% from foods classified as GEF
  keyhole_pct = if_else(total_kcal > 0, 100 * keyhole_kcal / total_kcal, NA_real_),

# E% within keyhole that's UPF
  upf_within_keyhole_pct = if_else(keyhole_kcal > 0, 100 * keyhole_upf_kcal / keyhole_kcal, NA_real_)
)
```

Persons with no GEF intake get keyhole_pct = 0 and upf_within_keyhole_pct = NA (denominator 0)

```{r}
GEF_summary <- person_results %>% 
  group_by(RM) %>% 
  summarise(
    n = n(),
    mean_keyhole_pct = mean(keyhole_pct, na.rm = TRUE),
    sd_keyhole_pct = sd(keyhole_pct, na.rm = TRUE),
    median_keyhole_pct = median(keyhole_pct, na.rm = TRUE),
    p25_keyhole_pct = quantile(keyhole_pct, 0.25, na.rm = TRUE),
    p75_keyhole_pct = quantile(keyhole_pct, 0.75, na.rm = TRUE),
    
    mean_upf_within_keyhole_pct = mean(upf_within_keyhole_pct, na.rm = TRUE),
    sd_upf_within_keyhole_pct = sd(upf_within_keyhole_pct, na.rm = TRUE),
    median_upf_within_keyhole_pct = median(upf_within_keyhole_pct, na.rm = TRUE),
    p25_upf_within_keyhole_pct = quantile(upf_within_keyhole_pct, 0.25, na.rm = TRUE),
    p75_upf_within_keyhole_pct = quantile(upf_within_keyhole_pct, 0.75, na.rm = TRUE),
    
    .groups = "drop"
  ) %>% 
  arrange(RM)

GEF_summary
```

Check number of NA per RM-group

```{r}
na_summary <- person_results %>% 
  group_by(RM) %>% 
  summarise(
    n_persons = n(),
    n_NA_upf_within_keyhole_pct = sum(is.na(upf_within_keyhole_pct)),
.groups = "drop"
  ) %>% 
  arrange(RM)

na_summary
```

```{r}
zero_summary <- person_results %>% 
  group_by(RM) %>% 
  summarise(
    n_persons = n(),
    n_0_keyhole_kcal = sum(keyhole_kcal == 0, na.rm = TRUE),
.groups = "drop"
  ) %>% 
  arrange(RM)

zero_summary
```

Export keyhole-results to one excelfile

```{r}

wb_gef <- createWorkbook()

addWorksheet(wb_gef, "GEF_summary")
writeData(wb_gef, "GEF_summary", GEF_summary)

saveWorkbook(wb_gef, 
             file = file.path(dir_output, "GEF_results.xlsx"), 
             overwrite = TRUE
            )

```

Calculate Energy density (excluding beverages) and collapse dataset to one row per person

```{r}
#1. Check for zeros in variable gram
sum(data$gram == 0, na.rm = TRUE)

```

```{r}
id_col <- "person_ID"

ed_person <- data %>% 
  filter(bev == "N", !is.na(gram), gram > 0, !is.na(kcal)) %>% 
  group_by(across(all_of(id_col))) %>% 
  summarise(
    ED_nobev = sum(kcal, na.rm = TRUE) / sum(gram, na.rm = TRUE),
    gram_nb_total = sum(gram, na.rm = TRUE),
    kcal_nb_total = sum(kcal, na.rm = TRUE),
    .groups = "drop"
  )

```

```{r}
sum(is.na(ed_person$ED_nobev))
```

Collapse dataset to get one row per person

```{r}

drop_vars <- c("food_ID", "food", "bev", "subgroup", "day_nr", "keyhole", "keyhole_ratio")
keep_first <- c("person_ID", "sex", "RM", "age_grp", "day_count")

df_collapsed <- data %>% 
  group_by(across(all_of(id_col))) %>% 
  summarise(
    across(any_of(keep_first), first),
    
    #mean daily intake for all remaining numeric variables
    across(
      .cols = where(is.numeric) & !any_of(c(drop_vars, keep_first)),
      .fns = ~sum(.x, na.rm = TRUE) / first(day_count)
    ),
    .groups = "drop"
  )

```

Join ED onto collapsed dataset

```{r}

df_collapsed <- df_collapsed %>% 
  left_join(ed_person %>% dplyr::select(all_of(id_col), ED_nobev), by = id_col)

```

Calculate energy percentage from UPF

```{r}
df_collapsed <- df_collapsed %>% 
  mutate(
    EP_upf_low = upf_low_kcal / kcal * 100,
    EP_upf_mid = upf_mid_kcal / kcal * 100,
    EP_upf_up = upf_up_kcal / kcal * 100
  )
```

Save collapsed dataset

```{r}
saveRDS(df_collapsed, file.path(dir_data, "Data_collapsed.rds"))
```
