---
title: "Foodgroups barplot"
format: html
---

```{r}
#| label: setup
#| include: false

project_root <- normalizePath(file.path(getwd(), ".."))

dir_data <- file.path(project_root, "data")
dir_output <- file.path(project_root, "output")

dir.create(dir_output, showWarnings = FALSE, recursive = TRUE)

```

analysis code starts here

```{r message=FALSE, warning=FALSE}

library(openxlsx)
library(writexl)
library(dplyr)
library(ggplot2)

```

Geet excel sheet names

```{r}
getSheetNames(
  file.path(dir_output, "Foodgroups_results_mid.xlsx")
)
```
Rename sheets

```{r}
#Sheet names in Excel
excel_sheets <- c(
  "by RM only",
  "by agegroup",
  "by_sex",
  "by age and sex"
)

#Short names to use in R

r_names <- c(
  "rm",
  "age",
  "sex",
  "age_sex"
)

```

Function to import one wb

```{r}
import_foodgroups <- function(filename) {
  
  wb <- lapply(
    excel_sheets, 
    function(sheet) {
      read.xlsx(
        file.path(dir_output, filename),
        sheet = sheet
      )
    }
  )
  
names(wb) <- r_names

wb
}
```

Apply function to import data

```{r}
foodgroups_low <- import_foodgroups("Foodgroups_results_low.xlsx")
foodgroups_mid <- import_foodgroups("Foodgroups_results_mid.xlsx")
foodgroups_up <- import_foodgroups("Foodgroups_results_up.xlsx")

```

```{r}
foodgroups_mid$rm
```
Bread
Sand_dish
Sugar
Meat
Fine_bakery
Condi
Ssb
Friedfood
Meat_dish
Milk

Inspect names

```{r}
names(foodgroups_mid$rm)
unique(foodgroups_mid$rm$subgroup)
unique(foodgroups_mid$rm$RM)
```
Check top 10 total

```{r}
subgroup_order_df <- foodgroups_mid$rm %>% 
  group_by(subgroup) %>% 
  summarise(
    rank_score = sum(upf_mid_E_pct),
    .groups = "drop"
  ) %>% 
  arrange(desc(rank_score))

```

```{r}
subgroup_order_df %>% 
  slice_head(n = 11)
```

Define subgroups names and plotting order

```{r}
subgroup_labels <- c(
  "Bread" = "Bread",
  "Sand_dish" = "Sandwiches, pizzas and similar",
  "Meat" = "Meat and meat products",
  "Sugar" = "Sugar and confectionery",
  "Fine_bakery" = "Fine bakery wares",
  "Condi" = "Savoury sauces and condiments",
  "Ssb" = "Sugar-sweetened beverages",
  "Friedfood" = "Savoury snacks",
  "Meat_dish" = "Meat dishes",
  "Milk" = "Milk, yoghurt and similar"
)
```

Define plotting order

```{r}
subgroup_order <- c(
  "Bread",
  "Sandwiches, pizzas and similar",
  "Meat and meat products",
  "Sugar and confectionery",
  "Fine bakery wares",
  "Savoury sauces and condiments",
  "Sugar-sweetened beverages",
  "Savoury snacks",
  "Meat dishes",
  "Milk, yoghurt and similar"
)
```

Define the RM labels and their order

```{r}
rm_labels <- c(
  "Vuxna" = "Adults",
  "Ungdom" = "Adolescents",
  "Barn" = "Young Children"
)
```

```{r}
rm_order <- c(
  "Adults",
  "Adolescents",
  "Young Children"
)

```


Create the main plotting table

```{r}
#Middle bound scenario

plot_mid <- foodgroups_mid$rm %>% 
  filter(subgroup %in% names(subgroup_labels)) %>% 
  transmute(
    subgroup = recode(subgroup, !!!subgroup_labels),
    RM = recode(RM, !!!rm_labels),
    estimate = upf_mid_E_pct
  ) %>% 
  mutate(
    subgroup = factor(subgroup, levels = subgroup_order),
    RM = factor(RM, levels = rm_order)
  ) %>% 
  arrange(subgroup, RM)

```

```{r}
plot_mid
```
Create corresponding lower- and upper-bound tables

```{r}
plot_low <- foodgroups_low$rm %>% 
  filter(subgroup %in% names(subgroup_labels)) %>% 
  transmute(
    subgroup = recode(subgroup, !!!subgroup_labels),
    RM = recode(RM, !!!rm_labels),
    lower = upf_low_E_pct
  ) %>% 
  mutate(
    subgroup = factor(subgroup, levels = subgroup_order),
    RM = factor(RM, levels = rm_order)
  ) %>% 
  arrange(subgroup, RM)


plot_up <- foodgroups_up$rm %>% 
  filter(subgroup %in% names(subgroup_labels)) %>% 
  transmute(
    subgroup = recode(subgroup, !!!subgroup_labels),
    RM = recode(RM, !!!rm_labels),
    upper = upf_up_E_pct
  ) %>% 
  mutate(
    subgroup = factor(subgroup, levels = subgroup_order),
    RM = factor(RM, levels = rm_order)
  ) %>% 
  arrange(subgroup, RM)

```

Join tables

```{r}
plot_data <- plot_mid %>% 
  left_join(
    plot_low,
    by = c("subgroup", "RM")
  ) %>% 
  left_join(
    plot_up,
    by = c("subgroup", "RM")
  ) %>% 
  arrange(subgroup, RM)

```

```{r}
plot_data
```

Define bar appearance

```{r}
bar_colors <- c(
  "Adults" = "royalblue3",
  "Adolescents" = "chartreuse3",
  "Young Children" = "indianred3"
)

bar_alpha <- 0.85

bar_border <- c(
  "Adults" = "royalblue4",
  "Adolescents" = "chartreuse4",
  "Young Children" = "indianred4"
)
```

```{r}
dodge_position <-  position_dodge2(
  width = 0.9,
  preserve = "single",
  padding = 0
)

error_position <- position_dodge(
  width = 0.75
)
```


Create plot

```{r}
barplot <- ggplot(
  plot_data,
  aes(
    x = subgroup,
    y = estimate,
    fill = RM,
    colour = RM,
    group = RM
  )
) +
  geom_col(
    position = dodge_position,
    width = 0.75,
    alpha = bar_alpha,
  ) +
  geom_errorbar(
    aes(
      ymin = lower,
      ymax = upper
    ),
    position = error_position,
      width = 0.15,
    ) +
  scale_fill_manual(
    values = bar_colors,
    name = NULL
  )+
  scale_color_manual(
    values = bar_border,
    name = NULL
  ) +
  scale_y_continuous(
    expand = expansion(mult = c(0, 0.05))
  ) +
  labs(
    x = NULL,
    y = "Contribution to total energy intake (E%)"
  ) +
  theme_classic() +
  theme(
    axis.text.x = element_text(
      angle = 45,
      hjust = 1
    ),
    legend.position = "bottom",
    legend.title = element_blank()
  )

barplot
  
```

Save images

```{r}

ggsave(
  filename = file.path(dir_output, "Foodgroups_figure.png"),
  plot = barplot,
  width = 8,
  height = 6,
  units = "in",
  dpi = 300,
  bg = "white"
)


```

And as pdf

```{r}
ggsave(
  filename = file.path(dir_output, "Foodgroups_figure.pdf"),
  plot = barplot,
  width = 8,
  height = 6,
  units = "in",
  dpi = 300,
  bg = "white"
)

```


Supplement table S7

```{r}
names(foodgroups_mid$rm)

names(foodgroups_mid$age)

names(foodgroups_mid$sex)

names(foodgroups_mid$age_sex)

```
Inspect the grouping values

```{r}
unique(foodgroups_mid$rm$RM)

unique(foodgroups_mid$age[, c("RM", "age_grp")])

unique(foodgroups_mid$sex[, c("RM", "sex")])

unique(foodgroups_mid$age_sex[, c("RM", "age_grp", "sex")])
```

Define order of sections

```{r}
section_order <- c(
  "Vuxna",
  "60-80",
  "40-59",
  "18-39",
  "Ungdom",
  "Ak2_g",
  "Ak8",
  "Ak5",
  "Barn",
  "4_ar",
  "18_manad"
)

sex_order <- c(
  "Overall",
  "M",
  "F"
)

sex_rm_order <- as.vector(
  outer(
    section_order,
    sex_order,
    paste,
    sep = ", "
  )
)

```


Build overall table

```{r}
top5_foodgroups <- function(df) {
  
  df %>% 
    arrange(desc(upf_mid_E_pct)) %>% 
    slice_head(n = 5) %>% 
    transmute(
      Foodgroup = subgroup, 
      Estimate = round(upf_mid_E_pct, 1)
    )
}
```


```{r}
make_top5_section <- function(df, group_label) {
  
  top5_foodgroups(df) %>% 
    mutate(Group = group_label, .before = 1)
}
```

Build the overall RM sections

```{r}
rm_sections <- lapply(
  unique(foodgroups_mid$rm$RM),
  function(rm_value) {
    
    make_top5_section(
      foodgroups_mid$rm %>% 
        filter(RM == rm_value),
      group_label = rm_value
    )
  }
)
```

Combine

```{r}
rm_sections <- bind_rows(rm_sections)
```

Create the age-group sections

```{r}
age_sections <- foodgroups_mid$age %>% 
  group_split(RM, age_grp) %>% 
  lapply(function(df) {
    
    make_top5_section(
      df,
      group_label = unique(df$age_grp)
    )
  }) %>% 
  bind_rows()

```

Combine and order

```{r}
overall_age_table <- bind_rows(
  rm_sections,
  age_sections
) %>% 
  mutate(
    Group = factor(Group, levels = section_order)
  ) %>% 
  arrange(Group, desc(Estimate))
```

```{r}
overall_age_table
```

Create sex sections

```{r}
sex_sections <- foodgroups_mid$sex %>% 
  group_split(RM, sex) %>% 
  lapply(function(df) { 
    
    group_name <- paste(
      unique(df$RM),
      unique(df$sex),
      sep = ", "
    )
    
    make_top5_section(
      df,
      group_label = group_name
    )
   }) %>% 
  bind_rows()
```

Order

```{r}
sex_sections <- sex_sections %>% 
  mutate(
    Group = factor(Group, levels = sex_rm_order)
  ) %>% 
  arrange(Group, desc(Estimate))

```

```{r}
sex_sections
```

Create sex x age sections

```{r}
age_sex_sections <- foodgroups_mid$age_sex %>% 
  group_split(age_grp, sex) %>% 
  lapply(function(df) {
    
    group_name <- paste(
      unique(df$age_grp),
      unique(df$sex),
      sep = ", "
    )
    
    make_top5_section(
      df,
      group_label = group_name
    )
  }) %>% 
  bind_rows
```

```{r}
age_sex_sections
```

Arrange

```{r}
age_sex_order <- as.vector(
  outer(
    section_order[-c(1,5,9)], 
    sex_order,
    paste,
    sep = ", "
  )
)
```

```{r}
age_sex_sections <- age_sex_sections %>% 
  mutate(
    Group = factor(Group, levels = age_sex_order)
  ) %>% 
  arrange(Group, desc(Estimate))
```

```{r}
age_sex_sections
```

Write to Excel

```{r}
wb <- createWorkbook()

addWorksheet(wb, "Overall and age")
writeData(wb, "Overall and age", overall_age_table)

addWorksheet(wb, "Sex")
writeData(wb, "Sex", sex_sections)

addWorksheet(wb, "Age and sex")
writeData(wb, "Age and sex", age_sex_sections)

saveWorkbook(
  wb,
  file.path(dir_output, "Foodgroups_top5.xlsx"),
  overwrite = TRUE
)
```




