---
title: "UPF intake by food group, UPF E% and GEF E%"
author: "Torsten Sällström"
format: html
editor: visual
---

In this script we: 1) summarize the foodgroups that contribute most kcal from UPF on population level. 2) calculate the dietary share of food items classified as GEF 3) collapses the dataset to one row per observatio and calculates E% UPF according to three classification scenarios.

Starting with the clean intake datasets created in step 02.1

```{r}
#| label: setup
#| include: false

project_root <- normalizePath(file.path(getwd(), ".."))

dir_data <- file.path(project_root, "data")
dir_output <- file.path(project_root, "output")

dir.create(dir_output, showWarnings = FALSE, recursive = TRUE)

```

analysis code starts here

```{r message=FALSE, warning=FALSE}

library(openxlsx)
library(writexl)
library(dplyr)
library (tidyr)
```

Import data (choose dataset)

```{r}
df_collapsed <- readRDS(file.path(dir_data, "Data_collapsed.rds"))

```

MIDDLE BOUND

1.  Summary stats

```{r}
summary_stats <- df_collapsed %>% 
  group_by(RM, sex) %>% 
  summarise(
    n = n(),
    median_upf = median(upf_mid_kcal, na.rm = TRUE),
    p25_upf = quantile(upf_mid_kcal, 0.25, na.rm = TRUE),
    p75_upf = quantile(upf_mid_kcal, 0.75, na.rm = TRUE),
    mean_EP = mean(EP_upf_mid, na.rm = TRUE),
    sd_EP = sd(EP_upf_mid, na.rm = TRUE),
    .groups = "drop"
  )
```

2.  Add subgroups totals (ignore sex)

```{r}
summary_totals <- df_collapsed %>% 
  group_by(RM) %>% 
  summarise(
    n = n(),
    median_upf = median(upf_mid_kcal, na.rm = TRUE),
    p25_upf = quantile(upf_mid_kcal, 0.25, na.rm = TRUE),
    p75_upf = quantile(upf_mid_kcal, 0.75, na.rm = TRUE),
    mean_EP = mean(EP_upf_mid, na.rm = TRUE),
    sd_EP = sd(EP_upf_mid, na.rm = TRUE),
    .groups = "drop"
  ) %>% 
  mutate(sex = "Total")

summary_all <- bind_rows(summary_totals, summary_stats)

summary_all

```

3.  Wilcoxon's test and t-test

```{r}
#Run per RM, comparing sexes
tests <- df_collapsed %>% 
  group_by(RM) %>% 
  group_map(~{
    dat <- .x
    
    w <- wilcox.test(
      upf_mid_kcal ~sex, 
      data = dat,
      paired = FALSE,
      conf.int = TRUE,
      conf.level = 0.95,
      exact = FALSE
      )
    #Welch t-test, mean diff + 95% CI
    tt <- t.test(
      EP_upf_mid ~sex, 
      data = dat,
      var.equal = FALSE
    )
    
    tibble(
      RM = .y$RM,
      
      wilcox_hl = w$estimate,
      wilcox_ci_low = w$conf.int[1],
      wilcox_ci_high = w$conf.int[2],
      wilcox_p = w$p.value,
      
      t_mean_diff = unname(diff(rev(tt$estimate))),
      t_ci_low = tt$conf.int[1],
      t_ci_high = tt$conf.int[2],
      t_p = tt$p.value
    )
  }) %>% 
  bind_rows()

tests
```

4.  Adjust p-values

```{r}
tests <- tests %>% 
  mutate(
    wilcox_p_adj = p.adjust(wilcox_p, method = "BH"),
    ttest_p_adj = p.adjust(t_p, method = "BH")
  )

tests
```

5.  Collect into one dataframe

```{r}
results_mid <- summary_all %>% 
  left_join(tests, by = "RM")

results_mid
```

6. Summary stats by age group

```{r}
age_grp_summary_mid <- df_collapsed %>% 
  group_by(RM, age_grp) %>% 
  summarise(
    n = n(),
    median_upf = median(upf_mid_kcal, na.rm = TRUE),
    p25_upf = quantile(upf_mid_kcal, 0.25, na.rm = TRUE),
    p75_upf = quantile(upf_mid_kcal, 0.75, na.rm = TRUE),
    mean_EP = mean(EP_upf_mid, na.rm = TRUE),
    sd_EP = sd(EP_upf_mid, na.rm = TRUE),
    .groups = "drop"
  )

age_grp_summary_mid
```




LOWER BOUND

1.  Summary stats

```{r}
summary_stats <- df_collapsed %>% 
  group_by(RM, sex) %>% 
  summarise(
    n = n(),
    median_upf = median(upf_low_kcal, na.rm = TRUE),
    p25_upf = quantile(upf_low_kcal, 0.25, na.rm = TRUE),
    p75_upf = quantile(upf_low_kcal, 0.75, na.rm = TRUE),
    mean_EP = mean(EP_upf_low, na.rm = TRUE),
    sd_EP = sd(EP_upf_low, na.rm = TRUE),
    .groups = "drop"
  )
```

2.  Add subgroups totals (ignore sex)

```{r}
summary_totals <- df_collapsed %>% 
  group_by(RM) %>% 
  summarise(
    n = n(),
    median_upf = median(upf_low_kcal, na.rm = TRUE),
    p25_upf = quantile(upf_low_kcal, 0.25, na.rm = TRUE),
    p75_upf = quantile(upf_low_kcal, 0.75, na.rm = TRUE),
    mean_EP = mean(EP_upf_low, na.rm = TRUE),
    sd_EP = sd(EP_upf_low, na.rm = TRUE),
    .groups = "drop"
  ) %>% 
  mutate(sex = "Total")

summary_all <- bind_rows(summary_totals, summary_stats)

```

3.  Wilcoxon's test and t-test

```{r}
##Run per RM, comparing sexes
tests <- df_collapsed %>% 
  group_by(RM) %>% 
  group_map(~{
    dat <- .x
    
    w <- wilcox.test(
      upf_low_kcal ~sex, 
      data = dat,
      paired = FALSE,
      conf.int = TRUE,
      conf.level = 0.95,
      exact = FALSE
      )
    #Welch t-test, mean diff + 95% CI
    tt <- t.test(
      EP_upf_low ~sex, 
      data = dat,
      var.equal = FALSE
    )
    
    tibble(
      RM = .y$RM,
      
      wilcox_hl = w$estimate,
      wilcox_ci_low = w$conf.int[1],
      wilcox_ci_high = w$conf.int[2],
      wilcox_p = w$p.value,
      
      t_mean_diff = unname(diff(rev(tt$estimate))),
      t_ci_low = tt$conf.int[1],
      t_ci_high = tt$conf.int[2],
      t_p = tt$p.value
    )
  }) %>% 
  bind_rows()

tests

```

4.  Adjust p-values

```{r}
tests <- tests %>% 
  mutate(
    wilcox_p_adj = p.adjust(wilcox_p, method = "BH"),
    ttest_p_adj = p.adjust(t_p, method = "BH")
  )

```

5.  Collect into one dataframe

```{r}
results_low <- summary_all %>% 
  left_join(tests, by = "RM")
results_low

```



6. Summary stats by age group

```{r}
age_grp_summary_low <- df_collapsed %>% 
  group_by(RM, age_grp) %>% 
  summarise(
    n = n(),
    median_upf = median(upf_low_kcal, na.rm = TRUE),
    p25_upf = quantile(upf_low_kcal, 0.25, na.rm = TRUE),
    p75_upf = quantile(upf_low_kcal, 0.75, na.rm = TRUE),
    mean_EP = mean(EP_upf_low, na.rm = TRUE),
    sd_EP = sd(EP_upf_low, na.rm = TRUE),
    .groups = "drop"
  )

age_grp_summary_low

```


UPPER BOUND

1.  Summary stats

```{r}
summary_stats <- df_collapsed %>% 
  group_by(RM, sex) %>% 
  summarise(
    n = n(),
    median_upf = median(upf_up_kcal, na.rm = TRUE),
    p25_upf = quantile(upf_up_kcal, 0.25, na.rm = TRUE),
    p75_upf = quantile(upf_up_kcal, 0.75, na.rm = TRUE),
    mean_EP = mean(EP_upf_up, na.rm = TRUE),
    sd_EP = sd(EP_upf_up, na.rm = TRUE),
    .groups = "drop"
  )
```

2.  Add subgroups totals (ignore sex)

```{r}
summary_totals <- df_collapsed %>% 
  group_by(RM) %>% 
  summarise(
    n = n(),
    median_upf = median(upf_up_kcal, na.rm = TRUE),
    p25_upf = quantile(upf_up_kcal, 0.25, na.rm = TRUE),
    p75_upf = quantile(upf_up_kcal, 0.75, na.rm = TRUE),
    mean_EP = mean(EP_upf_up, na.rm = TRUE),
    sd_EP = sd(EP_upf_up, na.rm = TRUE),
    .groups = "drop"
  ) %>% 
  mutate(sex = "Total")

summary_all <- bind_rows(summary_totals, summary_stats)

```

3.  Wilcoxon's test and t-test

```{r}
##Run per RM, comparing sexes
tests <- df_collapsed %>% 
  group_by(RM) %>% 
  group_map(~{
    dat <- .x
    
    w <- wilcox.test(
      upf_up_kcal ~sex, 
      data = dat,
      paired = FALSE,
      conf.int = TRUE,
      conf.level = 0.95,
      exact = FALSE
      )
    #Welch t-test, mean diff + 95% CI
    tt <- t.test(
      EP_upf_up ~sex, 
      data = dat,
      var.equal = FALSE
    )
    
    tibble(
      RM = .y$RM,
      
      wilcox_hl = w$estimate,
      wilcox_ci_low = w$conf.int[1],
      wilcox_ci_high = w$conf.int[2],
      wilcox_p = w$p.value,
      
      t_mean_diff = unname(diff(rev(tt$estimate))),
      t_ci_low = tt$conf.int[1],
      t_ci_high = tt$conf.int[2],
      t_p = tt$p.value
    )
  }) %>% 
  bind_rows()

tests
```

4.  Adjust p-values

```{r}
tests <- tests %>% 
  mutate(
    wilcox_p_adj = p.adjust(wilcox_p, method = "BH"),
    ttest_p_adj = p.adjust(t_p, method = "BH")
  )

```

5.  Collect into one dataframe

```{r}
results_up <- summary_all %>% 
  left_join(tests, by = "RM")
results_up

```

6. Summary stats by age group

```{r}
age_grp_summary_up <- df_collapsed %>% 
  group_by(RM, age_grp) %>% 
  summarise(
    n = n(),
    median_upf = median(upf_up_kcal, na.rm = TRUE),
    p25_upf = quantile(upf_up_kcal, 0.25, na.rm = TRUE),
    p75_upf = quantile(upf_up_kcal, 0.75, na.rm = TRUE),
    mean_EP = mean(EP_upf_up, na.rm = TRUE),
    sd_EP = sd(EP_upf_up, na.rm = TRUE),
    .groups = "drop"
  )

age_grp_summary_up

```



Export to one excelfile

```{r}

wb <- createWorkbook()

addWorksheet(wb, "results_mid")
writeData(wb, "results_mid", results_mid)

addWorksheet(wb, "age_groups_mid")
writeData(wb, "age_groups_mid", age_grp_summary_mid)

addWorksheet(wb, "results_low")
writeData(wb, "results_low", results_low)

addWorksheet(wb, "age_groups_low")
writeData(wb, "age_groups_low", age_grp_summary_low)

addWorksheet(wb, "results_up")
writeData(wb, "results_up", results_up)

addWorksheet(wb, "age_groups_up")
writeData(wb, "age_groups_up", age_grp_summary_up)

saveWorkbook(wb, 
             file = file.path(dir_output, "Table2_results.xlsx"), 
             overwrite = TRUE
          )

```
