---
title: "Cleaning nutrient variables and combining to masterfile"
author: "Torsten Sällström"
format: html
editor: visual
---

This script prepares the nutrient intake variables needed for subsequent analyses and combines intake data with background variables to one Masterfile. It starts with the collapsed and combined dataset created in step 3.0 as well as the clean background datasets for each survey.

Starting with the clean intake datasets created in step 02.1

```{r}
#| label: setup
#| include: false

project_root <- normalizePath(file.path(getwd(), ".."))

dir_data <- file.path(project_root, "data")
dir_output <- file.path(project_root, "output")

dir.create(dir_output, showWarnings = F, recursive = T)

```

analysis code starts here

```{r message=FALSE, warning=FALSE}

library(openxlsx)
library(writexl)
library(dplyr)
library (tidyr)
library(stringr)
library(purrr)

```

Import data

```{r}

data1 <- readRDS(file.path(dir_data, "Data_collapsed.rds"))

vux_bg <- readRDS(file.path(dir_data,"RM_vux_bg_clean.rds"))
ungdom_bg <- readRDS(file.path(dir_data,"RM_ungdom_bg_clean.rds"))
barn_bg <- readRDS(file.path(dir_data,"RM_barn_bg_clean.rds"))

```

Combine background datasets and join selected background variables to the main dataset

```{r}
#Define variables to join from each dataset
vars_vux <- c("person_ID", "age", "height_m", "height_cm", "bmi", "bmi_class", "weight", "overweight", "edu_combined", "leisure", "smoke")
vars_ungdom <- c("person_ID", "age", "height_m", "height_cm", "bmi", "bmi_z", "bmi_class", "weight", "overweight", "edu_combined", "leisure")
vars_barn <- c("person_ID", "age", "height_m", "height_cm", "bmi_z", "bmi_class", "weight", "overweight", "edu_combined", "screentime_combined")

#Define valid person_IDs (exist in background data and have intake data)
valid_ids <- data1 %>% 
  filter(!is.na(kcal)) %>% 
  pull(person_ID)

vux_sel <- vux_bg %>% dplyr::select(any_of(vars_vux)) %>% filter(person_ID %in% valid_ids)
ungdom_sel <- ungdom_bg %>% dplyr::select(any_of(vars_ungdom)) %>% filter(person_ID %in% valid_ids)
barn_sel <- barn_bg %>% dplyr::select(any_of(vars_barn)) %>% filter(person_ID %in% valid_ids)

bg_all <- bind_rows(vux_sel, ungdom_sel, barn_sel)

#Safety check: Stop if a person_ID appears in >1 source
dup <- bg_all %>% count(person_ID) %>% filter(n > 1)
if(nrow(dup) >0) stop("Stop")

#Collapse
bg_clean <- bg_all

#final join
data <- data1 %>% left_join(bg_clean, by = "person_ID")

```

```{r}
#data %>% group_by(RM) %>% 
  #median_x = median(EP_upf_mid)
```

Spotchecks after joining

```{r}
set.seed(1)
ids3 <- sample(unique(data$person_ID), 3)

check_vars <- c("age", "height_cm", "weight")

data %>% 
  filter(person_ID %in% ids3) %>% 
  dplyr::select(person_ID, any_of(check_vars)) %>% 
  arrange(person_ID)

```

```{r}
barn_bg %>% filter(person_ID %in% ids3) %>% dplyr::select(person_ID, any_of(check_vars)) %>% arrange(person_ID)
ungdom_bg %>% filter(person_ID %in% ids3) %>% dplyr::select(person_ID, any_of(check_vars)) %>% arrange(person_ID)

```

Compute EP protein, fat, SFA, added sugars and fibre density, wholegrain density (gram per 1000 kcal)

```{r}
data <- data %>% 
  mutate(
    EP_keyhole = keyhole_kcal / kcal * 100,
    EP_prot = protein * 4 / kcal * 100,
    EP_fat = fat * 9 / kcal * 100,
    EP_SFA = SFA * 9 / kcal * 100,
    EP_addedsugars = sugar_added * 4 / kcal * 100,
    fiber_dens = fiber / kcal * 1000,
    wholegrain_dens = wholegrain / kcal * 1000,
    naCl_dens = naCl / kcal * 1000,
  ) 
```

Compute nutrient density of vitamins and minerals (unit per 1000 kcal)

```{r}
data <- data %>%  
  mutate(
    vitA_dens = vitA / kcal * 1000,
    vitD_dens = vitD / kcal * 1000,
    vitE_dens = vitE / kcal * 1000,
    folate_dens = folate / kcal * 1000,
    vitB12_dens = vitB12 / kcal * 1000,
    vitC_dens = vitC / kcal * 1000,
    fe_dens = fe / kcal * 1000,
    ca_dens = ca / kcal * 1000,
    k_dens = k / kcal * 1000,
    mg_dens = mg / kcal * 1000,
    se_dens = se / kcal * 1000,
    zn_dens = zn / kcal * 1000,
    )

```

Create age bands + reference EI

```{r}
data <- data %>%
  mutate(
    age_band = case_when(
      age >= 1  & age < 4  ~ "1-3",
      age >= 4  & age < 7  ~ "4-6",
      age >= 7  & age < 11 ~ "7-10",
      age >= 11 & age < 15 ~ "11-14",
      age >= 15 & age < 18 ~ "15-17",
      age >= 18 & age < 25 ~ "18-24",
      age >= 25 & age < 51 ~ "25-50",
      age >= 51 & age < 71 ~ "51-70",
      age >= 71            ~ ">70",
      TRUE ~ NA_character_
    ),
    EI_ref_kcal = case_when(
      age_band == "1-3"  ~ 1099,
      age_band == "4-6"  ~ 1506,
      age_band == "7-10" ~ 1864,
      age_band == "11-14" & sex=="F" ~ 2199,
      age_band == "11-14" & sex=="M" ~ 2510,
      age_band == "15-17" & sex=="F" ~ 2414,
      age_band == "15-17" & sex=="M" ~ 3035,
      age_band == "18-24" & sex=="F" ~ 2247,
      age_band == "18-24" & sex=="M" ~ 2820,
      age_band == "25-50" & sex=="F" ~ 2151,
      age_band == "25-50" & sex=="M" ~ 2701,
      age_band == "51-70" & sex=="F" ~ 1984,
      age_band == "51-70" & sex=="M" ~ 2462,
      age_band == ">70"   & sex=="F" ~ 1960,
      age_band == ">70"   & sex=="M" ~ 2414,
      TRUE ~ NA_real_
    )
  )

```

Define AR-values for selected vitamins and minerals based on sex and age

```{r}
data <- data %>% 
  mutate(
    vitD_AR = 7.5, 
    
    vitC_AR = case_when(
      age_band == "1-3" ~20,
      age_band == "4-6" ~30,
      age_band == "7-10" ~45,
      age_band == "11-14" & sex == "F" ~60,
      age_band == "11-14" & sex == "M" ~65,
      age_band == "15-17" & sex == "F" ~75,
      age_band == "15-17" & sex == "M" ~85,
      age_band %in% c("18-24", "25-50", "51-70", ">70") & sex == "F" ~75,
      age_band %in% c("18-24", "25-50", "51-70", ">70") & sex == "M" ~90,
      TRUE ~NA_real_
    ),

folate_AR = case_when(
      age_band == "1-3" ~90,
      age_band == "4-6" ~110,
      age_band == "7-10" ~160,
      age_band == "11-14" & sex == "F" ~220,
      age_band == "11-14" & sex == "M" ~200,
      age_band == "15-17" & sex == "F" ~240,
      age_band == "15-17" & sex == "M" ~250,
      age_band %in% c("18-24", "25-50", "51-70", ">70") ~250,
      TRUE ~NA_real_
    ),

ca_AR = case_when(
     age_band == "1-3" ~400,
     age_band == "4-6" ~700,
     age_band == "7-10" ~675,
     age_band %in% c("11-14", "15-17")~980,
     age_band %in% c("18-24", "25-50")~870,
     age_band %in% c("51-70", ">70")  ~750,
     TRUE ~NA_real_
),

fe_AR = case_when(
   age_band == "1-3" ~6,
      age_band == "4-6" ~5,
      age_band == "7-10" ~7,
      age_band == "11-14" & sex == "F" ~10,
      age_band == "11-14" & sex == "M" ~9,
      age_band == "15-17" ~9,
      age_band %in% c("18-24", "25-50") & sex == "F" ~9,
      age_band %in% c("18-24", "25-50") & sex == "M" ~7,
      age_band %in% c("51-70", ">70") & sex == "F" ~6,
      age_band %in% c("51-70", ">70") & sex == "M" ~7,
      TRUE ~NA_real_
    ),
  vitE_AR = case_when(
      age_band == "1-3" ~6,
      age_band %in% c("4-6","7-10") ~7,
      age_band == "11-14" & sex == "F" ~8,
      age_band == "11-14" & sex == "M" ~9,
      age_band == "15-17" & sex == "F" ~9,
      age_band == "15-17" & sex == "M" ~10,
      age_band %in% c("18-24", "25-50", "51-70", ">70") & sex == "F" ~8,
      age_band %in% c("18-24", "25-50", "51-70", ">70") & sex == "M" ~9,
      TRUE ~NA_real_
    ),

  vitB12_AR = case_when(
      age_band == "1-3" ~1.2,
      age_band == "4-6" ~1.4,
      age_band == "7-10" ~2.0,
      age_band == "11-14" & sex == "F" ~2.8,
      age_band == "11-14" & sex == "M" ~2.6,
      age_band == "15-17" & sex == "F" ~3.1,
      age_band == "15-17" & sex == "M" ~3.2,
      age_band %in% c("18-24", "25-50", "51-70", ">70") ~3.2,
      TRUE ~NA_real_
    ),

   k_AR = case_when(
      age_band == "1-3" ~700,
      age_band == "4-6" ~900,
      age_band == "7-10" ~1450,
      age_band == "11-14" & sex == "F" ~1900,
      age_band == "11-14" & sex == "M" ~2050,
      age_band == "15-17" & sex == "F" ~2250,
      age_band == "15-17" & sex == "M" ~2700,
      age_band %in% c("18-24", "25-50", "51-70", ">70") & sex == "F" ~2800,
      age_band %in% c("18-24", "25-50", "51-70", ">70") & sex == "M" ~2800,
      TRUE ~NA_real_
    ),

    se_AR = case_when(
      age_band == "1-3" ~15,
      age_band == "4-6" ~20,
      age_band == "7-10" ~35,
      age_band == "11-14" ~50,
      age_band == "15-17" & sex == "F" ~55,
      age_band == "15-17" & sex == "M" ~70,
      age_band %in% c("18-24", "25-50", "51-70", ">70") & sex == "F" ~60,
      age_band %in% c("18-24", "25-50", "51-70", ">70") & sex == "M" ~70,
      TRUE ~NA_real_
    ),
)

    
  
```

```{r}
data %>% summarise(across(ends_with("_AR"), ~sum(is.na(.x))))

```

```{r}
problem_row <- data %>% 
  filter(is.na(vitC_AR)) %>% 
  select(person_ID, age, age_band, sex, vitC_AR, vitD_AR, folate_AR, ca_AR, fe_AR, se_AR)

problem_row
```

Calculate how large percent of AR each participant meets

```{r}
data <- data %>%
  mutate(
    pct_vitD = (vitD / vitD_AR) * 100,
    pct_vitE = (vitE/ vitE_AR) * 100,
    pct_folate = (folate / folate_AR) * 100,
    pct_vitB12 = (vitB12/ vitB12_AR) * 100,
    pct_vitC = (vitC/ vitC_AR) * 100,
    pct_fe = (fe / fe_AR) * 100,
    pct_ca = (ca / ca_AR) * 100,
    pct_k = (k / k_AR) * 100,
    pct_se = (se / se_AR) * 100
  )

```

NRF calculation on separate data. Starts with uncapped version.

```{r}
d1 <- data

```

Create RI/MRI denominators per day

```{r}
d2 <- d1 %>%
  mutate(
    # --- RI per DAY  ---
    vitA_drv = case_when(
      age < 4 ~300,
      age >= 4 & age < 7 ~350,
      age >= 7 & age < 11 ~450,
      sex == "F" & age >= 11 & age < 18 ~650,
      sex == "F" & age >= 18 ~700,
      sex == "M" & age >= 11 & age < 15 ~700,
      sex == "M" & age >= 15 & age < 18 ~750,
      sex == "M" & age >= 18 ~800,
      TRUE ~ NA_real_
    ),
    vitC_drv = case_when(
      age < 4 ~25,
      age >= 4 & age < 7 ~35,
      age >= 7 & age < 11 ~55,
      sex == "F" & age >= 11 & age < 15 ~75,
      sex == "M" & age >= 11 & age < 15 ~80,
      sex == "F" & age >= 15 & age < 18 ~90,
      sex == "M" & age >= 15 & age < 18 ~105,
      sex == "F" & age >= 18 ~95,
      sex == "M" & age >= 18 ~110,
      TRUE ~ NA_real_
    ),
    vitD_drv = case_when(
      age < 70 ~10,
      age >= 70 ~20,
      TRUE ~ NA_real_
    ),
    vitE_drv = case_when(
      age < 4 ~7,
      age >= 4 & age < 7 ~8,
      age >= 7 & age < 11 ~9,
      sex == "F" & age >= 11 & age < 15 ~10,
      sex == "F" & age >= 15 & age < 18 ~11,
      sex == "F" & age >= 18 ~10,
      sex == "M" & age >= 11 & age < 15 ~11,
      sex == "M" & age >= 15 & age < 18 ~12,
      sex == "M" & age >= 18 ~11,
      TRUE ~ NA_real_
    ),
    folate_drv = case_when(
      age < 4 ~120,
      age >= 4 & age < 7 ~140,
      age >= 7 & age < 11 ~200,
      sex == "F" & age >= 11 & age < 15 ~280,
      sex == "F" & age >= 15 & age < 18 ~340,
      sex == "F" & age >= 18 ~330,
      sex == "M" & age >= 11 & age < 15 ~260,
      sex == "M" & age >= 15 & age < 18 ~320,
      sex == "M" & age >= 18 ~330,
      TRUE ~ NA_real_
    ),
    fe_drv = case_when(
      age < 4 ~7,
      age >= 4 & age < 7 ~7,
      age >= 7 & age < 11 ~9,
      sex == "F" & age >= 11 & age < 15 ~13,
      sex == "M" & age >= 11 & age < 15 ~11,
      sex == "F" & age >= 15 ~15,
      sex == "M" & age >= 15 & age < 18 ~11,
      sex == "M" & age >= 18 ~9,
      TRUE ~ NA_real_ 
      ),
    ca_drv = case_when(
      age < 4 ~450,
      age >= 4 & age < 7 ~800,
      age >= 7 & age < 11 ~800,
      age >= 11 & age < 18 ~1150,
      age >= 18 & age < 25 ~1000,
      age >= 25 ~950,
      TRUE ~ NA_real_
    ),
     k_drv = case_when(
      age < 4 ~850,
      age >= 4 & age < 7 ~1150,
      age >= 7 & age < 11 ~1800,
      sex == "F" & age >= 11 & age < 15 ~2400,
      sex == "M" & age >= 11 & age < 15 ~2550,
      sex == "F" & age >= 15 & age < 18 ~2850,
      sex == "M" & age >= 15 & age < 18 ~3400,
      sex == "F" & age >= 18 ~3500,
      sex == "M" & age >= 18 ~3500,
      TRUE ~ NA_real_
    ),
      mg_drv = case_when(
      age < 4 ~170,
      age >= 4 & age < 7 ~230,
      age >= 7 & age < 11 ~230,
      sex == "F" & age >= 11 & age < 15 ~250,
      sex == "M" & age >= 11 & age < 15 ~300,
      sex == "F" & age >= 15 & age < 18 ~250,
      sex == "M" & age >= 15 & age < 18 ~300,
      sex == "F" & age >= 18 ~300,
      sex == "M" & age >= 18 ~350,
      TRUE ~ NA_real_
    ),
    
    naCl_drv = case_when(
      age < 4 ~2.75,
      age >= 4 & age < 7 ~3.5,
      age >= 7 & age < 11 ~4.25,
      age >= 11 & age < 18 ~5,
      age >= 18 ~5.75,
      TRUE ~ NA_real_
    )
  )

```

Convert denominators to "per 100 kcal" (\*\_drv100)

a)  Micronutrients + NaCl: RI / day -\> RI per 100 kcal using reference EI

```{r}
d3 <- d2 %>%
  mutate(
    vitA_drv100   = vitA_drv   / EI_ref_kcal * 100,
    vitC_drv100   = vitC_drv   / EI_ref_kcal * 100,
    vitD_drv100   = vitD_drv   / EI_ref_kcal * 100,
    vitE_drv100   = vitE_drv   / EI_ref_kcal * 100,
    folate_drv100 = folate_drv / EI_ref_kcal * 100,
    fe_drv100     = fe_drv     / EI_ref_kcal * 100,
    ca_drv100     = ca_drv     / EI_ref_kcal * 100,
    k_drv100      = k_drv      / EI_ref_kcal * 100,
    mg_drv100     = mg_drv     / EI_ref_kcal * 100,
    naCl_drv100   = naCl_drv   / EI_ref_kcal * 100
  )

```

b)  E% based nutrients: constants per 100 kcal (not using observed kcal)

```{r}
d3 <- d3 %>%
  mutate(
    prot_drv100        = 0.15 * 100 / 4,  # 3.75 g/100kcal
    SFA_drv100         = 0.10 * 100 / 9,  # 1.111... g/100kcal
    sugar_added_drv100 = 0.10 * 100 / 4   # 2.5 g/100kcal
  )

```

c)  Fibre: g/MJ -\> g/100 kcal

```{r}
MJ_per_100kcal <- 100 / 239.006

d3 <- d3 %>% 
  mutate(
    fiber_drv100 = case_when(
      age < 11 ~2.0 * MJ_per_100kcal,
      age >=11 & age < 18 ~2.5 * MJ_per_100kcal,
      age >= 18 ~3.0 * MJ_per_100kcal,
      TRUE ~NA_real_
    )
  )
```

Nutrient per 100 kcal

```{r}
d4 <- d3 %>%
  mutate(
    kcal_ok = !is.na(kcal) & kcal > 0,
    prot_100kcal = if_else(kcal_ok, (protein / kcal) * 100, NA_real_),
    fiber_100kcal = if_else(kcal_ok, (fiber / kcal) * 100, NA_real_),
    vitA_100kcal = if_else(kcal_ok, (vitA / kcal) * 100, NA_real_),
    vitC_100kcal = if_else(kcal_ok, (vitC / kcal) * 100, NA_real_),
    vitD_100kcal = if_else(kcal_ok, (vitD / kcal) * 100, NA_real_),
    vitE_100kcal = if_else(kcal_ok, (vitE / kcal) * 100, NA_real_),
    folate_100kcal = if_else(kcal_ok, (folate / kcal) * 100, NA_real_),
    fe_100kcal = if_else(kcal_ok, (fe / kcal) * 100, NA_real_),
    ca_100kcal = if_else(kcal_ok, (ca / kcal) * 100, NA_real_),
    k_100kcal = if_else(kcal_ok, (k / kcal) * 100, NA_real_),
    mg_100kcal = if_else(kcal_ok, (mg / kcal) * 100, NA_real_),
    SFA_100kcal = if_else(kcal_ok, (SFA / kcal) * 100, NA_real_),
    sugar_added_100kcal = if_else(kcal_ok, (sugar_added / kcal) * 100, NA_real_),
    naCl_100kcal = if_else(kcal_ok, (naCl / kcal) * 100, NA_real_)
  )

```

Compute quotients

```{r}
d5 <- d4 %>%
  mutate(
    prot_quote = prot_100kcal / prot_drv100, 
    fiber_quote = fiber_100kcal / fiber_drv100, 
    vitA_quote = vitA_100kcal / vitA_drv100, 
    vitC_quote = vitC_100kcal / vitC_drv100, 
    vitD_quote = vitD_100kcal / vitD_drv100, 
    vitE_quote = vitE_100kcal / vitE_drv100, 
    folate_quote = folate_100kcal / folate_drv100, 
    fe_quote = fe_100kcal / fe_drv100, 
    ca_quote = ca_100kcal / ca_drv100, 
    k_quote = k_100kcal / k_drv100, 
    mg_quote = mg_100kcal / mg_drv100, 
    SFA_quote = SFA_100kcal / SFA_drv100, 
    sugar_added_quote = sugar_added_100kcal / sugar_added_drv100, 
    naCl_quote = naCl_100kcal / naCl_drv100 
  )

```

NRF sums

```{r}
qual_vars <- c("prot_quote","fiber_quote","vitA_quote","vitC_quote","vitD_quote",
               "vitE_quote","folate_quote","fe_quote","ca_quote","k_quote","mg_quote")
disq_vars <- c("SFA_quote","sugar_added_quote","naCl_quote")

d6 <- d5 %>%
  mutate(
    nrf_positive = rowSums(across(all_of(qual_vars)), na.rm = TRUE),
    nrf_negative = rowSums(across(all_of(disq_vars)), na.rm = TRUE),
    nrf_uncapped = nrf_positive - nrf_negative
  )

```

Calculate nrf capped at 100% of RI for all qualifying nutrients except for fiber

```{r}
#Define variables to cap
qual_vars_to_cap <- setdiff(qual_vars, "fiber_quote")

d6 <-d6 %>% 
  mutate(
    across(all_of(qual_vars_to_cap), ~pmin(.x, 1), .names = "{.col}_cap"),
    
    #keep fibre as-is (uncapped) by just reusing the original column
    fiber_quote_cap = fiber_quote,
    
    #Compute capped NRF
    nrf_capped = rowSums(across(all_of(paste0(qual_vars_to_cap, "_cap"))), na.rm = TRUE) +
      fiber_quote_cap -
      rowSums(across(all_of(disq_vars)), na.rm = TRUE)
  )

```

```{r}
summary(d6$nrf_uncapped)
summary(d6$nrf_capped)
cor(d6$nrf_uncapped, d6$nrf_capped, use="complete.obs", method = "spearman")

```

```{r}
hist(
  d6$nrf_uncapped,
  breaks = 50
)
```

```{r}
hist(
  d6$nrf_capped,
  breaks = 50
)
```

```{r}
boxplot(
  d6$nrf_uncapped,
  d6$nrf_capped,
  horizontal = TRUE
)
```

```{r}
top10_nrf_capped <- d6 %>% 
  select(person_ID, kcal, nrf_capped) %>% 
  arrange(desc(nrf_capped)) %>% 
  slice_head(n=20)

top10_nrf_capped
```

```{r}
cor(d6$nrf_capped, d6$kcal, use = "complete.obs")
cor(d6$nrf_uncapped, d6$kcal, use = "complete.obs")
```

The capped version of NRF is chosen

```{r}
d6 <- d6 %>% 
  rename(nrf = nrf_capped)
```

Join back nrf to original data

```{r}
data <- data %>% 
  left_join(
    d6 %>% select(person_ID, nrf),
    by = "person_ID"
  )

```

Select variables to keep

```{r}
master <- data %>% 
  dplyr::select(
    person_ID:zn_dens,
    pct_vitD:pct_se,
    nrf
  )
```

and save as RDS

```{r}
saveRDS(master, file.path(dir_output, "Masterfile.rds"))
```
