---
title: "Cleaning nutrient variables and combining to masterfile"
author: "Torsten Sällström"
format: html
editor: visual
---

This script prepares the nutrient intake variables needed for subsequent analyses and combines intake data with background variables to one Masterfile. 
It starts with the collapsed and combined dataset created in step 3.0 as well as the clean background datasets for each survey.

Starting with the clean intake datasets created in step 02.1

```{r}
#| label: setup
#| include: false

project_root <- normalizePath(file.path(getwd(), ".."))

dir_data <- file.path(project_root, "data")
dir_output <- file.path(project_root, "output")

dir.create(dir_output, showWarnings = F, recursive = T)

```

analysis code starts here


```{r message=FALSE, warning=FALSE}

library(openxlsx)
library(writexl)
library(dplyr)
library (tidyr)
library(stringr)
library(purrr)

```

Import data

```{r}

data1 <- readRDS(file.path(dir_data, "Data_collapsed.rds"))

vux_bg <- readRDS(file.path(dir_data,"RM_vux_bg_clean.rds"))
ungdom_bg <- readRDS(file.path(dir_data,"RM_ungdom_bg_clean.rds"))
barn_bg <- readRDS(file.path(dir_data,"RM_barn_bg_clean.rds"))

```

Combine background datasets and join selected background variables to the main dataset

```{r}
#Define variables to join from each dataset
vars_vux <- c("person_ID", "age", "height_m", "height_cm", "bmi", "bmi_class", "weight", "overweight", "edu_combined", "leisure", "smoke")
vars_ungdom <- c("person_ID", "age", "height_m", "height_cm", "bmi", "bmi_z", "bmi_class", "weight", "overweight", "edu_combined", "leisure")
vars_barn <- c("person_ID", "age", "height_m", "height_cm", "bmi_z", "bmi_class", "weight", "overweight", "edu_combined", "screentime_combined")

#Define valid person_IDs (exist in background data and have intake data)
valid_ids <- data1 %>% 
  filter(!is.na(kcal)) %>% 
  pull(person_ID)

vux_sel <- vux_bg %>% dplyr::select(any_of(vars_vux)) %>% filter(person_ID %in% valid_ids)
ungdom_sel <- ungdom_bg %>% dplyr::select(any_of(vars_ungdom)) %>% filter(person_ID %in% valid_ids)
barn_sel <- barn_bg %>% dplyr::select(any_of(vars_barn)) %>% filter(person_ID %in% valid_ids)

bg_all <- bind_rows(vux_sel, ungdom_sel, barn_sel)

#Safety check: Stop if a person_ID appears in >1 source
dup <- bg_all %>% count(person_ID) %>% filter(n > 1)
if(nrow(dup) >0) stop("Stop")

#Collapse
bg_clean <- bg_all

#final join
data <- data1 %>% left_join(bg_clean, by = "person_ID")

```

```{r}
#data %>% group_by(RM) %>% 
  #median_x = median(EP_upf_mid)
```

Spotchecks after joining

```{r}
set.seed(1)
ids3 <- sample(unique(data$person_ID), 3)

check_vars <- c("age", "height_cm", "weight")

data %>% 
  filter(person_ID %in% ids3) %>% 
  dplyr::select(person_ID, any_of(check_vars)) %>% 
  arrange(person_ID)

```

```{r}
barn_bg %>% filter(person_ID %in% ids3) %>% dplyr::select(person_ID, any_of(check_vars)) %>% arrange(person_ID)
ungdom_bg %>% filter(person_ID %in% ids3) %>% dplyr::select(person_ID, any_of(check_vars)) %>% arrange(person_ID)

```

Compute EP protein, fat, SFA, added sugars and fibre density, wholegrain density (gram per 1000 kcal)

```{r}
data <- data %>% 
  mutate(
    EP_keyhole = keyhole_kcal / kcal * 100,
    EP_prot = protein * 4 / kcal * 100,
    EP_fat = fat * 9 / kcal * 100,
    EP_SFA = SFA * 9 / kcal * 100,
    EP_addedsugars = sugar_added * 4 / kcal * 100,
    fiber_dens = fiber / kcal * 1000,
    wholegrain_dens = wholegrain / kcal * 1000,
    naCl_dens = naCl / kcal * 1000,
  ) 
```

Compute nutrient density of vitamins and minerals (unit per 1000 kcal)

```{r}
data <- data %>%  
  mutate(
    vitA_dens = vitA / kcal * 1000,
    vitD_dens = vitD / kcal * 1000,
    vitE_dens = vitE / kcal * 1000,
    folate_dens = folate / kcal * 1000,
    vitB12_dens = vitB12 / kcal * 1000,
    vitC_dens = vitC / kcal * 1000,
    fe_dens = fe / kcal * 1000,
    ca_dens = ca / kcal * 1000,
    k_dens = k / kcal * 1000,
    mg_dens = mg / kcal * 1000,
    se_dens = se / kcal * 1000,
    zn_dens = zn / kcal * 1000,
    )

```

Define AR-values for selected vitamins and minerals based on sex and age

```{r}
data <- data %>%
  mutate(
    vitE_AR = case_when(
      age < 4 ~6,
      age >= 4 & age < 11 ~7,
      sex == "F" & age >= 11 & age < 15 ~8,
      sex == "F" & age >= 15 & age < 18 ~9,
      sex == "F" & age >= 18 ~8,
      sex == "M" & age >= 11 & age < 15 ~9,
      sex == "M" & age >= 15 & age < 18 ~10,
      sex == "M" & age >= 18 ~9
    ),
    vitD_AR = 7.5,
    
    folate_AR = case_when(
      age < 4 ~90,
      age >= 4 & age < 7 ~110,
      age >= 7 & age < 11 ~160,
      sex == "F" & age >= 11 & age < 15 ~220,
      sex == "F" & age >= 15 & age < 18 ~240,
      sex == "F" & age >= 18 ~250,
      sex == "M" & age >= 11 & age < 15 ~200,
      sex == "M" & age >= 15 ~250
    ),
    vitB12_AR = case_when(
      age < 4 ~1.2,
      age >= 4 & age < 7 ~1.4,
      age >= 7 & age < 11 ~2.0,
      sex == "F" & age >= 11 & age < 15 ~2.8,
      sex == "F" & age >= 15 & age < 18 ~3.1,
      sex == "M" & age >= 11 & age < 15 ~2.6,
      age >= 15 ~3.2
    ),
    vitC_AR = case_when(
      age < 4 ~20,
      age >= 4 & age < 7 ~30,
      age >= 7 & age < 11 ~45,
      sex == "F" & age >= 11 & age < 15 ~60,
      sex == "F" & age >= 15 ~75,
      sex == "M" & age >= 11 & age < 15 ~65,
      sex == "M" & age >= 15 & age < 18 ~85,
      sex == "M" & age >= 18 ~90
    ),
    fe_AR = case_when(
      age < 4 ~6,
      age >= 4 & age < 7 ~5,
      age >= 7 & age < 11 ~7,
      sex == "F" & age >= 11 & age < 15 ~10,
      sex == "F" & age >= 15 & age < 51 ~9,
      sex == "F" & age >= 51 ~6,
      sex == "M" & age >= 11 & age < 18 ~9,
      sex == "M" & age >= 18 ~7
    ),
    ca_AR = case_when(
      age < 4 ~400,
      age >= 4 & age < 7 ~700,
      age >= 7 & age < 18 ~980,
      age >= 18 & age < 25 ~870,
      age >= 25 ~750
    ),
     k_AR = case_when(
      age < 4 ~700,
      age >= 4 & age < 7 ~900,
      age >= 7 & age < 11 ~1450,
      sex == "F" & age >= 11 & age < 15 ~1900,
      sex == "F" & age >= 15 & age < 18 ~2250,
      sex == "F" & age >= 18 ~2800,
      sex == "M" & age >= 11 & age < 15 ~2050,
      sex == "M" & age >= 15 & age < 18 ~2700,
      sex == "M" & age >= 18 ~2700
    ),
    se_AR = case_when(
      age < 4 ~15,
      age >= 4 & age < 7 ~20,
      age >= 7 & age < 11 ~35,
      sex == "F" & age >= 11 & age < 15 ~50,
      sex == "F" & age >= 15 & age < 18 ~55,
      sex == "F" & age >= 18 ~60,
      sex == "M" & age >= 11 & age < 15 ~50,
      sex == "M" & age >= 15 ~70
    ),
  )

```

Calculate how large percent of AR each participant meets

```{r}
data <- data %>%
  mutate(
    pct_vitD = (vitD / vitD_AR) * 100,
    pct_vitE = (vitE/ vitE_AR) * 100,
    pct_folate = (folate / folate_AR) * 100,
    pct_vitB12 = (vitB12/ vitB12_AR) * 100,
    pct_vitC = (vitC/ vitC_AR) * 100,
    pct_fe = (fe / fe_AR) * 100,
    pct_ca = (ca / ca_AR) * 100,
    pct_k = (k / k_AR) * 100,
    pct_se = (se / se_AR) * 100
  )

```

Define reference values based on sex and age

```{r}
data <- data %>%
  mutate(
     prot_drv = 0.10 * kcal / 4,
    
     fiber_drv = case_when(
      age < 11 ~kcal / 1000 * 8.368,
      age >= 11 & age < 18 ~kcal / 1000 * 10.46,
      age >= 18 ~kcal / 1000 * 12.552,
      TRUE ~ NA_real_
    ),
    vitA_drv = case_when(
      age < 4 ~300,
      age >= 4 & age < 7 ~350,
      age >= 7 & age < 11 ~450,
      sex == "F" & age >= 11 & age < 18 ~650,
      sex == "F" & age >= 18 ~700,
      sex == "M" & age >= 11 & age < 15 ~700,
      sex == "M" & age >= 15 & age < 18 ~750,
      sex == "M" & age >= 18 ~800,
      TRUE ~ NA_real_
    ),
     vitC_drv = case_when(
      age < 4 ~25,
      age >= 4 & age < 7 ~35,
      age >= 7 & age < 11 ~55,
      sex == "F" & age >= 11 & age < 15 ~75,
      sex == "M" & age >= 11 & age < 15 ~80,
      sex == "F" & age >= 15 & age < 18 ~90,
      sex == "M" & age >= 15 & age < 18 ~105,
      sex == "F" & age >= 18 ~95,
      sex == "M" & age >= 18 ~110,
      TRUE ~ NA_real_
    ),
     vitD_drv = case_when(
      age < 70 ~10,
      age >= 70 ~20,
      TRUE ~ NA_real_
    ),
    vitE_drv = case_when(
      age < 4 ~7,
      age >= 4 & age < 7 ~8,
      age >= 7 & age < 11 ~9,
      sex == "F" & age >= 11 & age < 15 ~10,
      sex == "F" & age >= 15 & age < 18 ~11,
      sex == "F" & age >= 18 ~10,
      sex == "M" & age >= 11 & age < 15 ~11,
      sex == "M" & age >= 15 & age < 18 ~12,
      sex == "M" & age >= 18 ~11,
      TRUE ~ NA_real_
    ),
    folate_drv = case_when(
      age < 4 ~120,
      age >= 4 & age < 7 ~140,
      age >= 7 & age < 11 ~200,
      sex == "F" & age >= 11 & age < 15 ~280,
      sex == "F" & age >= 15 & age < 18 ~340,
      sex == "F" & age >= 18 ~330,
      sex == "M" & age >= 11 & age < 15 ~260,
      sex == "M" & age >= 15 & age < 18 ~320,
      sex == "M" & age >= 18 ~330,
      TRUE ~ NA_real_
    ),
    fe_drv = case_when(
      age < 4 ~7,
      age >= 4 & age < 7 ~7,
      age >= 7 & age < 11 ~9,
      sex == "F" & age >= 11 & age < 15 ~13,
      sex == "M" & age >= 11 & age < 15 ~11,
      sex == "F" & age >= 15 ~15,
      sex == "M" & age >= 15 & age < 18 ~11,
      sex == "M" & age >= 18 ~9,
      TRUE ~ NA_real_ 
      ),
    ca_drv = case_when(
      age < 4 ~450,
      age >= 4 & age < 7 ~800,
      age >= 7 & age < 11 ~800,
      age >= 11 & age < 18 ~1150,
      age >= 18 & age < 25 ~1000,
      age >= 25 ~950,
      TRUE ~ NA_real_
    ),
     k_drv = case_when(
      age < 4 ~850,
      age >= 4 & age < 7 ~1150,
      age >= 7 & age < 11 ~1800,
      sex == "F" & age >= 11 & age < 15 ~2400,
      sex == "M" & age >= 11 & age < 15 ~2550,
      sex == "F" & age >= 15 & age < 18 ~2850,
      sex == "M" & age >= 15 & age < 18 ~3400,
      sex == "F" & age >= 18 ~3500,
      sex == "M" & age >= 18 ~3500,
      TRUE ~ NA_real_
    ),
      mg_drv = case_when(
      age < 4 ~170,
      age >= 4 & age < 7 ~230,
      age >= 7 & age < 11 ~230,
      sex == "F" & age >= 11 & age < 15 ~250,
      sex == "M" & age >= 11 & age < 15 ~300,
      sex == "F" & age >= 15 & age < 18 ~250,
      sex == "M" & age >= 15 & age < 18 ~300,
      sex == "F" & age >= 18 ~300,
      sex == "M" & age >= 18 ~350,
      TRUE ~ NA_real_
    ),
    
     mg_drv = case_when(
      age < 4 ~170,
      age >= 4 & age < 7 ~230,
      age >= 7 & age < 11 ~230,
      sex == "F" & age >= 11 & age < 15 ~250,
      sex == "M" & age >= 11 & age < 15 ~300,
      sex == "F" & age >= 15 & age < 18 ~250,
      sex == "M" & age >= 15 & age < 18 ~300,
      sex == "F" & age >= 18 ~300,
      sex == "M" & age >= 18 ~350,
      TRUE ~ NA_real_
    ),
    SFA_drv = 0.1 * kcal / 9,
    
    sugar_added_drv = 0.1 * kcal / 4,
    
    naCl_drv = case_when(
      age < 4 ~2.75,
      age >= 4 & age < 7 ~3.5,
      age >= 7 & age < 11 ~4.25,
      age >= 11 & age < 18 ~5,
      age >= 18 ~5.75,
      TRUE ~ NA_real_
    )
)
    
```

Calculate nutrient content per 100 kcal

```{r}
data <- data %>%
  mutate(
    prot_100kcal = (protein / kcal) *100,
    fiber_100kcal = (fiber / kcal) * 100,
    vitA_100kcal = (vitA / kcal) * 100,
    vitC_100kcal = (vitC / kcal) * 100,
    vitD_100kcal = (vitD / kcal) * 100,
    vitE_100kcal = (vitE / kcal) * 100,
    folate_100kcal = (folate / kcal) * 100,
    fe_100kcal = (fe / kcal) * 100,
    ca_100kcal = (ca / kcal) * 100,
    k_100kcal = (k / kcal) * 100,
    mg_100kcal = (mg / kcal) * 100,
    SFA_100kcal = (SFA / kcal) * 100,
    sugar_added_100kcal = (sugar_added / kcal) * 100,
    naCl_100kcal = (naCl / kcal) * 100
  )

```

Calculate quotients

```{r}
data <- data %>% 
  mutate(
  prot_quote = (prot_100kcal / prot_drv * 100),
  fiber_quote = (fiber_100kcal / fiber_drv * 100),
  vitA_quote = (vitA_100kcal / vitA_drv * 100),
  vitC_quote = (vitC_100kcal / vitC_drv * 100),
  vitD_quote = (vitD_100kcal / vitD_drv * 100),
  vitE_quote = (vitE_100kcal / vitE_drv * 100),
  folate_quote = (folate_100kcal / folate_drv * 100),
  fe_quote = (fe_100kcal / fe_drv * 100),
  ca_quote = (ca_100kcal / ca_drv * 100),
  k_quote = (k_100kcal / k_drv * 100),
  mg_quote = (mg_100kcal / mg_drv * 100),
  SFA_quote = (SFA_100kcal / SFA_drv * 100),
  sugar_added_quote = (sugar_added_100kcal / sugar_added_drv * 100),
  naCl_quote = (naCl_100kcal / naCl_drv * 100)
)
  
```

Sum up quotients

```{r}
data <- data %>% 
  mutate(
  nrf_positive = rowSums(dplyr::select(., prot_quote:mg_quote), na.rm = T),
  nrf_negative = rowSums(dplyr::select(., SFA_quote:naCl_quote), na.rm = T),
  nrf = nrf_positive - nrf_negative
  )
```

Check

```{r}
summary(data$nrf)
```

Investigate extreme values

```{r}
top_10 <- data %>% 
  arrange(desc(nrf)) %>% 
  slice_head(n = 10)

bottom_10 <- data %>% 
  arrange(nrf) %>% 
  slice_head(n = 10)

top_10
bottom_10
```

Select variables to keep

```{r}
master <- data %>% 
  dplyr::select(
    person_ID:zn_dens,
    pct_vitD:pct_se,
    nrf
  )
```

and save as RDS

```{r}
saveRDS(master, file.path(dir_output, "Masterfile.rds"))
```
