---
title: "Energy- and nutrient density by quintiles of UPF-intake"
author: "Torsten Sällström"
format: html
editor: visual
---

This script divides the dataset into quintiles, computes density of energy, nutrients and vitamins and minerals within each quintile.

It starts with the Masterfile created in step 4.0

```{r}
#| label: setup
#| include: false

project_root <- normalizePath(file.path(getwd(), ".."))

dir_data <- file.path(project_root, "data")
dir_output <- file.path(project_root, "output")

dir.create(dir_output, showWarnings = FALSE, recursive = TRUE)

```

analysis code starts here

```{r message=FALSE, warning=FALSE}

library(dplyr)
library (tidyr)
library(openxlsx)

```

Import file

```{r message=FALSE, warning=FALSE}

nutrients <- readRDS(file.path(dir_data, "Masterfile.rds")) 

```

Define inputs

```{r}
df <- nutrients

rm_levels <- c("Vuxna", "Ungdom", "Barn")


```

MIDDLE BOUND SCENARIO

```{r}
vars_to_median_mid <- c("EP_upf_mid", "kcal", "ED_nobev", "EP_prot",  "fiber_dens", "wholegrain_dens", "EP_SFA", "EP_addedsugars", "naCl_dens", "EP_keyhole", "nrf", "vitD_dens", "vitE_dens", "folate_dens", "vitB12_dens", "vitC_dens", "fe_dens", "ca_dens", "k_dens", "se_dens")

ep <- "EP_upf_mid"
outfile <- file.path(dir_output, paste0("Quintiles_summary_", ep, ".xlsx"))
```

Block 1. Basic input checks

```{r}
stopifnot(is.data.frame(df))
stopifnot(all(c("RM", ep, vars_to_median_mid) %in% names(df)))

```

Block 2 - working copy & coerce median vars to numeric

```{r}
df <- df %>% 
  mutate(across(all_of(vars_to_median_mid), ~suppressWarnings(as.numeric(.))))

```

Block 3 - Create Excel workbook

```{r}
wb <- createWorkbook()
```

RM = Vuxna

```{r}


d <- df %>% filter(RM == "Vuxna")
n_total <- nrow(d)

d <- d %>% filter(!is.na(.data[[ep]]))
n_assigned <- nrow(d)

q <- d %>% 
  mutate(Q = ntile(.data[[ep]], 5)) %>% 
  group_by(Q) %>% 
  summarise(
    across(all_of(vars_to_median_mid), ~median(.x, na.rm = TRUE)),
    .groups = "drop"
  ) %>% 
  pivot_longer(
    cols = all_of(vars_to_median_mid),
    names_to = "variable",
    values_to = "median"
  ) %>% 
  pivot_wider(
    names_from = Q,
    values_from = median,
    names_prefix = "Q"
  )


addWorksheet(wb, "Vuxna")
writeData(wb, "Vuxna", 
          tibble(metric = c("n_total", "n_assigned"),
                 value = c(n_total, n_assigned)),
          colNames = FALSE)
writeData(wb, "Vuxna", q, startRow = 5)

```

RM = Ungdom

```{r}
d <- df %>% filter(RM == "Ungdom")
n_total <- nrow(d)

d <- d %>% filter(!is.na(.data[[ep]]))
n_assigned <- nrow(d)

q <- d %>% 
  mutate(Q = ntile(.data[[ep]], 5)) %>% 
  group_by(Q) %>% 
  summarise(
    across(all_of(vars_to_median_mid), ~median(.x, na.rm = TRUE)),
    .groups = "drop"
  ) %>% 
  pivot_longer(
    cols = all_of(vars_to_median_mid),
    names_to = "variable",
    values_to = "median"
  ) %>% 
  pivot_wider(
    names_from = Q,
    values_from = median,
    names_prefix = "Q"
  )


addWorksheet(wb, "Ungdom")
writeData(wb, "Ungdom", 
          tibble(metric = c("n_total", "n_assigned"),
                 value = c(n_total, n_assigned)),
          colNames = FALSE)
writeData(wb, "Ungdom", q, startRow = 5)

```

RM = Barn

```{r}
d <- df %>% filter(RM == "Barn")
n_total <- nrow(d)

d <- d %>% filter(!is.na(.data[[ep]]))
n_assigned <- nrow(d)

q <- d %>% 
  mutate(Q = ntile(.data[[ep]], 5)) %>% 
  group_by(Q) %>% 
  summarise(
    across(all_of(vars_to_median_mid), ~median(.x, na.rm = TRUE)),
    .groups = "drop"
  ) %>% 
  pivot_longer(
    cols = all_of(vars_to_median_mid),
    names_to = "variable",
    values_to = "median"
  ) %>% 
  pivot_wider(
    names_from = Q,
    values_from = median,
    names_prefix = "Q"
  )


addWorksheet(wb, "Barn")
writeData(wb, "Barn", 
          tibble(metric = c("n_total", "n_assigned"),
                 value = c(n_total, n_assigned)),
          colNames = FALSE)
writeData(wb, "Barn", q, startRow = 5)

```

n per quintile

```{r}
n_per_quintile <- df %>% 
  filter(!is.na(.data[[ep]])) %>% 
  group_by(RM) %>% 
  mutate(Q = ntile(.data[[ep]], 5)) %>% 
  count(RM, Q) %>% 
  arrange(RM, Q)

n_per_quintile

addWorksheet(wb, "Counts_per_quintile")
writeData(wb, "Counts_per_quintile", n_per_quintile)
```

Ranges per quintile

```{r}
range_per_quintile <- df %>% 
  filter(!is.na(.data[[ep]])) %>% 
  group_by(RM) %>%
  mutate(Q = ntile(.data[[ep]], 5)) %>% 
  group_by(RM, Q) %>%
  summarise(
    min_EP = min(.data[[ep]], na.rm = TRUE), 
    max_EP = max(.data[[ep]], na.rm = TRUE),
    .groups = "drop"
  ) %>% 
  arrange(RM, Q)

range_per_quintile

addWorksheet(wb, "Range_per_quintile")
writeData(wb, "Range_per_quintile", range_per_quintile)
```

Save Excel file

```{r}
saveWorkbook(wb, outfile, overwrite = TRUE)

```

LOWER BOUND SCENARIO

Define inputs

```{r}

vars_to_median_low <- c("EP_upf_low", "kcal", "ED_nobev", "EP_prot",  "fiber_dens", "wholegrain_dens", "EP_SFA", "EP_addedsugars", "naCl_dens", "EP_keyhole", "nrf", "vitD_dens", "vitE_dens", "folate_dens", "vitB12_dens", "vitC_dens", "fe_dens", "ca_dens", "k_dens", "se_dens")


ep <- "EP_upf_low"
outfile <- file.path(dir_output, paste0("Quintiles_summary_", ep, ".xlsx"))
```

Create new Excel workbook

```{r}
wb <- createWorkbook()
```

RM = Vuxna

```{r}


d <- df %>% filter(RM == "Vuxna")
n_total <- nrow(d)

d <- d %>% filter(!is.na(.data[[ep]]))
n_assigned <- nrow(d)

q <- d %>% 
  mutate(Q = ntile(.data[[ep]], 5)) %>% 
  group_by(Q) %>% 
  summarise(
    across(all_of(vars_to_median_low), ~median(.x, na.rm = TRUE)),
    .groups = "drop"
  ) %>% 
  pivot_longer(
    cols = all_of(vars_to_median_low),
    names_to = "variable",
    values_to = "median"
  ) %>% 
  pivot_wider(
    names_from = Q,
    values_from = median,
    names_prefix = "Q"
  )


addWorksheet(wb, "Vuxna")
writeData(wb, "Vuxna", 
          tibble(metric = c("n_total", "n_assigned"),
                 value = c(n_total, n_assigned)),
          colNames = FALSE)
writeData(wb, "Vuxna", q, startRow = 5)

```

RM = Ungdom

```{r}
d <- df %>% filter(RM == "Ungdom")
n_total <- nrow(d)

d <- d %>% filter(!is.na(.data[[ep]]))
n_assigned <- nrow(d)

q <- d %>% 
  mutate(Q = ntile(.data[[ep]], 5)) %>% 
  group_by(Q) %>% 
  summarise(
    across(all_of(vars_to_median_low), ~median(.x, na.rm = TRUE)),
    .groups = "drop"
  ) %>% 
  pivot_longer(
    cols = all_of(vars_to_median_low),
    names_to = "variable",
    values_to = "median"
  ) %>% 
  pivot_wider(
    names_from = Q,
    values_from = median,
    names_prefix = "Q"
  )


addWorksheet(wb, "Ungdom")
writeData(wb, "Ungdom", 
          tibble(metric = c("n_total", "n_assigned"),
                 value = c(n_total, n_assigned)),
          colNames = FALSE)
writeData(wb, "Ungdom", q, startRow = 5)

```

RM = Barn

```{r}
d <- df %>% filter(RM == "Barn")
n_total <- nrow(d)

d <- d %>% filter(!is.na(.data[[ep]]))
n_assigned <- nrow(d)

q <- d %>% 
  mutate(Q = ntile(.data[[ep]], 5)) %>% 
  group_by(Q) %>% 
  summarise(
    across(all_of(vars_to_median_low), ~median(.x, na.rm = TRUE)),
    .groups = "drop"
  ) %>% 
  pivot_longer(
    cols = all_of(vars_to_median_low),
    names_to = "variable",
    values_to = "median"
  ) %>% 
  pivot_wider(
    names_from = Q,
    values_from = median,
    names_prefix = "Q"
  )


addWorksheet(wb, "Barn")
writeData(wb, "Barn", 
          tibble(metric = c("n_total", "n_assigned"),
                 value = c(n_total, n_assigned)),
          colNames = FALSE)
writeData(wb, "Barn", q, startRow = 5)

```

n per quintile

```{r}
n_per_quintile <- df %>% 
  filter(!is.na(.data[[ep]])) %>% 
  group_by(RM) %>% 
  mutate(Q = ntile(.data[[ep]], 5)) %>% 
  count(RM, Q) %>% 
  arrange(RM, Q)

n_per_quintile

addWorksheet(wb, "Counts_per_quintile")
writeData(wb, "Counts_per_quintile", n_per_quintile)
```

Ranges per quintile

```{r}
range_per_quintile <- df %>% 
  filter(!is.na(.data[[ep]])) %>% 
  group_by(RM) %>%
  mutate(Q = ntile(.data[[ep]], 5)) %>% 
  group_by(RM, Q) %>%
  summarise(
    min_EP = min(.data[[ep]], na.rm = TRUE), 
    max_EP = max(.data[[ep]], na.rm = TRUE),
    .groups = "drop"
  ) %>% 
  arrange(RM, Q)

range_per_quintile

addWorksheet(wb, "Range_per_quintile")
writeData(wb, "Range_per_quintile", range_per_quintile)
```

Save Excel file

```{r}
saveWorkbook(wb, outfile, overwrite = TRUE)

```

UPPER BOUND SCENARIO

Define inputs

```{r}
vars_to_median_up <- c("EP_upf_up", "kcal", "ED_nobev", "EP_prot",  "fiber_dens", "wholegrain_dens", "EP_SFA", "EP_addedsugars", "naCl_dens", "EP_keyhole", "nrf", "vitD_dens", "vitE_dens", "folate_dens", "vitB12_dens", "vitC_dens", "fe_dens", "ca_dens", "k_dens", "se_dens")


ep <- "EP_upf_up"
outfile <- file.path(dir_output, paste0("Quintiles_summary_", ep, ".xlsx"))
```

Create new Excel workbook

```{r}
wb <- createWorkbook()
```

RM = Vuxna

```{r}


d <- df %>% filter(RM == "Vuxna")
n_total <- nrow(d)

d <- d %>% filter(!is.na(.data[[ep]]))
n_assigned <- nrow(d)

q <- d %>% 
  mutate(Q = ntile(.data[[ep]], 5)) %>% 
  group_by(Q) %>% 
  summarise(
    across(all_of(vars_to_median_up), ~median(.x, na.rm = TRUE)),
    .groups = "drop"
  ) %>% 
  pivot_longer(
    cols = all_of(vars_to_median_up),
    names_to = "variable",
    values_to = "median"
  ) %>% 
  pivot_wider(
    names_from = Q,
    values_from = median,
    names_prefix = "Q"
  )


addWorksheet(wb, "Vuxna")
writeData(wb, "Vuxna", 
          tibble(metric = c("n_total", "n_assigned"),
                 value = c(n_total, n_assigned)),
          colNames = FALSE)
writeData(wb, "Vuxna", q, startRow = 5)

```

RM = Ungdom

```{r}
d <- df %>% filter(RM == "Ungdom")
n_total <- nrow(d)

d <- d %>% filter(!is.na(.data[[ep]]))
n_assigned <- nrow(d)

q <- d %>% 
  mutate(Q = ntile(.data[[ep]], 5)) %>% 
  group_by(Q) %>% 
  summarise(
    across(all_of(vars_to_median_up), ~median(.x, na.rm = TRUE)),
    .groups = "drop"
  ) %>% 
  pivot_longer(
    cols = all_of(vars_to_median_up),
    names_to = "variable",
    values_to = "median"
  ) %>% 
  pivot_wider(
    names_from = Q,
    values_from = median,
    names_prefix = "Q"
  )


addWorksheet(wb, "Ungdom")
writeData(wb, "Ungdom", 
          tibble(metric = c("n_total", "n_assigned"),
                 value = c(n_total, n_assigned)),
          colNames = FALSE)
writeData(wb, "Ungdom", q, startRow = 5)

```

RM = Barn

```{r}
d <- df %>% filter(RM == "Barn")
n_total <- nrow(d)

d <- d %>% filter(!is.na(.data[[ep]]))
n_assigned <- nrow(d)

q <- d %>% 
  mutate(Q = ntile(.data[[ep]], 5)) %>% 
  group_by(Q) %>% 
  summarise(
    across(all_of(vars_to_median_up), ~median(.x, na.rm = TRUE)),
    .groups = "drop"
  ) %>% 
  pivot_longer(
    cols = all_of(vars_to_median_up),
    names_to = "variable",
    values_to = "median"
  ) %>% 
  pivot_wider(
    names_from = Q,
    values_from = median,
    names_prefix = "Q"
  )


addWorksheet(wb, "Barn")
writeData(wb, "Barn", 
          tibble(metric = c("n_total", "n_assigned"),
                 value = c(n_total, n_assigned)),
          colNames = FALSE)
writeData(wb, "Barn", q, startRow = 5)

```

n per quintile

```{r}
n_per_quintile <- df %>% 
  filter(!is.na(.data[[ep]])) %>% 
  group_by(RM) %>% 
  mutate(Q = ntile(.data[[ep]], 5)) %>% 
  count(RM, Q) %>% 
  arrange(RM, Q)

n_per_quintile

addWorksheet(wb, "Counts_per_quintile")
writeData(wb, "Counts_per_quintile", n_per_quintile)
```

Ranges per quintile

```{r}
range_per_quintile <- df %>% 
  filter(!is.na(.data[[ep]])) %>% 
  group_by(RM) %>%
  mutate(Q = ntile(.data[[ep]], 5)) %>% 
  group_by(RM, Q) %>%
  summarise(
    min_EP = min(.data[[ep]], na.rm = TRUE), 
    max_EP = max(.data[[ep]], na.rm = TRUE),
    .groups = "drop"
  ) %>% 
  arrange(RM, Q)

range_per_quintile

addWorksheet(wb, "Range_per_quintile")
writeData(wb, "Range_per_quintile", range_per_quintile)
```

Save Excel file

```{r}
saveWorkbook(wb, outfile, overwrite = TRUE)

```
