---
title: "Quantile regression"
format: html
editor: visual
---

This script runs quantile regression at the median on nutrient intake variables with E% from UPF (scaled by division of 10) as predictor. Adult dataset.

```{r}
#| label: setup
#| include: false

project_root <- normalizePath(file.path(getwd(), ".."))

dir_data <- file.path(project_root, "data")
dir_output <- file.path(project_root, "output")

dir.create(dir_output, showWarnings = F, recursive = T)

```

Analytic code starts here

```{r message=FALSE, warning=FALSE}
library(openxlsx)
library(dplyr)
library (tidyr)
library(stringr)
library(purrr)
library(quantreg)
```

Import data

```{r}
master <- readRDS(file.path(dir_data,"Masterfile.rds")) 

```

Filter for adults. Select variables of interest

```{r}
vux <- master %>% 
  filter(RM == "Vuxna") %>% 
  dplyr::select(person_ID, sex, age, bmi, edu_combined, leisure, smoke, kcal, ED_nobev, EP_prot, fiber_dens, wholegrain_dens, EP_SFA, EP_addedsugars, naCl_dens, EP_keyhole, nrf, vitD_dens, vitE_dens, folate_dens, vitB12_dens, vitC_dens, fe_dens, ca_dens, k_dens, se_dens, EP_upf_low, EP_upf_up, EP_upf_mid)
```

Convert selected variables to factors

```{r}
vux <- vux %>% 
  mutate(
    sex = as.factor(sex),
    edu_combined = as.factor(edu_combined),
    leisure = as.factor(leisure),
    smoke = as.factor(smoke)
  )

```

Create quintiles for EP UPF mid, low and up

```{r}
vux <- vux %>% 
  mutate(
    quintile_mid = as.numeric(ntile(EP_upf_mid, 5)),
    quintile_low = as.numeric(ntile(EP_upf_low, 5)),
    quintile_up = as.numeric(ntile(EP_upf_up, 5))
  )

```

Define dependent variables

```{r}
dep_vars <- names(vux)[which(names(vux) == "kcal"):which(names(vux) == "se_dens")]
```

Scale predictor variable

```{r}
vux$EP_upf_low_10 <- vux$EP_upf_low / 10
vux$EP_upf_mid_10 <- vux$EP_upf_mid / 10
```

Define covariates

```{r}
covars <- c("age", "sex", "edu_combined", "leisure", "smoke")
```

Check for zero reporters for intake variables

```{r}
zero_counts <- sapply(vux[dep_vars], function(x) sum(x == 0))

zero_counts_df <- data.frame(
  variable = names(zero_counts),
  n_zeros = as.integer(zero_counts),
  row.names = NULL
)

print(zero_counts_df)
```

Count complete cases

```{r}

n_total <- nrow(vux)

n_complete <- sum(complete.cases(vux[, covars]))


n_obs_df <- data.frame(
  n_total = n_total,
  n_complete = n_complete,
  perc_complete = round(100 * n_complete / n_total, 1)
)
```

QUANTILE REGRESSION

1.  MIDDLE BOUND

Define function to run qr on all outcomes

```{r}
qreg_orig <- function(data, outcome, predictor, covars, 
                      tau = 0.5,
                      R = 2000, 
                      seed = 123) {
  
  form <- as.formula(
    paste0(outcome, "~", predictor, "+", paste(covars, collapse = "+"))
  )
  
  fit <- rq(form, tau = tau, data = data)
  set.seed(seed)
  sm <- summary(fit, se = "boot", R = R)
  
  #Extract row for predictor
  row <- sm$coefficients[predictor, ]
  beta <- row["Value"]
  se <- row["Std. Error"]
  pval <- row["Pr(>|t|)"]
  
  ci <- c(beta -1.96 * se, beta + 1.96 * se)
  
  #Return as a one-row data frame
  data.frame(
    outcome = outcome,
    beta = beta,
    ci_lower = ci[1],
    ci_upper = ci[2],
    p_value = pval,
    stringsAsFactors = F
  )
  
}
```

Apply function to all outcomes

```{r}
results <- do.call(
  rbind,
  lapply(dep_vars, qreg_orig,
         data = vux,
         predictor = "EP_upf_mid_10",
         covars = covars,
         tau = 0.5,
         R = 2000,
         seed = 123)
)
```

Check stability

```{r}
qreg_orig(vux, outcome="kcal", predictor = "EP_upf_mid_10", covars = covars, seed=1)$beta

qreg_orig(vux, outcome="kcal", predictor = "EP_upf_mid_10", covars = covars, seed=2)$beta
```

```{r}
print(results)
```

Adjust for multiple testing

```{r}
results$padj <- p.adjust(results$p_value, method = "BH")
```

Function to customize rounding

```{r}
round_by_var <- function(outcome, value, stat = c("beta", "q5_vs_q1", "ci")) {
  stat <-  match.arg(stat)
  
  if(stat == "ci") {
    return(round(value, 2))
  }
  
  if(outcome %in% c("kcal", "folate_dens", "ca_dens", "k_dens")) {
    return(round(value, 0))
  } else if (outcome == "ED_nobev"){
    return(round(value, 2))
  } else {
    return(round(value, 1))
  }
}

```

Apply function

```{r}
results_clean <- results %>% 
  rowwise() %>% 
  mutate(
    beta = round_by_var(outcome, beta, "beta"),
    ci_lower = round_by_var(outcome, ci_lower, "ci"),
    ci_upper = round_by_var(aoutcome, ci_upper, "ci")
  ) %>% 
  ungroup()
```

Format

```{r}
results_clean$CI <- paste0("(", results_clean$ci_lower, ", ", results_clean$ci_upper, ")")

results_clean <- results_clean[, c("outcome", "beta", "CI", "padj")]
colnames(results_clean)[4] <- "p_adj"

results_clean
```

Test of linearity: Spline vs linear QR

Set knots

```{r}
library(splines)

#Chose outcome variable by indexing
yname <- dep_vars[5]
xname <- "EP_upf_mid_10"

vars_needed <- c(yname, xname, covars)
dat <- vux[, vars_needed]
dat <- dat[complete.cases(dat), ]

```

Step 1 define y and x

```{r}
y <- dat[[yname]]
x <- dat[[xname]]
```

Step 2 build covariate matrix

```{r}
Z <- model.matrix(~., data = dat[, covars, drop = FALSE])

Z <- Z[, colnames(Z) != "(Intercept)", drop = FALSE]

```

Step 3 build linear vs spline design matrices

```{r}
X_lin <- cbind(Intercept = 1, x = x, Z)

kn <- as.numeric(quantile(x, probs = c(0.10, 0.50, 0.90)))
B <- ns(x, knots = kn)

X_spl <- cbind(Intercept = 1, B, Z)

```

Step 4 - fit median quantile regression (matrix interface)

```{r}
fit_lin <- rq.fit(X_lin, y, tau = 0.5)
fit_spl <- rq.fit(X_spl, y, tau = 0.5)
```
Step 5 - compare fitted conditional medians over x (visual check)

```{r}
yhat_lin <- as.vector(X_lin %*% fit_lin$coefficients)
yhat_spl <- as.vector(X_spl %*% fit_spl$coefficients)

ord <- order(x)
plot(x[ord], yhat_lin[ord], type = "l", xlab = xname, ylab = paste("Predicted median", yname))
lines(x[ord], yhat_spl[ord], type = "l")

```

Step 6 - Smooth grid

```{r}
xg <- seq(min(x, na.rm = T), max(x, na.rm = T), length.out = 200)

Zg <- matrix(colMeans(Z), nrow = length(xg), ncol = ncol(Z), byrow = T)

Xg_lin <- cbind(Intercept = 1, x = xg, Zg)

Bg <- ns(xg, knots = kn)
Xg_spl <-  cbind(Intercept = 1, Bg, Zg)

yhat_lin_g <- as.vector(Xg_lin %*% fit_lin$coefficients)
yhat_spl_g <- as.vector(Xg_spl %*% fit_spl$coefficients)

plot(xg, yhat_lin_g, type = "l", lwd = 2,
     xlab = xname, ylab = paste("Predicted median", yname))
lines(xg, yhat_spl_g, lwd = 2)
legend("topright", legend = c("Linear", "Spline"), lty = 1, bty = "n")

```


CONTRAST BETWEEN Q5 AND Q1 (MODEL-BASED)

```{r}

qreg_contrast <- function(data, outcome, quintile_var, covars, 
                      tau = 0.5,
                      R = 2000, 
                      seed = 123) {
  
  #make Q1 the reference
  data[[quintile_var]] <- factor(data[[quintile_var]], levels = 1:5)
  data[[quintile_var]] <- stats::relevel(data[[quintile_var]], ref = "1")
  
  form <- as.formula(
    paste0(outcome, "~", quintile_var, "+", paste(covars, collapse = "+"))
  )
  
  fit <- rq(form, tau = tau, data = data)
  
  set.seed(seed)
  sm <- summary(fit, se = "boot", R = R)
  
  #Extract rows
  coef_name <- paste0(quintile_var, "5")
  row <- sm$coefficients[coef_name, ]

  beta <- row["Value"]
  se <- row["Std. Error"]
  ci <- c(beta -1.96 * se, beta + 1.96 * se)
  
  #Return as a one-row data frame
  data.frame(
    outcome = outcome,
    q5_vs_q1 = beta,
    ci_lower = ci[1],
    ci_upper = ci[2],
    stringsAsFactors = F
  )
}
```

Apply function to all outcomes

```{r}
results_contrast <- do.call(
  rbind,
  lapply(dep_vars, qreg_contrast,
         data = vux,
         quintile_var = "quintile_mid",
         covars = covars,
         tau = 0.5,
         R = 2000,
         seed = 123)
)

```

```{r}
print(results_contrast)
```

Apply rounding function

```{r}
results_contrast_clean <- results_contrast %>% 
  rowwise() %>% 
  mutate(
    q5_vs_q1 = round_by_var(outcome, q5_vs_q1, "q5_vs_q1"),
    ci_lower = round_by_var(outcome, ci_lower, "ci"),
    ci_upper = round_by_var(aoutcome, ci_upper, "ci")
  ) %>% 
  ungroup()

results_contrast_clean
```

Format

```{r}
results_contrast_clean$CI <- paste0("(", results_contrast_clean$ci_lower, ", ", results_contrast_clean$ci_upper, ")")

results_contrast_clean <- results_contrast_clean[, c("outcome", "q5_vs_q1", "CI")]

results_contrast_clean
```

2.  LOWER BOUND

Apply function for quantile regression to all outcomes

```{r}
results_low <- do.call(
  rbind,
  lapply(dep_vars, qreg_orig,
         data = vux,
         predictor = "EP_upf_low_10",
         covars = covars,
         tau = 0.5,
         R = 2000,
         seed = 123)
)
```

Check stability

```{r}
qreg_orig(vux, outcome="kcal", predictor = "EP_upf_low_10", covars = covars, seed=1)$beta

qreg_orig(vux, outcome="kcal", predictor = "EP_upf_low_10", covars = covars, seed=2)$beta
```

```{r}
print(results_low)
```

Adjust for multiple testing

```{r}
results_low$padj <- p.adjust(results_low$p_value, method = "BH")
```

Apply rounding function

```{r}
results_clean_low <- results_low %>% 
  rowwise() %>% 
  mutate(
    beta = round_by_var(outcome, beta, "beta"),
    ci_lower = round_by_var(outcome, ci_lower, "ci"),
    ci_upper = round_by_var(aoutcome, ci_upper, "ci")
  ) %>% 
  ungroup()
```

Format

```{r}
results_clean_low$CI <- paste0("(", results_clean_low$ci_lower, ", ", results_clean_low$ci_upper, ")")

results_clean_low <- results_clean_low[, c("outcome", "beta", "CI", "padj")]
colnames(results_clean_low)[4] <- "p_adj"

results_clean_low
```

Contrast

Apply function to all outcomes

```{r}
results_contrast_low <- do.call(
  rbind,
  lapply(dep_vars, qreg_contrast,
         data = vux,
         quintile_var = "quintile_low",
         covars = covars,
         tau = 0.5,
         R = 2000,
         seed = 123)
)

```

```{r}
print(results_contrast_low)
```

Apply rounding function

```{r}
results_contrast_clean_low <- results_contrast_low %>% 
  rowwise() %>% 
  mutate(
    q5_vs_q1 = round_by_var(outcome, q5_vs_q1, "q5_vs_q1"),
    ci_lower = round_by_var(outcome, ci_lower, "ci"),
    ci_upper = round_by_var(aoutcome, ci_upper, "ci")
  ) %>% 
  ungroup()

results_contrast_clean_low
```

Format

```{r}
results_contrast_clean_low$CI <- paste0("(", results_contrast_clean_low$ci_lower, ", ", results_contrast_clean_low$ci_upper, ")")

results_contrast_clean_low <- results_contrast_clean_low[, c("outcome", "q5_vs_q1", "CI")]

results_contrast_clean_low
```

Export to excel

```{r}

wb <- createWorkbook()

addWorksheet(wb, "N complete cases")
writeData(wb, sheet = "N complete cases", n_obs_df)

addWorksheet(wb, "N zero counts")
writeData(wb, sheet = "N zero counts", zero_counts_df)

addWorksheet(wb, "Clean_mid")
writeData(wb, sheet = "Clean_mid", results_clean)

addWorksheet(wb, "Contrast_mid")
writeData(wb, sheet = "Contrast_mid", results_contrast_clean)

addWorksheet(wb, "Clean_low")
writeData(wb, sheet = "Clean_low", results_clean_low)

addWorksheet(wb, "Contrast_low")
writeData(wb, sheet = "Contrast_low", results_contrast_clean_low)


saveWorkbook(wb, 
             file = file.path(dir_output, "Quantreg_results_vux.xlsx"),
             overwrite = T
          )
```
