---
title: "Quantile regression with intake as pct of AR as dependent variable"
format: html
editor: visual
---

This script runs quantile regression with intake of vitamins and mineras as % of average requirement (AR) as dependent variables. Adolescents dataset.

```{r}
#| label: setup
#| include: false

project_root <- normalizePath(file.path(getwd(), ".."))

dir_data <- file.path(project_root, "data")
dir_output <- file.path(project_root, "output")

dir.create(dir_output, showWarnings = FALSE, recursive = TRUE)

```

Analytic code starts here

```{r message=FALSE, warning=FALSE}
library(openxlsx)
library(dplyr)
library(quantreg)
```

Import data (choose dataset)

```{r}
master <- readRDS(file.path(dir_data, "Masterfile.rds")) 

```

Create dataset for regression

```{r}
ungdom <- master %>% 
  filter(RM == "Ungdom") %>% 
  dplyr::select(person_ID, sex, age, age_grp, bmi, edu_combined, leisure, kcal, pct_vitD, pct_vitE, pct_folate, pct_vitB12, pct_vitC, pct_fe, pct_ca, pct_k, pct_se, EP_upf_low, EP_upf_up, EP_upf_mid)
```

Convert selected variables to factors

```{r}
ungdom <- ungdom %>% 
  mutate(
    sex = as.factor(sex),
    edu_combined = as.factor(edu_combined),
    leisure = as.factor(leisure),
    age_grp = as.factor(age_grp)
  )

```

Scale predictor variable

```{r}
ungdom$EP_upf_low_10 <- ungdom$EP_upf_low / 10
ungdom$EP_upf_mid_10 <- ungdom$EP_upf_mid / 10
```

Define dependent variables

```{r}
dep_vars <- names(ungdom)[which(names(ungdom) == "pct_vitD"):which(names(ungdom) == "pct_se")]
```

Define covariates for model 1 and model 2

```{r}
covars_m1 <- c("age_grp", "sex", "edu_combined", "leisure")
covars_m2 <- c("age_grp", "sex", "edu_combined", "leisure", "kcal")

```

Count complete cases

```{r}

n_total <- nrow(ungdom)

n_complete <- sum(complete.cases(ungdom[, covars_m1]))


n_obs_df <- data.frame(
  n_total = n_total,
  n_complete = n_complete,
  perc_complete = round(100 * n_complete / n_total, 1)
)
```

QUANTILE REGRESSION - Middle bound scenario

Model 1 (main model)

Define function to run qr on all outcomes

```{r}
qreg_orig <- function(data, outcome, predictor, covars_m1, 
                      tau = 0.5,
                      R = 2000, 
                      seed = 123) {
  
  form <- as.formula(
    paste0(outcome, "~", predictor, "+", paste(covars_m1, collapse = "+"))
  )
  
  fit <- rq(form, tau = tau, data = data)
  set.seed(seed)
  sm <- summary(fit, se = "boot", R = R)
  
  #Extract row for predictor
  row <- sm$coefficients[predictor, ]
  beta <- row["Value"]
  se <- row["Std. Error"]
  pval <- row["Pr(>|t|)"]
  
  ci <- c(beta -1.96 * se, beta + 1.96 * se)
  
  #Return as a one-row data frame
  data.frame(
    outcome = outcome,
    beta = beta,
    ci_lower = ci[1],
    ci_upper = ci[2],
    p_value = pval,
    stringsAsFactors = F
  )
  
}
```

Apply function to all outcomes

```{r}
results <- do.call(
  rbind,
  lapply(dep_vars, qreg_orig,
         data = ungdom,
         predictor = "EP_upf_mid_10",
         covars = covars_m1,
         tau = 0.5,
         R = 2000,
         seed = 123)
)
```

Check stability

```{r}
qreg_orig(ungdom, outcome="pct_vitD", predictor = "EP_upf_mid_10", covars = covars_m1, seed=1)$beta

qreg_orig(ungdom, outcome="pct_vitD", predictor = "EP_upf_mid_10", covars = covars_m1, seed=2)$beta
```

```{r}
print(results)
```

Adjust for multiple testing

```{r}
results$padj <- p.adjust(results$p_value, method = "BH")
```

Rounding

```{r}
results_m1 <- results %>% 
  mutate(
    across(c(beta, ci_lower, ci_upper), ~round(.x, 1)),
    padj = round(padj, 3)
  )
```

```{r}
print(results_m1)
```

Model 2 (energy adjusted)

Apply function for quantile regression to all outcomes

```{r}
results_m2 <- do.call(
  rbind,
  lapply(dep_vars, qreg_orig,
         data = ungdom,
         predictor = "EP_upf_mid_10",
         covars = covars_m2,
         tau = 0.5,
         R = 2000,
         seed = 123)
)
```

Check stability

```{r}
qreg_orig(ungdom, outcome="pct_fe", predictor = "EP_upf_mid_10", covars = covars_m2, seed=1)$beta

qreg_orig(ungdom, outcome="pct_fe", predictor = "EP_upf_mid_10", covars = covars_m2, seed=2)$beta
```

```{r}
print(results_m2)
```

Adjust for multiple testing

```{r}
results_m2$padj <- p.adjust(results_m2$p_value, method = "BH")
```

Rounding

```{r}
results_m2 <- results_m2 %>% 
  mutate(
    across(c(beta, ci_lower, ci_upper), ~round(.x, 1)),
    padj = round(padj, 3)
  )
```

```{r}
print(results_m2)
```

Export to excel

```{r}

wb <- createWorkbook()

addWorksheet(wb, "N complete cases")
writeData(wb, sheet = "N complete cases", n_obs_df)

addWorksheet(wb, "model_1")
writeData(wb, sheet = "model_1", results_m1)

addWorksheet(wb, "model_2")
writeData(wb, sheet = "model_2", results_m2)

saveWorkbook(wb, 
             file = file.path("AR_results_ungdom.xlsx"), 
             overwrite = TRUE
            )
```
