---
title: "Nutrients vs AR, diagrams"
author: "Torsten Sällström"
format: html
editor: visual
---

This script creates lineplots over vitamins and minerals intake based on quintiles of E% upf.

```{r}
#| label: setup
#| include: false

project_root <- normalizePath(file.path(getwd(), ".."))

dir_data <- file.path(project_root, "data")
dir_output <- file.path(project_root, "output")

dir.create(dir_output, showWarnings = FALSE, recursive = TRUE)

```

Analytic code starts here

```{r message=FALSE, warning=FALSE}

library(patchwork)
library(openxlsx)
library(dplyr)
library (tidyr)
library(ggplot2)

```

```{r}
master <- readRDS(file.path(dir_data,"Masterfile.rds"))
         
```

Prepare plotdata for adults

```{r}

# Divide EP_upf_mid into quintiles
vux <- master %>%
  filter(RM == "Vuxna") %>% 
  mutate(quintile = paste0("Q", ntile(EP_upf_mid, 5))
)
#Group by quintiles, calculate summary statistics
vux_summary <- vux %>% 
  group_by(quintile) %>% 
  summarise(across(pct_vitD:pct_se, list(
      median = ~median(.x, na.rm = T),
      p25 = ~quantile(.x, 0.25, na.rm = T),
      p75 = ~quantile(.x, 0.75, na.rm = T)
    ), .names = "{.col}_{.fn}"))

```

Histograms to check distribution

```{r}

ggplot(vux[vux$quintile == "Q3", ], aes(x = pct_vitC))+
  geom_histogram(bins = 30, fill = "skyblue", color = "black")
```

Data is heavily skewed. Medians will be reported.

Reshape to long format

```{r}
vux_long <- vux_summary %>% 
  pivot_longer(
    cols = -quintile,
    names_to = c("variable", "stat"),
    names_pattern = "^(.*)_(median|p25|p75)$"
  ) %>% 
  pivot_wider(
    names_from = stat,
    values_from = value
  ) %>% 
  mutate(group = "Adults")

```

Prepare plotdata for adolescents

```{r}

# Divide EP_upf_mid into quintiles
ungdom <- master %>%
  filter(RM == "Ungdom") %>% 
  mutate(quintile = paste0("Q", ntile(EP_upf_mid, 5))
)
#Group by quintiles, calculate summary statistics
ungdom_summary <- ungdom %>% 
  group_by(quintile) %>% 
  summarise(across(pct_vitD:pct_se, list(
      median = ~median(.x, na.rm = T),
      p25 = ~quantile(.x, 0.25, na.rm = T),
      p75 = ~quantile(.x, 0.75, na.rm = T)
    ), .names = "{.col}_{.fn}"))

```

Reshape to long format

```{r}
ungdom_long <- ungdom_summary %>% 
  pivot_longer(
    cols = -quintile,
    names_to = c("variable", "stat"),
    names_pattern = "^(.*)_(median|p25|p75)$"
  ) %>% 
  pivot_wider(
    names_from = stat,
    values_from = value
  ) %>% 
  mutate(group = "Adolescents")

```

Prepare plot data for children

```{r}

# Divide EP_upf_mid into quintiles
barn <- master %>%
  filter(RM == "Barn") %>% 
  mutate(quintile = paste0("Q", ntile(EP_upf_mid, 5))
)

#Group by quintiles, calculate summary statistics
barn_summary <- barn %>% 
  group_by(quintile) %>% 
  summarise(across(pct_vitD:pct_se, list(
      median = ~median(.x, na.rm = T),
      p25 = ~quantile(.x, 0.25, na.rm = T),
      p75 = ~quantile(.x, 0.75, na.rm = T)
    ), .names = "{.col}_{.fn}"))

```

Reshape to long format

```{r}
barn_long <- barn_summary %>% 
  pivot_longer(
    cols = -quintile,
    names_to = c("variable", "stat"),
    names_pattern = "^(.*)_(median|p25|p75)$"
  ) %>% 
  pivot_wider(
    names_from = stat,
    values_from = value
  ) %>% 
  mutate(group = "Young Children")

```

```{r}
colnames(ungdom_long)
```

Merge datasets, add p-values

```{r}
df_long <- bind_rows(vux_long, ungdom_long, barn_long)
```

```{r}
unique(df_long$group)

```

Convert grouping variable to factor

```{r}

df_long$group <- factor(df_long$group, levels = c("Adults", "Adolescents", "Young Children"))
```

1)  Define custom variable titles

```{r}
unique(df_long$variable)
```

```{r}
plot_titles <- c(
  pct_vitD = "Vitamin D",
  pct_vitE = "Vitamin E",
  pct_folate = "Folate",
  pct_vitB12 = "Vitamin B12",
  pct_vitC = "Vitamin C",
  pct_fe = "Iron",
  pct_ca = "Calcium",
  pct_k = "Potassium",
  pct_se = "Selenium"
 )
```

2)  Create variables object and a list to store plots

```{r}
variables <- unique(df_long$variable)

#Create a list to store plots
plot_list <-  list()
```

IQR is used as error bars

```{r}
for(var in variables){
  data_subset <- df_long %>% filter(variable == var)
 
  p <- ggplot(data_subset, aes(x = quintile, y = median, group = group, color = group, linetype = group, shape = group)) +
    theme_classic()+
    geom_line(size = 1)+
     scale_linetype_manual(values = c("Adults" = "solid", "Adolescents" = "dashed", "Young Children" = "dashed"))+
    geom_point(position = position_dodge(width = 0.3), size = 4)+
     scale_shape_manual(values = c("Adults" = 16, "Adolescents" = 15, "Young Children" = 17)) +
     scale_color_manual(values = c("Adults" = "royalblue3", "Adolescents" = "chartreuse3", "Young Children" = "indianred3"))+
     guides(
      color = guide_legend(override.aes = list(shape = c(16, 15, 17),
                                               linetype = c("solid", "dashed", "dashed"))),
      shape = "none",
      linetype = "none"
    ) +

    theme(axis.text.x = element_text(size = 16),
          axis.text.y = element_text(size = 14)) +

    geom_errorbar(aes(ymin = p25, ymax = p75), 
                  position = position_dodge(width = 0.3),
                  width = 0.1,
                  linetype = "solid") +
    geom_hline(yintercept = 100, linetype = "dotted", color = "black") +
    annotate("text", x = 1, y = 102, label = "AR", hjust = 2, vjust = 0, size = 4) +
  
     labs(title = plot_titles[[var]],
           x = "quintiles of E% UPF",
           y = "Median intake (% of AR)",
           color = NULL,
           linetype = NULL)+

      theme(plot.title = element_text(size = 16, face = "bold"),
            panel.grid = element_blank(),
            legend.position = "bottom",
            plot.margin = margin(10, 10, 10, 30), # (top, right, bottom, left)
            axis.line = element_line(color = "black"),
            axis.ticks = element_line(color = "black"),
            axis.title.x = element_blank(),
            axis.title.y = element_blank()
          )
  
 
  print(p)
  
  plot_list[[var]] <-p
}

```

Combine plots into one image

```{r}

library(patchwork)

th <- theme(
        legend.position = "none",
        legend.box.margin = margin(t=40),
        plot.margin = margin(10, 10, 10, 30), # (top, right, bottom, left)
        legend.title = element_blank(),
        legend.text = element_text(size = 20),
        legend.key.size = unit(1.2, "lines")
  ) 

plot_list2 <- lapply(plot_list, function(p) p + theme(legend.position = "none"))

lineplots <- wrap_plots(plot_list2, ncol = 3) +
  plot_layout(guides = "collect")

print(lineplots)

```

Combine plots into one figure

```{r}
# Save the combined plot
ggsave(
  filename = file.path(dir_output, "Lineplots.png"), 
  plot = lineplots, 
  width = 16, 
  height = 12,
  dpi = 300
)

```

```{r}
# Save the combined plot
ggsave(
  filename = file.path(dir_output, "Lineplots.pdf"), 
  plot = lineplots, 
  width = 16, 
  height = 12,
  dpi = 300
)

```


Extract legend only

```{r}
library(grid)
library(gridExtra)

p1 <- plot_list[[1]] + theme(legend.position = "bottom")
g <- ggplotGrob(p1)
idx <- which(g$layout$name == "guide-box-bottom")

leg <- g$grobs[[idx]]

grid.draw(leg)

legend_grob <- arrangeGrob(leg)

```

Save legend as image

```{r}
ggsave(
  filename = file.path(dir_output, "Legend_only.png"),
  plot = legend_grob,
  width = 6,
  height = 1.2,
  units = "in",
  dpi = 300,
  bg = "white"
)
```


```{r}
ggsave(
  filename = file.path(dir_output, "Legend_only.pdf"),
  plot = legend_grob,
  width = 6,
  height = 1.2,
  units = "in",
  dpi = 300,
  bg = "white"
)
```





