---
title: "Linear regression of BMI-z and weight status among adolescents"
format: html
editor: visual
---

1)  Multiple imputation
2)  Linear regression with logBMI as dependent variable
3)  Poisson regression with robust variance for Prevalence ratios of overweight/obesity Dataset: Adolescents

```{r}
#| label: setup
#| include: false

project_root <- normalizePath(file.path(getwd(), ".."))

dir_data <- file.path(project_root, "data")
dir_output <- file.path(project_root, "output")

dir.create(dir_output, showWarnings = FALSE, recursive = TRUE)

```

Analytic code starts here

```{r message=FALSE, warning=FALSE}
library(openxlsx)
library(dplyr)
library (tidyr)
library(stringr)
library(purrr)
library(forcats)
```

Import data

```{r}
master <- readRDS(file.path(dir_data, "Masterfile.rds")) 
```

Create dataset for regression

```{r}
ungdom1 <- master %>% 
  filter(RM == "Ungdom") %>% 
  dplyr::select(person_ID, sex, age, age_grp, edu_combined, leisure, nrf, kcal, ED_nobev, bmi, bmi_z, overweight, EP_upf_low, EP_upf_up, EP_upf_mid)

```

Exclude individuals older than 18 years

```{r}
sum(ungdom1$age >=18, na.rm = T)

ungdom <- ungdom1[ungdom1$age <18, ]
```

Convert selected variables to factors

```{r}
ungdom <- ungdom %>% 
  mutate(
    sex = as.factor(sex),
    edu_combined = as.factor(edu_combined),
    leisure = as.factor(leisure),
    age_grp = as.factor(age_grp),
    overweight = as.factor(overweight),
  )

```

Divide the dataset in tertiles based on the variable EP_upf

```{r}
ungdom <- ungdom %>% 
  mutate(
    tertile_low = factor(paste0("T", ntile(EP_upf_low, 3)), levels = paste0("T", 1:3)),
    tertile_mid = factor(paste0("T", ntile(EP_upf_mid, 3)), levels = paste0("T", 1:3)),
    tertile_up = factor(paste0("T", ntile(EP_upf_up, 3)), levels = paste0("T", 1:3)),
  )

```

Scale predictor variable

```{r}
ungdom$EP_upf_low_10 <- ungdom$EP_upf_low / 10
ungdom$EP_upf_mid_10 <- ungdom$EP_upf_mid / 10
ungdom$EP_upf_up_10 <- ungdom$EP_upf_up / 10
```

Count number of individuals with NA for bmi_z

```{r}
sum(is.na(ungdom$bmi_z))
```

Exclude individuals with NA for the outcome before imputation

```{r}
ungdom <- ungdom[!is.na(ungdom$bmi_z),]

nrow(ungdom)

```

Double check so that overweight category is correctly defined

```{r}
ungdom <- ungdom %>% 
  mutate(
    overweight_check = case_when(
      is.na(bmi_z) ~NA_character_,
      sex == "M" & bmi_z >= 1.310 ~"yes",
      sex == "F" & bmi_z >= 1.244 ~"yes",
      TRUE ~"no"
    ),
overweight_check = factor(overweight_check, levels = levels(overweight))
 )

#compare
ungdom %>% count(overweight, overweight_check)
```

Values for BMI-z score in the range -5 to +5 are deemed biologically plausible. Other values are removed.

```{r}
ungdom <- ungdom[ungdom$bmi_z >= -5 & ungdom$bmi_z <=5, ]

nrow(ungdom)
```

Fit model for standard diagnostic tests

```{r}
library(car)
fit <- lm(bmi_z ~EP_upf_mid_10 + age_grp + sex + leisure + edu_combined + kcal + nrf, data = ungdom)

fit
plot(fit)
vif(fit)
```


MULTIPLE IMPUTATION

```{r}
library(mice)
library(miceadds)
library(broom)

```

Prepare for imputation

```{r}

imp_data <- as.data.frame(ungdom)

#Define variables to exclude from imputation
exclude_vars <- c( "bmi", "bmi_z", "overweight", "tertile_low", "tertile_mid", "tertile_up")

pred <- make.predictorMatrix(imp_data)

#Outcomes not used to predict, not imputed
pred[, exclude_vars] <- 0
pred[exclude_vars,] <- 0

#ID not used either
pred[, "person_ID"] <- 0
pred["person_ID", ] <- 0

#Method vector
meth <- make.method(imp_data)

#Outcomes and ID should not be imputed
meth[exclude_vars] <- ""
meth["person_ID"] <- ""

#Sanity checks
stopifnot(is.matrix(pred))
stopifnot(all(rownames(pred) == names(imp_data)))
stopifnot(all(colnames(pred)== names(imp_data)))

```

Run mice

```{r}
ungdom_imputed <- mice(
  imp_data,
  m = 20,
  predictorMatrix = pred,
  method = meth,
  seed = 123
)

```

```{r}
summary(ungdom_imputed)
```

Inspect methods

```{r}
ungdom_imputed$method
```

```{r}
colSums(ungdom_imputed$where)
```

MIDDLE BOUND

CRUDE MODEL

```{r}
crude <- with(
  ungdom_imputed,
  lm(bmi_z ~EP_upf_mid_10)
)

crude_pooled <- pool(crude)
crude_pooled_sum <- summary(crude_pooled, conf.int = T)


n_per_imp <- sapply(crude$analyses, function(mod) nrow(model.frame(mod)))
unique_n <- unique(n_per_imp)
unique_n
crude_pooled_sum
```

MODEL 1 - sociodemographics only

```{r}
model1 <- with(
  ungdom_imputed,
  lm(bmi_z ~EP_upf_mid_10 + age_grp + sex + edu_combined + leisure )
)

model1_pooled <- pool(model1)
model1_pooled_sum <- summary(model1_pooled, conf.int = T)


n_per_imp <- sapply(model1$analyses, function(mod) nrow(model.frame(mod)))
unique_n <- unique(n_per_imp)
unique_n
model1_pooled_sum
```

MODEL 2 - sociodemographics + energy intake

```{r}
model2 <- with(
  ungdom_imputed,
  lm(bmi_z ~EP_upf_mid_10 + age_grp + sex + edu_combined + leisure  + kcal)
)

model2_pooled <- pool(model2)
model2_pooled_sum <- summary(model2_pooled, conf.int = T)


n_per_imp <- sapply(model2$analyses, function(mod) nrow(model.frame(mod)))
unique_n <- unique(n_per_imp)
unique_n
model2_pooled_sum
```

MODEL 3 - sodiodemographics + energy + nrf

```{r}
model3 <- with(
  ungdom_imputed,
  lm(bmi_z ~EP_upf_mid_10 + age_grp + sex + edu_combined + leisure + kcal + nrf)
)

model3_pooled <- pool(model3)
model3_pooled_sum <- summary(model3_pooled, conf.int = T)


n_per_imp <- sapply(model3$analyses, function(mod) nrow(model.frame(mod)))
unique_n <- unique(n_per_imp)
unique_n
model3_pooled_sum
```

Export to excel

```{r}
write.xlsx(
  list(
    Crude = crude_pooled_sum,
    Model1 = model1_pooled_sum,
    Model2 = model2_pooled_sum,
    Model3 = model3_pooled_sum
  ),
  file = file.path(dir_output, "Regression_BMI_ungdom_mid.xlsx"),
   overwrite = TRUE
)
```

LOWER BOUND

CRUDE MODEL

```{r}
crude <- with(
  ungdom_imputed,
  lm(bmi_z ~EP_upf_low_10)
)

crude_pooled <- pool(crude)
crude_pooled_sum_low <- summary(crude_pooled, conf.int = T)


n_per_imp <- sapply(crude$analyses, function(mod) nrow(model.frame(mod)))
unique_n <- unique(n_per_imp)
unique_n
crude_pooled_sum_low
```

MODEL 1 - Sociodemographics only

```{r}
model1 <- with(
  ungdom_imputed,
  lm(bmi_z ~EP_upf_low_10 + age_grp + sex + edu_combined + leisure )
)

model1_pooled <- pool(model1)
model1_pooled_sum_low <- summary(model1_pooled, conf.int = T)


n_per_imp <- sapply(model1$analyses, function(mod) nrow(model.frame(mod)))
unique_n <- unique(n_per_imp)
unique_n
model1_pooled_sum_low
```

MODEL 2 - Sociodemographics + EI

```{r}
model2 <- with(
  ungdom_imputed,
  lm(bmi_z ~EP_upf_low_10 + age_grp + sex + edu_combined + leisure  + kcal)
)

model2_pooled <- pool(model2)
model2_pooled_sum_low <- summary(model2_pooled, conf.int = T)


n_per_imp <- sapply(model2$analyses, function(mod) nrow(model.frame(mod)))
unique_n <- unique(n_per_imp)
unique_n
model2_pooled_sum_low
```

MODEL 3

```{r}
model3 <- with(
  ungdom_imputed,
  lm(bmi_z ~EP_upf_low_10 + age_grp + sex + edu_combined + leisure  + kcal + nrf)
)

model3_pooled <- pool(model3)
model3_pooled_sum_low <- summary(model3_pooled, conf.int = T)


n_per_imp <- sapply(model3$analyses, function(mod) nrow(model.frame(mod)))
unique_n <- unique(n_per_imp)
unique_n
model3_pooled_sum_low
```

Export to excel

```{r}
write.xlsx(
  list(
    Crude = crude_pooled_sum_low,
    Model1 = model1_pooled_sum_low,
    Model2 = model2_pooled_sum_low,
    Model3 = model3_pooled_sum_low
  ),
  file = file.path(dir_output, "Regression_BMI_ungdom_low.xlsx"),
   overwrite = TRUE
)
```

UPPER BOUND

CRUDE MODEL

```{r}
crude <- with(
  ungdom_imputed,
  lm(bmi_z ~EP_upf_up_10)
)

crude_pooled <- pool(crude)
crude_pooled_sum_up <- summary(crude_pooled, conf.int = T)


n_per_imp <- sapply(crude$analyses, function(mod) nrow(model.frame(mod)))
unique_n <- unique(n_per_imp)
unique_n
crude_pooled_sum_up
```

MODEL 1 - Sociodemographics only

```{r}
model1 <- with(
  ungdom_imputed,
  lm(bmi_z ~EP_upf_up_10 + age_grp + sex + edu_combined + leisure )
)

model1_pooled <- pool(model1)
model1_pooled_sum_up <- summary(model1_pooled, conf.int = T)


n_per_imp <- sapply(model1$analyses, function(mod) nrow(model.frame(mod)))
unique_n <- unique(n_per_imp)
unique_n
model1_pooled_sum_up
```

MODEL 2 - Sociodemographics + EI

```{r}
model2 <- with(
  ungdom_imputed,
  lm(bmi_z ~EP_upf_up_10 + age_grp + sex + edu_combined + leisure  + kcal)
)

model2_pooled <- pool(model2)
model2_pooled_sum_up <- summary(model2_pooled, conf.int = T)


n_per_imp <- sapply(model2$analyses, function(mod) nrow(model.frame(mod)))
unique_n <- unique(n_per_imp)
unique_n
model2_pooled_sum_up
```

MODEL 3 - Sociodemographics + kcal + nrf

```{r}
model3 <- with(
  ungdom_imputed,
  lm(bmi_z ~EP_upf_up_10 + age_grp + sex + edu_combined + leisure  + kcal + nrf)
)

model3_pooled <- pool(model3)
model3_pooled_sum_up <- summary(model3_pooled, conf.int = T)


n_per_imp <- sapply(model3$analyses, function(mod) nrow(model.frame(mod)))
unique_n <- unique(n_per_imp)
unique_n
model3_pooled_sum_up
```

Export to excel

```{r}
write.xlsx(
  list(
    Crude = crude_pooled_sum_up,
    Model1 = model1_pooled_sum_up,
    Model2 = model2_pooled_sum_up,
    Model3 = model3_pooled_sum_up
  ),
  file = file.path(dir_output, "Regression_BMI_ungdom_up.xlsx"),
   overwrite = TRUE
)

```

PREVALENCE RATIOS

Load packages for robust SE

```{r}
library(mitools)
library(sandwich)
library(tibble)
library(miceadds)
library(lmtest)
```

define covars

```{r}
covars0 <- character(0)
covars1 <- c("age_grp", "sex", "edu_combined", "leisure")
covars2 <- c("age_grp", "sex", "edu_combined", "leisure", "kcal")
covars3 <- c("age_grp", "sex", "edu_combined", "leisure", "kcal", "nrf")
```

MIDDLE-BOUND

CRUDE MODEL

```{r}

covars <- covars0

```

```{r}
models <- lapply(1:20, function(i) {
  dat <- complete(ungdom_imputed, i)
  
  dat$overweight <- as.numeric(dat$overweight == "yes")
  
  dat <- subset(dat, !is.na(overweight))
  
  dat$tertile_mid <- factor(dat$tertile_mid, levels = c("T1", "T2", "T3"))
  
  form <- reformulate(
    termlabels = c("tertile_mid", covars),
    response = "overweight"
  )
  
  glm(form, family = poisson(link = "log"), data = dat)
})

```

Extract robust estimates

```{r}
robust <- lapply(models, function(m) {
  ct <- coeftest(m, vcov. = vcovHC(m, type = "HC0"))
  list(
    coef = coef(m),
    vcov = vcovHC(m, type = "HC0"),
    ct = ct
  )
})
```

Confirm coefficient names

```{r}
names(robust[[1]]$coef)
```

Chose the terms to pool

```{r}
terms_to_pool <- c("tertile_midT2", "tertile_midT3")
```

Create a pooling function (Rubin's rules for one term)

```{r}
pool_pr_one_term <- function(robust, term) {
  m <- length(robust)
  
  Q <- sapply(robust, function(x) x$coef[term])
  U <- sapply(robust, function(x) x$vcov[term, term])
  
  Qbar <- mean(Q)
  Ubar <- mean(U)
  B <- var(Q)
  T <- Ubar + (1+1/m)*B
  se <- sqrt(T)
  
  z <- Qbar / se
  p <- 2 * pnorm(-abs(z)) #two-sided Wald p-value
  
  data.frame(
    term = term,
    PR <- exp(Qbar),
    LCL <- exp(Qbar - 1.96 * se),
    UCL <- exp(Qbar + 1.96 * se),
    logPR = Qbar,
    SE = se,
    z = z,
    p_value = p,
    stringsAsFactors = FALSE
  )
}
```

Build summary table

```{r}
PR_summary_crude_mid <- do.call(rbind, lapply(terms_to_pool, function(trm) {
  pool_pr_one_term(robust, trm)
}))

PR_summary_crude_mid
```

MODEL 1

```{r}

covars <- covars1

```

```{r}
models <- lapply(1:20, function(i) {
  dat <- complete(ungdom_imputed, i)
  
  dat$overweight <- as.numeric(dat$overweight == "yes")
  
  dat <- subset(dat, !is.na(overweight))
  
  dat$tertile_mid <- factor(dat$tertile_mid, levels = c("T1", "T2", "T3"))
  
  form <- reformulate(
    termlabels = c("tertile_mid", covars),
    response = "overweight"
  )
  
  glm(form, family = poisson(link = "log"), data = dat)
})

```

Extract robust estimates

```{r}
robust <- lapply(models, function(m) {
  ct <- coeftest(m, vcov. = vcovHC(m, type = "HC0"))
  list(
    coef = coef(m),
    vcov = vcovHC(m, type = "HC0"),
    ct = ct
  )
})
```

Confirm coefficient names

```{r}
names(robust[[1]]$coef)
```

Chose the terms to pool

```{r}
terms_to_pool <- c("tertile_midT2", "tertile_midT3")
```

Build summary table

```{r}
PR_summary_model1_mid <- do.call(rbind, lapply(terms_to_pool, function(trm) {
  pool_pr_one_term(robust, trm)
}))

PR_summary_model1_mid
```

MODEL 2

```{r}

covars <- covars2

```

```{r}
models <- lapply(1:20, function(i) {
  dat <- complete(ungdom_imputed, i)
  
  dat$overweight <- as.numeric(dat$overweight == "yes")
  
  dat <- subset(dat, !is.na(overweight))
  
  dat$tertile_mid <- factor(dat$tertile_mid, levels = c("T1", "T2", "T3"))
  
  form <- reformulate(
    termlabels = c("tertile_mid", covars),
    response = "overweight"
  )
  
  glm(form, family = poisson(link = "log"), data = dat)
})

```

Extract robust estimates

```{r}
robust <- lapply(models, function(m) {
  ct <- coeftest(m, vcov. = vcovHC(m, type = "HC0"))
  list(
    coef = coef(m),
    vcov = vcovHC(m, type = "HC0"),
    ct = ct
  )
})
```

Confirm coefficient names

```{r}
names(robust[[1]]$coef)
```

Chose the terms to pool

```{r}
terms_to_pool <- c("tertile_midT2", "tertile_midT3")
```

Build summary table

```{r}
PR_summary_model2_mid <- do.call(rbind, lapply(terms_to_pool, function(trm) {
  pool_pr_one_term(robust, trm)
}))

PR_summary_model2_mid
```

MODEL 3

```{r}

covars <- covars3

```

```{r}
models <- lapply(1:20, function(i) {
  dat <- complete(ungdom_imputed, i)
  
  dat$overweight <- as.numeric(dat$overweight == "yes")
  
  dat <- subset(dat, !is.na(overweight))
  
  dat$tertile_mid <- factor(dat$tertile_mid, levels = c("T1", "T2", "T3"))
  
  form <- reformulate(
    termlabels = c("tertile_mid", covars),
    response = "overweight"
  )
  
  glm(form, family = poisson(link = "log"), data = dat)
})

```

Extract robust estimates

```{r}
robust <- lapply(models, function(m) {
  ct <- coeftest(m, vcov. = vcovHC(m, type = "HC0"))
  list(
    coef = coef(m),
    vcov = vcovHC(m, type = "HC0"),
    ct = ct
  )
})
```

Confirm coefficient names

```{r}
names(robust[[1]]$coef)
```

Chose the terms to pool

```{r}
terms_to_pool <- c("tertile_midT2", "tertile_midT3")
```

Build summary table

```{r}
PR_summary_model3_mid <- do.call(rbind, lapply(terms_to_pool, function(trm) {
  pool_pr_one_term(robust, trm)
}))

PR_summary_model3_mid
```

LOWER BOUND - MODEL 1

```{r}

covars <- covars1

```

```{r}
models <- lapply(1:20, function(i) {
  dat <- complete(ungdom_imputed, i)
  
  dat$overweight <- as.numeric(dat$overweight == "yes")
  
  dat <- subset(dat, !is.na(overweight))
  
  dat$tertile_low <- factor(dat$tertile_low, levels = c("T1", "T2", "T3"))
  
  form <- reformulate(
    termlabels = c("tertile_low", covars),
    response = "overweight"
  )
  
  glm(form, family = poisson(link = "log"), data = dat)
})

```

Extract robust estimates

```{r}
robust <- lapply(models, function(m) {
  ct <- coeftest(m, vcov. = vcovHC(m, type = "HC0"))
  list(
    coef = coef(m),
    vcov = vcovHC(m, type = "HC0"),
    ct = ct
  )
})
```

Confirm coefficient names

```{r}
names(robust[[1]]$coef)
```

Chose the terms to pool

```{r}
terms_to_pool <- c("tertile_lowT2", "tertile_lowT3")
```

Build summary table

```{r}
PR_summary_model1_low <- do.call(rbind, lapply(terms_to_pool, function(trm) {
  pool_pr_one_term(robust, trm)
}))

PR_summary_model1_low
```

UPPER BOUND - MODEL 1

```{r}

covars <- covars1

```

```{r}
models <- lapply(1:20, function(i) {
  dat <- complete(ungdom_imputed, i)
  
  dat$overweight <- as.numeric(dat$overweight == "yes")
  
  dat <- subset(dat, !is.na(overweight))
  
  dat$tertile_up <- factor(dat$tertile_up, levels = c("T1", "T2", "T3"))
  
  form <- reformulate(
    termlabels = c("tertile_up", covars),
    response = "overweight"
  )
  
  glm(form, family = poisson(link = "log"), data = dat)
})

```

Extract robust estimates

```{r}
robust <- lapply(models, function(m) {
  ct <- coeftest(m, vcov. = vcovHC(m, type = "HC0"))
  list(
    coef = coef(m),
    vcov = vcovHC(m, type = "HC0"),
    ct = ct
  )
})
```

Confirm coefficient names

```{r}
names(robust[[1]]$coef)
```

Chose the terms to pool

```{r}
terms_to_pool <- c("tertile_upT2", "tertile_upT3")
```

Build summary table

```{r}
PR_summary_model1_up <- do.call(rbind, lapply(terms_to_pool, function(trm) {
  pool_pr_one_term(robust, trm)
}))

PR_summary_model1_up
```

Create a descriptive table with n and % of overweight per tertile

1)  Subset to those non-missing overweight

```{r}
dat <- subset(ungdom, !is.na(overweight))
```

2)Create counts per tertile

```{r}
tab_n <- table(dat$tertile_mid, dat$overweight)
tab_n
```

3)  Extract number overweight == yes and calculate percentage_grps

```{r}
n_yes <- tab_n[, "yes"]

pct_yes <- prop.table(tab_n, margin = 1) [, "yes"] * 100

```

4)  Combine into a summary table

```{r}
desc_table <- data.frame(
  tertile = rownames(tab_n),
  n_overweight = as.numeric(n_yes),
  percent_overweight = round(pct_yes, 1)
)

desc_table
```

Write to excel

```{r}
sheets <- list(
  desc_table = desc_table,
  PR_summary_crude_mid = PR_summary_crude_mid,
  PR_summary_model1_mid = PR_summary_model1_mid,
  PR_summary_model2_mid = PR_summary_model2_mid,
  PR_summary_model3_mid = PR_summary_model3_mid,
  PR_summary_model1_low = PR_summary_model1_low,
  PR_summary_model1_up = PR_summary_model1_up
)

write.xlsx(sheets, 
           file.path(dir_output, file = "PR_summaries_ungdom.xlsx"), 
           overwrite = TRUE
          )
```

Sensitivity analysis of Poisson regression on pre-imputed dataset

1.  Prepare data for complete cases

```{r}
#Define covars

covars <- covars3
```

```{r}
dat <- ungdom
dat$overweight <- as.numeric(dat$overweight == "yes")

dat$tertile_mid <-  factor(dat$tertile_mid, levels = c("T1", "T2", "T3"))

vars_model <- c("overweight", "tertile_mid", covars)
dat_cc <- dat[complete.cases(dat[, vars_model]), ]
```

2.  Fit Poisson regression with log link

```{r}
form <- reformulate(
  termlabels = c("tertile_mid", covars),
  response = "overweight"
)

fit <- glm(
  form,
  family = poisson(link = "log"),
  data = dat_cc
)

```

3.  Obtain robust (sandwich) se

```{r}
robust_fit <- coeftest(fit, vcov. = vcovHC(fit, type = "HC0"))
robust_fit
```

4)  Convert to PR and CI

```{r}
est <- robust_fit[, "Estimate"]
se <- robust_fit[, "Std. Error"]

PR <- exp(est)
LCL <- exp(est - 1.96 * se)
UCL <- exp(est + 1.96 * se)

results <- data.frame(
  term = names(est),
  PR = PR,
  LCL = LCL,
  UCL = UCL,
  p_value = robust_fit[, "Pr(>|z|)"]
)

results
```
