---
title: "Suspected misreporters"
format: html
editor: visual
---
This script calculates prediction interval to identify suspected misreporters


```{r}
#| label: setup
#| include: false

project_root <- normalizePath(file.path(getwd(), ".."))

dir_data <- file.path(project_root, "data")
dir_output <- file.path(project_root, "output")

dir.create(dir_output, showWarnings = FALSE, recursive = TRUE)

```

Analytic code starts here


```{r}
library(dplyr)
library(openxlsx)

```

```{r}
master3 <- readRDS(file.path("Masterfile.rds")) 
```

Check number of NA for height and weight

```{r}
na_by_RM <- master3 %>% 
  group_by(RM) %>% 
  summarise(
    n = n(),
    height_NA = sum(is.na(height)),
    weight_NA = sum(is.na(weight)),
    either_NA = sum(is.na(height) | is.na(weight)),
    height_NA_pct = 100 * height_NA / n,
    weight_NA_pct = 100 * weight_NA / n,
    either_NA_pct = 100 * either_NA / n,
    .gropus = "drop"
  )

na_by_RM
```

Exclude cases with NA for height or weight

```{r}
master2 <- master3[!is.na(master3$height) | !is.na(master3$weight), ]
```

Exclude individuals with age \>18 years from RM Adolescents

```{r}
master1 <- master2 %>% 
  filter(!(RM == "Ungdom" & age > 18))

```

Keep only individuals in Age_group 4 ys among Young Children

```{r}
master1 <- master1[!(master1$RM == "Barn" & master1$age_grp == "18_manad"), ]
```

Exclude individuals with BMI or BMI-z score outside biological plausible range

```{r}
master <- master1[!(
  (master1$RM %in% c("Barn", "Ungdom") & (master1$bmi_z < -5 | master1$bmi_z > 5)) | (master1$RM == "Vuxna" & (master1$bmi < 12 | master1$bmi > 60))
), ]

```

Create new variables

```{r}

#sex code
master$sex_code <- ifelse(master$sex == "M", -1,
                          ifelse(master$sex == "F", 1, NA))

#MJ per day
master$MJ_perday <- master$kcal * 0.004184

```

Define elevation (weighted average between Stockholm, Göteborg and Malmö)

```{r}
#Define elevation
elevation <- 30
```

Create needed variables separately

```{r}
master <- master %>% 
  mutate(
    ln_weight = log(weight),
    age2 = age^2,
    age3 = age^3,
    ln_elevate = log(.env$elevation)
  )

```

Calculate equation

```{r}

master <- master %>% 
  mutate(
    ln_TEE = -0.2172 +
      ln_weight * 0.4167 +
      height_cm * 0.006565 -
      age * 0.02054 +
      age2 * 0.0003308 - 
      age3 * 0.000001852 +
      ln_elevate * 0.09126 -
      sex_code * 0.04092 +
      1 * 0.003589 -
      height_cm * ln_elevate * 0.0006759 +
      age * ln_elevate * 0.002018 -
      age2 * ln_elevate * 0.00002262 -
      sex_code * ln_elevate * 0.006947
  )

```

Exponentiate back

```{r}
master$TEE_MJ <- exp(master$ln_TEE)

```

Convert to kcal for double checking

```{r}

master$TEE_kcal <- master$TEE_MJ * 239.006

```

```{r}
summary(master$TEE_kcal)
```

Calculate lower and upper prediction intervals

```{r}

master$PI_low <- (master$TEE_MJ * 0.7466) - 1.5405
master$PI_high <- (master$TEE_MJ * 1.3395) + 2.7668

```

Calculate number and percent of excluded individuals in each RM-group

```{r}
excluded_summary <- master %>% 
  group_by(RM) %>% 
  summarise(
    total = n(),
    excluded_under = sum(MJ_perday < PI_low, na.rm = T),
    excluded_over = sum(MJ_perday > PI_high, na.rm = T),
    .groups = "drop"
  ) %>% 
  
mutate(
  excluded = excluded_under + excluded_over,
  kept = total - excluded,
  percent_excluded = 100 * excluded / total
)

excluded_summary

```

Convert PI from MJ to kcal

```{r}
master$PI_low_kcal <- master$PI_low * 239.006
master$PI_high_kcal <- master$PI_high * 239.006
```

Select variables of interest for sensitivity analysis

```{r}

vars_of_interest <- c("person_ID", "RM", "sex", "age", "age_grp", "kcal", "MJ_perday", "ED_nobev", "bmi", "bmi_class", "leisure", "edu_combined", "overweight", "smoke", "bmi_z", "screentime_combined", "EP_upf_mid", "EP_upf_low", "EP_upf_up", "PI_low", "PI_high", "TEE_MJ", "TEE_kcal")
  
```

```{r}
sensitivity <- master %>% 
  dplyr::select(all_of(vars_of_interest))
```

Write to excel

```{r}
saveRDS(
  sensitivity, 
  file = file.path(dir_output, "Sensitivity.rds")
)

```
