---
title: "Background Table 1"
format: html
editor: visual
---

This script compiles data for Table 1 (background table)

```{r}
#| label: setup
#| include: false

project_root <- normalizePath(file.path(getwd(), ".."))

dir_data <- file.path(project_root, "data")
dir_output <- file.path(project_root, "output")

dir.create(dir_output, showWarnings = FALSE, recursive = TRUE)

```

Analytic code starts here

```{r}
library(dplyr)
library(openxlsx)

```

```{r}
master <- readRDS(file.path(dir_data,"Masterfile.rds"))

```

```{r}
barn <- master %>% filter(RM == "Barn") %>% 
  dplyr::select(person_ID, RM, sex, age, age_grp, weight, height_cm, bmi, bmi_class, bmi_z, edu_combined, screentime_combined)


ungdom <- master %>% filter(RM == "Ungdom") %>% 
  dplyr::select(person_ID, RM, sex, age, age_grp, weight, height_cm, bmi, bmi_class, bmi_z, edu_combined, leisure, screentime_combined)


vux <- master %>% filter(RM == "Vuxna") %>% 
  dplyr::select(person_ID, RM, sex, age, age_grp, weight, height_cm, bmi, bmi_class, edu_combined, leisure, smoke)

```

First, create grouping variables and get number of observations (n) per group

```{r}
master2 <- master %>% 
  mutate(
    table_grp = case_when(
      RM == "Vuxna" ~"Vuxna (all)",
      RM == "Barn" ~paste0("Barn: ", age_grp),
      RM == "Ungdom" ~paste0("Ungdom: ", age_grp),
      TRUE ~NA_character_
  )
)

```

Different grouping used only for BMI-variable

```{r}
master2 <- master2 %>% 
  mutate(
    bmi_grp = case_when(
      RM == "Vuxna" ~"Vuxna (all)",
      
      RM == "Barn" ~paste0("Barn: ", age_grp),
      
      RM == "Ungdom" & age_grp != "Ak2_g" ~paste0("Ungdom: ", age_grp),
      
      RM == "Ungdom" & age_grp == "Ak2_g" & age < 18 ~"Ungdom: Ak2_g (<18)",
      RM == "Ungdom" & age_grp == "Ak2_g" & age >=18 ~"Ungdom: Ak2_g (>=18)",
      
      TRUE ~NA_character_
  )
)
```

List of group names

```{r}
sort(unique(master2$table_grp))
sort(unique(master2$bmi_grp))
```

Set ordering of groups for outputs

```{r}
table_grp_levels <- c(
  "Vuxna (all)",
  "Ungdom: Ak2_g",
  "Ungdom: Ak8",
  "Ungdom: Ak5",
  "Barn: 4_ar",
  "Barn: 18_manad"
)

bmi_grp_levels <- c(
  "Vuxna (all)",
  "Ungdom: Ak2_g (>=18)",
  "Ungdom: Ak2_g (<18)",
  "Ungdom: Ak8",
  "Ungdom: Ak5",
  "Barn: 4_ar",
  "Barn: 18_manad"
)

master2 <- master2 %>% 
  mutate(
    table_grp = factor(table_grp, levels = table_grp_levels),
    bmi_grp = factor(bmi_grp, levels = bmi_grp_levels)
  )

```

Count n observations in each group (used as denominators for percentages)

```{r}
group_n <- master2 %>% count(table_grp, name = "N")
bmi_n <- master2 %>% count(bmi_grp, name = "N")
```

General rules:

-For categorical variables: Percentages based on total group N, including missing. Missing are shown as its own row.

-For continuous variabes: n used (non-missing) reported separately. Missing values not included in median calculations.

1.  SEX

```{r}
library(forcats)

sex_tab <- master2 %>% 
  mutate(
    sex = fct_na_value_to_level(factor(sex), level = "Missing")
  ) %>% 
  count(table_grp, sex) %>% 
  left_join(group_n, by = "table_grp") %>% 
  mutate(
    pct = 100 * n / N, 
    n_pct = paste0(n, " (", round(pct, 1), "%)")
  )

sex_tab
```

2.  AGE

Create an integer age variable representing compleded years

```{r}
master2 <- master2 %>% 
  mutate(
    age_int = floor(age)
  )

```

```{r}
age_tab <- master2 %>% 
  group_by(table_grp) %>% 
  summarise(
    n_total = n(),
    n_missing = sum(is.na(age_int)),
    n_used = sum(!is.na(age_int)),
    median_age_int = median(age_int, na.rm = T),
    min_age_int = min(age_int, na.rm = T),
    max_age_int = max(age_int, na.rm = T),
    .groups = "drop"
  )
```

Format

```{r}
age_tab <- age_tab %>% 
  mutate(
    age_median_range = paste0(
      median_age_int,
      " (", min_age_int, "-", max_age_int, ")"
    ),
missing = paste0(
  n_missing,
  " (", round(100 * n_missing / n_total, 1), "%)"
    )
  )

age_tab
```

3.  WEIGHT

```{r}
weight_tab <- master2 %>% 
  group_by(table_grp) %>% 
  summarise(
    N_total = n(),
    N_missing = sum(is.na(weight)),
    median_wt = median(weight, na.rm = T),
    p25 = quantile(weight, 0.25, na.rm = T),
    p75 = quantile(weight, 0.75, na.rm = T),
    .groups = "drop"
  )
```

```{r}
weight_tab <- weight_tab %>% 
  mutate(
    weight_median_iqr = paste0(
      round(median_wt, 1),
      " (", round(p25, 1), "-", round(p75, 1), ")"
    ),
    missing = paste0(
      N_missing,
      " (", round(100 * N_missing / N_total, 1), "%)"
    )
  )

weight_tab
```

4.  HEIGHT

```{r}
height_tab <- master2 %>% 
  group_by(table_grp) %>% 
  summarise(
    N_total = n(),
    N_missing = sum(is.na(height_cm)),
    median_wt = median(height_cm, na.rm = T),
    p25 = quantile(height_cm, 0.25, na.rm = T),
    p75 = quantile(height_cm, 0.75, na.rm = T),
    .groups = "drop"
  ) %>% 
  mutate(
    weight_median_iqr = paste0(
      round(median_wt, 1),
      " (", round(p25, 1), "-", round(p75, 1), ")"
    ),
    missing = paste0(
      N_missing,
      " (", round(100 * N_missing / N_total, 1), "%)"
    )
  )

height_tab
```

5.  BMI / BMI-z score

<!-- -->

a)  BMI for adults only

```{r}
bmi_tab <- master2 %>% 
  filter(
    RM == "Vuxna" |
      (RM == "Ungdom" & age >=18)
  ) %>% 
  group_by(bmi_grp) %>% 
  summarise(
    N_total = n(),
    N_missing = sum(is.na(bmi)),
    median_bmi = median(bmi, na.rm = T),
    p25 = quantile(bmi, 0.25, na.rm = T),
    p75 = quantile(bmi, 0.75, na.rm = T),
    .groups = "drop"
  ) %>% 
  mutate(
    bmi_median_iqr = paste0(
      round(median_bmi, 1),
      " (", round(p25, 1), "-", round(p75, 1), ")"
    ),
    missing = paste0(
      N_missing,
      " (", round(100 * N_missing / N_total, 1), "%)"
    )
  )

bmi_tab
```

```{r}

bmiz_tab <- master2 %>% 
  filter(
    !(RM == "Vuxna" |
      (RM == "Ungdom" & age >=18))
  ) %>% 
  group_by(bmi_grp) %>% 
  summarise(
    N_total = n(),
    N_missing = sum(is.na(bmi_z)),
    median_bmiz = median(bmi_z, na.rm = T),
    p25 = quantile(bmi_z, 0.25, na.rm = T),
    p75 = quantile(bmi_z, 0.75, na.rm = T),
    .groups = "drop"
  ) %>% 
  mutate(
    bmiz_median_iqr = paste0(
      round(median_bmiz, 1),
      " (", round(p25, 1), "-", round(p75, 1), ")"
    ),
    missing = paste0(
      N_missing,
      " (", round(100 * N_missing / N_total, 1), "%)"
    )
  )



bmiz_tab
```

Quick sanity check

```{r}
table(master2$bmi_grp, useNA = "ifany")

bmi_tab %>% summarise(sum_N = sum(N_total))
bmiz_tab %>% summarise(sum_N = sum(N_total))

```

6.  BMI-class

Inspect current values

```{r}
sort(unique(master2$bmi_class))
```

Convert to factor and set order

```{r}
master2 <- master2 %>% 
  mutate(
    bmi_class = factor(
      bmi_class, 
      levels = c(
        "undervikt",
        "normalvikt",
        "övervikt",
        "fetma"
      )
    )
  )
```

Tabulate

```{r}
bmi_class_tab <- master2 %>% 
  mutate(
    bmi_class = fct_na_value_to_level(bmi_class, level = "Missing")
  ) %>% 
  count(table_grp, bmi_class) %>% 
  left_join(group_n, by = "table_grp") %>% 
  mutate(
    pct = 100 * n / N,
    n_pct = paste0(n, " (", round(pct, 1), "%)"
  )
)

bmi_class_tab
```

7.  EDUCATION

Inspect current values

```{r}
sort(unique(master2$edu_combined))
```

Convert to factor

```{r}
master2 <- master2 %>% 
  mutate(
    edu_combined = factor(
      edu_combined, 
      levels = c("lower", "higher")
    )
  )
```

Tabulate

```{r}
edu_tab <- master2 %>% 
  mutate(
    edu_combined = fct_na_value_to_level(edu_combined, level = "Missing")
  ) %>% 
  count(table_grp, edu_combined) %>% 
  left_join(group_n, by = "table_grp") %>% 
  mutate(
    pct = 100 * n / N,
    n_pct = paste0(n, " (", round(pct, 1), "%)"
  )
)

edu_tab

```

8.  PHYSICAL ACTIVITY

Inspect values

```{r}
sort(unique(master$leisure))
```

Convert to factor and order

```{r}
master2 <- master2 %>% 
  mutate(
    leisure = factor(
      leisure, 
      levels = c(
        "sedentary",
        "light",
        "moderate",
        "hard"
      )
    )
  )
```

Tabulate

```{r}
pa_tab <- master2 %>% 
  mutate(
    leisure = fct_na_value_to_level(leisure, level = "Missing")
  ) %>% 
  count(table_grp, leisure) %>% 
  left_join(group_n, by = "table_grp") %>% 
  mutate(
    pct = 100 * n / N,
    n_pct = paste0(n, " (", round(pct, 1), "%)"
  )
)

pa_tab
```

9.  SCREENTIME

```{r}
screen_tab <- master2 %>% 
  filter(RM == "Barn") %>% 
  group_by(table_grp) %>% 
  summarise(
    N_total = n(),
    N_missing = sum(is.na(screentime_combined)),
    median = median(screentime_combined, na.rm = T),
    p25 = quantile(screentime_combined, 0.25, na.rm = T),
    p75 = quantile(screentime_combined, 0.75, na.rm = T),
    .groups = "drop"
  ) %>% 
  mutate(
    screentime_median_iqr = paste0(
      round(median, 1),
      " (", round(p25, 1), "-", round(p75, 1), ")"
    ),
    missing = paste0(
      N_missing, 
      " (", round(100 * N_missing / N_total, 1), "%)"
    )
  )

screen_tab
```

10. SMOKING

Inspect values

```{r}
sort(unique(master$smoke))
```

Convert to factor and order

```{r}
master2 <- master2 %>% 
  mutate(
    smoke = factor(
      smoke, 
      levels = c(
        "never",
        "former",
        "occasional",
        "daily"
      )
    )
  )
```

Tabulate

```{r}
smoke_tab <- master2 %>% 
  filter(RM == "Vuxna") %>% 
  mutate(
    smoke = fct_na_value_to_level(smoke, level = "Missing")
  ) %>% 
  count(table_grp, smoke) %>% 
  left_join(group_n, by = "table_grp") %>% 
  mutate(
    pct = 100 * n / N,
    n_pct = paste0(n, " (", round(pct, 1), "%)"
  )
)

smoke_tab
```

Export tables to Excel

```{r}

library(writexl)

write_xlsx(
  list(
    Sex = sex_tab,
    Age = age_tab,
    Weight = weight_tab,
    Height = height_tab,
    BMI = bmi_tab,
    BMIz = bmiz_tab,
    BMIclass = bmi_class_tab,
    Education = edu_tab,
    Screentime = screen_tab,
    PA = pa_tab,
    Smoking = smoke_tab
  ),
  path = file.path(dir_output, "Table1_background.xlsx")
)

```
