---
title: "Linear regression weight outcomes"
format: html
editor: visual
---

Sensitivity analysis excluding suspected misreporters

1)  Multiple imputation
2)  Linear regression with logBMI as dependent variable
3)  Poisson regression with robust variance for Prevalence ratios of overweight/obesity Dataset: Adults

```{r}
#| label: setup
#| include: false

project_root <- normalizePath(file.path(getwd(), ".."))

dir_data <- file.path(project_root, "data")
dir_output <- file.path(project_root, "output")

dir.create(dir_output, showWarnings = FALSE, recursive = TRUE)

```

Analytic code starts here

```{r message=FALSE, warning=FALSE}
library(openxlsx)
library(dplyr)
library(forcats)
```

Import data

```{r}
master <- readRDS(file.path(dir_data,"Masterfile.rds")) 

sens <- readRDS(file.path(dir_data, "Sensitivity.rds"))

```

Join misrep column

```{r}
master <- master %>% 
  left_join(
    sens %>% select(person_ID, misrep),
    by = "person_ID"
  )
```

Create dataset for regression

```{r}
vux1 <- master %>% 
  filter(RM == "Vuxna" & misrep == "no") %>% 
  dplyr::select(person_ID, sex, age, edu_combined, leisure, nrf, smoke, kcal, ED_nobev, bmi, overweight, EP_upf_low, EP_upf_up, EP_upf_mid, misrep)

```

Check

```{r}
table(vux1$misrep, useNA = "ifany")
```

Convert selected variables to factors, log-transform bmi-variable

```{r}
vux1 <- vux1 %>% 
  mutate(
    sex = as.factor(sex),
    edu_combined = as.factor(edu_combined),
    leisure = as.factor(leisure),
    smoke = as.factor(smoke),
    overweight = as.factor(overweight),
    logBMI = log(bmi)
  )

```

Divide the dataset in tertiles based on the variable EP_upf

```{r}
vux1 <- vux1 %>% 
  mutate(
    tertile_low = factor(paste0("T", ntile(EP_upf_low, 3)), levels = paste0("T", 1:3)),
    tertile_mid = factor(paste0("T", ntile(EP_upf_mid, 3)), levels = paste0("T", 1:3)),
    tertile_up = factor(paste0("T", ntile(EP_upf_up, 3)), levels = paste0("T", 1:3)),
  )

```

Scale predictor variable

```{r}
vux1$EP_upf_low_10 <- vux1$EP_upf_low / 10
vux1$EP_upf_mid_10 <- vux1$EP_upf_mid / 10
vux1$EP_upf_up_10 <- vux1$EP_upf_up / 10
```

Count number of individuals with NA for logBMI

```{r}
sum(is.na(vux1$logBMI))
```

Exclude individuals with NA for the outcome before imputation

```{r}
vux <- vux1[!is.na(vux1$logBMI),]

nrow(vux)

```

Values outside the range of 12-60 were judged biologically implausible values

```{r}
vux$person_ID[vux$bmi <=12|vux$bmi >=60]

```

```{r}
vux <- vux[vux$bmi >=12 & vux$bmi <=60, ]

nrow(vux)
```

Double check so that "overweight" variable is correctly defined for all individuals

```{r}
#Define what should be overweight
should_be_overweight <- vux$bmi > 25

#Define what is actually coded as overweight
is_overweight_yes <- vux$overweight == "yes"

#Look for divergencies
divergent_cases <- vux[
  !is.na(vux$bmi) & !is.na(vux$overweight) &
  (should_be_overweight !=is_overweight_yes),
]

#Inspect divergent cases
divergent_cases[, c("person_ID", "bmi", "overweight")]

```

Check

```{r}
library(car)
fit <- lm(logBMI ~EP_upf_mid_10 + age + sex + edu_combined + kcal + nrf, data = vux)
fit
plot(fit)
vif(fit)
```

MULTIPLE IMPUTATION

```{r}
library(mice)
library(miceadds)
library(broom)

```

Prepare for imputation

```{r}

imp_data <- as.data.frame(vux)

#Define variables to exclude from imputation
exclude_vars <- c("logBMI", "bmi", "overweight", "tertile_low", "tertile_mid", "tertile_up", "EP_upf_mid_10", "EP_upf_low_10", "EP_upf_up_10")

pred <- make.predictorMatrix(imp_data)

#Outcomes not used to predict, not imputed
pred[, exclude_vars] <- 0
pred[exclude_vars,] <- 0

#ID not used either
pred[, "person_ID"] <- 0
pred["person_ID", ] <- 0

#Method vector
meth <- make.method(imp_data)

#Outcomes and ID should not be imputed
meth[exclude_vars] <- ""
meth["person_ID"] <- ""

#Sanity checks
stopifnot(is.matrix(pred))
stopifnot(all(rownames(pred) == names(imp_data)))
stopifnot(all(colnames(pred)== names(imp_data)))

```

Run mice

```{r}
vux_imputed <- mice(
  imp_data,
  m = 20,
  predictorMatrix = pred,
  method = meth,
  seed = 123
)

```

```{r}
summary(vux_imputed)
```

Inspect methods

```{r}
vux_imputed$method
```

How many individuals values were imputed?

```{r}
colSums(vux_imputed$where)
```

CRUDE MODEL

```{r}
crude <- with(
  vux_imputed,
  lm(logBMI ~EP_upf_mid_10)
)

crude_pooled <- pool(crude)
crude_pooled_sum <- summary(crude_pooled, conf.int = T)


n_per_imp <- sapply(crude$analyses, function(mod) nrow(model.frame(mod)))
unique_n <- unique(n_per_imp)
unique_n
crude_pooled_sum
```

```{r}
log_table <- crude_pooled_sum[2, c("term", "estimate", "conf.low", "conf.high", "p.value")]
colnames(log_table) <- c("term", "estimate", "conf.low", "conf.high", "p.value")

#Back-transform
back_table <- log_table
back_table$estimate = exp(log_table$estimate)
back_table$conf.low = exp(log_table$conf.low)
back_table$conf.high = exp(log_table$conf.high)

#Calculate percentual change 
perc_table <- log_table
perc_table$estimate <- (exp(log_table$estimate) -1)*100
perc_table$conf.low <- (exp(log_table$conf.low) -1)*100
perc_table$conf.high <- (exp(log_table$conf.high) -1)*100

final_table_crude <- list(
  log_scale = log_table,
  back_trans = back_table,
  perc_table = perc_table
)

final_table_crude

```

Define sets of covaritates

```{r}
covars1 <- c("age", "sex", "edu_combined", "leisure", "smoke")
covars2 <- c("age", "sex", "edu_combined", "leisure", "smoke", "kcal")
covars3 <- c("age", "sex", "edu_combined", "leisure", "smoke", "kcal", "nrf")
```

MODEL 1 - sociodemographics only

```{r}
model1 <- with(
  vux_imputed,
  lm(logBMI ~EP_upf_mid_10 + age + sex + edu_combined + leisure + smoke)
)

model1_pooled <- pool(model1)
model1_pooled_sum <- summary(model1_pooled, conf.int = T)


n_per_imp <- sapply(model1$analyses, function(mod) nrow(model.frame(mod)))
unique_n <- unique(n_per_imp)
unique_n
model1_pooled_sum
```

```{r}
log_table <- model1_pooled_sum[2, c("term", "estimate", "conf.low", "conf.high", "p.value")]
colnames(log_table) <- c("term", "estimate", "conf.low", "conf.high", "p.value")

#Back-transform
back_table <- log_table
back_table$estimate = exp(log_table$estimate)
back_table$conf.low = exp(log_table$conf.low)
back_table$conf.high = exp(log_table$conf.high)

#Calculate percentual change 
perc_table <- log_table
perc_table$estimate <- (exp(log_table$estimate) -1)*100
perc_table$conf.low <- (exp(log_table$conf.low) -1)*100
perc_table$conf.high <- (exp(log_table$conf.high) -1)*100

final_table_model1 <- list(
  log_scale = log_table,
  back_trans = back_table,
  perc_table = perc_table
)

final_table_model1

```

MODEL 2 - sociodemographics + energy intake

```{r}
model2 <- with(
  vux_imputed,
  lm(logBMI ~EP_upf_mid_10 + age + sex + edu_combined + leisure + smoke + kcal)
)

model2_pooled <- pool(model2)
model2_pooled_sum <- summary(model2_pooled, conf.int = T)


n_per_imp <- sapply(model2$analyses, function(mod) nrow(model.frame(mod)))
unique_n <- unique(n_per_imp)
unique_n
model2_pooled_sum
```

```{r}
log_table <- model2_pooled_sum[2, c("term", "estimate", "conf.low", "conf.high", "p.value")]
colnames(log_table) <- c("term", "estimate", "conf.low", "conf.high", "p.value")

#Back-transform
back_table <- log_table
back_table$estimate = exp(log_table$estimate)
back_table$conf.low = exp(log_table$conf.low)
back_table$conf.high = exp(log_table$conf.high)

#Calculate percentual change 
perc_table <- log_table
perc_table$estimate <- (exp(log_table$estimate) -1)*100
perc_table$conf.low <- (exp(log_table$conf.low) -1)*100
perc_table$conf.high <- (exp(log_table$conf.high) -1)*100

final_table_model2 <- list(
  log_scale = log_table,
  back_trans = back_table,
  perc_table = perc_table
)

final_table_model2

```

MODEL 3 - sodiodemographics + energy + nrf

```{r}
model3 <- with(
  vux_imputed,
  lm(logBMI ~EP_upf_mid_10 + age + sex + edu_combined + leisure + kcal + nrf)
)

model3_pooled <- pool(model3)
model3_pooled_sum <- summary(model3_pooled, conf.int = T)


n_per_imp <- sapply(model3$analyses, function(mod) nrow(model.frame(mod)))
unique_n <- unique(n_per_imp)
unique_n
model3_pooled_sum
```

```{r}
log_table <- model3_pooled_sum[2, c("term", "estimate", "conf.low", "conf.high", "p.value")]
colnames(log_table) <- c("term", "estimate", "conf.low", "conf.high", "p.value")

#Back-transform
back_table <- log_table
back_table$estimate = exp(log_table$estimate)
back_table$conf.low = exp(log_table$conf.low)
back_table$conf.high = exp(log_table$conf.high)

#Calculate percentual change 
perc_table <- log_table
perc_table$estimate <- (exp(log_table$estimate) -1)*100
perc_table$conf.low <- (exp(log_table$conf.low) -1)*100
perc_table$conf.high <- (exp(log_table$conf.high) -1)*100

final_table_model3 <- list(
  log_scale = log_table,
  back_trans = back_table,
  perc_table = perc_table
)

final_table_model3

```

Export to excel

```{r}
write.xlsx(
  list(
    Crude = final_table_crude,
    Model1 = final_table_model1,
    Model2 = final_table_model2,
    Model3 = final_table_model3
  ),
  file = file.path(dir_output,"Regression_BMI_vux_mid_sens.xlsx"),
  overwrite = TRUE
)
```

LOWER BOUND

CRUDE MODEL

```{r}
crude <- with(
  vux_imputed,
  lm(logBMI ~EP_upf_low_10)
)

crude_pooled <- pool(crude)
crude_pooled_sum <- summary(crude_pooled, conf.int = T)


n_per_imp <- sapply(crude$analyses, function(mod) nrow(model.frame(mod)))
unique_n <- unique(n_per_imp)
unique_n
crude_pooled_sum
```

```{r}
log_table <- crude_pooled_sum[2, c("term", "estimate", "conf.low", "conf.high", "p.value")]
colnames(log_table) <- c("term", "estimate", "conf.low", "conf.high", "p.value")

#Back-transform
back_table <- log_table
back_table$estimate = exp(log_table$estimate)
back_table$conf.low = exp(log_table$conf.low)
back_table$conf.high = exp(log_table$conf.high)

#Calculate percentual change 
perc_table <- log_table
perc_table$estimate <- (exp(log_table$estimate) -1)*100
perc_table$conf.low <- (exp(log_table$conf.low) -1)*100
perc_table$conf.high <- (exp(log_table$conf.high) -1)*100

final_table_crude <- list(
  log_scale = log_table,
  back_trans = back_table,
  perc_table = perc_table
)

final_table_crude

```

Interpretation of estimate 1.0025: Since 1.0025 - 1 = 0.0025, it means that each 10% increase in E% upf was associated with 0.25% higher BMI.

MODEL 1 - Sociodemographics only

```{r}
model1 <- with(
  vux_imputed,
  lm(logBMI ~EP_upf_low_10 + age + sex + edu_combined + leisure + smoke)
)

model1_pooled <- pool(model1)
model1_pooled_sum <- summary(model1_pooled, conf.int = T)


n_per_imp <- sapply(model1$analyses, function(mod) nrow(model.frame(mod)))
unique_n <- unique(n_per_imp)
unique_n
model1_pooled_sum
```

```{r}
log_table <- model1_pooled_sum[2, c("term", "estimate", "conf.low", "conf.high", "p.value")]
colnames(log_table) <- c("term", "estimate", "conf.low", "conf.high", "p.value")

#Back-transform
back_table <- log_table
back_table$estimate = exp(log_table$estimate)
back_table$conf.low = exp(log_table$conf.low)
back_table$conf.high = exp(log_table$conf.high)

#Calculate percentual change 
perc_table <- log_table
perc_table$estimate <- (exp(log_table$estimate) -1)*100
perc_table$conf.low <- (exp(log_table$conf.low) -1)*100
perc_table$conf.high <- (exp(log_table$conf.high) -1)*100

final_table_model1 <- list(
  log_scale = log_table,
  back_trans = back_table,
  perc_table = perc_table
)

final_table_model1

```

MODEL 2 - Sociodemographics + EI

```{r}
model2 <- with(
  vux_imputed,
  lm(logBMI ~EP_upf_low_10 + age + sex + edu_combined + leisure + smoke + kcal)
)

model2_pooled <- pool(model2)
model2_pooled_sum <- summary(model2_pooled, conf.int = T)


n_per_imp <- sapply(model2$analyses, function(mod) nrow(model.frame(mod)))
unique_n <- unique(n_per_imp)
unique_n
model2_pooled_sum
```

```{r}
log_table <- model2_pooled_sum[2, c("term", "estimate", "conf.low", "conf.high", "p.value")]
colnames(log_table) <- c("term", "estimate", "conf.low", "conf.high", "p.value")

#Back-transform
back_table <- log_table
back_table$estimate = exp(log_table$estimate)
back_table$conf.low = exp(log_table$conf.low)
back_table$conf.high = exp(log_table$conf.high)

#Calculate percentual change 
perc_table <- log_table
perc_table$estimate <- (exp(log_table$estimate) -1)*100
perc_table$conf.low <- (exp(log_table$conf.low) -1)*100
perc_table$conf.high <- (exp(log_table$conf.high) -1)*100

final_table_model2 <- list(
  log_scale = log_table,
  back_trans = back_table,
  perc_table = perc_table
)

final_table_model2

```

MODEL 3 - Sociodemographics + EI + diet quality

```{r}
model3 <- with(
  vux_imputed,
  lm(logBMI ~EP_upf_low_10 + age + sex + edu_combined + leisure + smoke + kcal + nrf)
)

model3_pooled <- pool(model3)
model3_pooled_sum <- summary(model3_pooled, conf.int = T)


n_per_imp <- sapply(model3$analyses, function(mod) nrow(model.frame(mod)))
unique_n <- unique(n_per_imp)
unique_n
model3_pooled_sum
```

```{r}
log_table <- model3_pooled_sum[2, c("term", "estimate", "conf.low", "conf.high", "p.value")]
colnames(log_table) <- c("term", "estimate", "conf.low", "conf.high", "p.value")

#Back-transform
back_table <- log_table
back_table$estimate = exp(log_table$estimate)
back_table$conf.low = exp(log_table$conf.low)
back_table$conf.high = exp(log_table$conf.high)

#Calculate percentual change 
perc_table <- log_table
perc_table$estimate <- (exp(log_table$estimate) -1)*100
perc_table$conf.low <- (exp(log_table$conf.low) -1)*100
perc_table$conf.high <- (exp(log_table$conf.high) -1)*100

final_table_model3 <- list(
  log_scale = log_table,
  back_trans = back_table,
  perc_table = perc_table
)

final_table_model3

```

Export to excel

```{r}
write.xlsx(
  list(
    Crude = final_table_crude,
    Model1 = final_table_model1,
    Model2 = final_table_model2,
    Model3 = final_table_model3
  ),
  file = file.path(dir_output, "Regression_BMI_vux_low_sens.xlsx"),
  overwrite = TRUE
)
```

UPPER BOUND

CRUDE MODEL

```{r}
crude <- with(
  vux_imputed,
  lm(logBMI ~EP_upf_up_10)
)

crude_pooled <- pool(crude)
crude_pooled_sum <- summary(crude_pooled, conf.int = T)


n_per_imp <- sapply(crude$analyses, function(mod) nrow(model.frame(mod)))
unique_n <- unique(n_per_imp)
unique_n
crude_pooled_sum
```

```{r}
log_table <- crude_pooled_sum[2, c("term", "estimate", "conf.low", "conf.high", "p.value")]
colnames(log_table) <- c("term", "estimate", "conf.low", "conf.high", "p.value")

#Back-transform
back_table <- log_table
back_table$estimate = exp(log_table$estimate)
back_table$conf.low = exp(log_table$conf.low)
back_table$conf.high = exp(log_table$conf.high)

#Calculate percentual change 
perc_table <- log_table
perc_table$estimate <- (exp(log_table$estimate) -1)*100
perc_table$conf.low <- (exp(log_table$conf.low) -1)*100
perc_table$conf.high <- (exp(log_table$conf.high) -1)*100

final_table_crude <- list(
  log_scale = log_table,
  back_trans = back_table,
  perc_table = perc_table
)

final_table_crude

```

Interpretation of estimate 1.0025: Since 1.0025 - 1 = 0.0025, it means that each 10% increase in E% upf was associated with 0.25% higher BMI.

MODEL 1 - Sociodemographics only

```{r}
model1 <- with(
  vux_imputed,
  lm(logBMI ~EP_upf_up_10 + age + sex + edu_combined + leisure + smoke)
)

model1_pooled <- pool(model1)
model1_pooled_sum <- summary(model1_pooled, conf.int = T)


n_per_imp <- sapply(model1$analyses, function(mod) nrow(model.frame(mod)))
unique_n <- unique(n_per_imp)
unique_n
model1_pooled_sum
```

```{r}
log_table <- model1_pooled_sum[2, c("term", "estimate", "conf.low", "conf.high", "p.value")]
colnames(log_table) <- c("term", "estimate", "conf.low", "conf.high", "p.value")

#Back-transform
back_table <- log_table
back_table$estimate = exp(log_table$estimate)
back_table$conf.low = exp(log_table$conf.low)
back_table$conf.high = exp(log_table$conf.high)

#Calculate percentual change 
perc_table <- log_table
perc_table$estimate <- (exp(log_table$estimate) -1)*100
perc_table$conf.low <- (exp(log_table$conf.low) -1)*100
perc_table$conf.high <- (exp(log_table$conf.high) -1)*100

final_table_model1 <- list(
  log_scale = log_table,
  back_trans = back_table,
  perc_table = perc_table
)

final_table_model1

```

MODEL 2 - Sociodemographics + EI

```{r}
model2 <- with(
  vux_imputed,
  lm(logBMI ~EP_upf_up_10 + age + sex + edu_combined + leisure + smoke + kcal)
)

model2_pooled <- pool(model2)
model2_pooled_sum <- summary(model2_pooled, conf.int = T)


n_per_imp <- sapply(model2$analyses, function(mod) nrow(model.frame(mod)))
unique_n <- unique(n_per_imp)
unique_n
model2_pooled_sum
```

```{r}
log_table <- model2_pooled_sum[2, c("term", "estimate", "conf.low", "conf.high", "p.value")]
colnames(log_table) <- c("term", "estimate", "conf.low", "conf.high", "p.value")

#Back-transform
back_table <- log_table
back_table$estimate = exp(log_table$estimate)
back_table$conf.low = exp(log_table$conf.low)
back_table$conf.high = exp(log_table$conf.high)

#Calculate percentual change 
perc_table <- log_table
perc_table$estimate <- (exp(log_table$estimate) -1)*100
perc_table$conf.low <- (exp(log_table$conf.low) -1)*100
perc_table$conf.high <- (exp(log_table$conf.high) -1)*100

final_table_model2 <- list(
  log_scale = log_table,
  back_trans = back_table,
  perc_table = perc_table
)

final_table_model2

```

MODEL 3 - Sociodemographics + kcal + nrf

```{r}
model3 <- with(
  vux_imputed,
  lm(logBMI ~EP_upf_up_10 + age + sex + edu_combined + leisure + smoke + kcal + nrf)
)

model3_pooled <- pool(model3)
model3_pooled_sum <- summary(model3_pooled, conf.int = T)


n_per_imp <- sapply(model3$analyses, function(mod) nrow(model.frame(mod)))
unique_n <- unique(n_per_imp)
unique_n
model3_pooled_sum
```

```{r}
log_table <- model3_pooled_sum[2, c("term", "estimate", "conf.low", "conf.high", "p.value")]
colnames(log_table) <- c("term", "estimate", "conf.low", "conf.high", "p.value")

#Back-transform
back_table <- log_table
back_table$estimate = exp(log_table$estimate)
back_table$conf.low = exp(log_table$conf.low)
back_table$conf.high = exp(log_table$conf.high)

#Calculate percentual change 
perc_table <- log_table
perc_table$estimate <- (exp(log_table$estimate) -1)*100
perc_table$conf.low <- (exp(log_table$conf.low) -1)*100
perc_table$conf.high <- (exp(log_table$conf.high) -1)*100

final_table_model3 <- list(
  log_scale = log_table,
  back_trans = back_table,
  perc_table = perc_table
)

final_table_model3

```

Export to excel

```{r}
write.xlsx(
  list(
    Crude = final_table_crude,
    Model1 = final_table_model1,
    Model2 = final_table_model2,
    Model3 = final_table_model3
  ),
  file = file.path(dir_output,"Regression_BMI_vux_up_sens.xlsx"),
  overwrite = TRUE
)

```

PREVALENCE RATIOS

Load packages for robust SE

```{r}
library(mitools)
library(sandwich)
library(tibble)
library(miceadds)
library(lmtest)
```

define covars

```{r}
covars0 <- character(0)
covars1 <- c("age", "sex", "edu_combined", "leisure", "smoke")
covars2 <- c("age", "sex", "edu_combined", "leisure", "smoke", "kcal")
covars3 <- c("age", "sex", "edu_combined", "leisure", "smoke", "kcal", "nrf")
```

MIDDLE-BOUND

CRUDE MODEL

```{r}

covars <- covars0

```

```{r}
models <- lapply(1:20, function(i) {
  dat <- complete(vux_imputed, i)
  
  dat$overweight <- as.numeric(dat$overweight == "yes")
  
  dat <- subset(dat, !is.na(overweight))
  
  dat$tertile_mid <- factor(dat$tertile_mid, levels = c("T1", "T2", "T3"))
  
  form <- reformulate(
    termlabels = c("tertile_mid", covars),
    response = "overweight"
  )
  
  glm(form, family = poisson(link = "log"), data = dat)
})

```

Extract robust estimates

```{r}
robust <- lapply(models, function(m) {
  ct <- coeftest(m, vcov. = vcovHC(m, type = "HC0"))
  list(
    coef = coef(m),
    vcov = vcovHC(m, type = "HC0"),
    ct = ct
  )
})
```

Confirm coefficient names

```{r}
names(robust[[1]]$coef)
```

Chose the terms to pool

```{r}
terms_to_pool <- c("tertile_midT2", "tertile_midT3")
```

Create a pooling function (Rubin's rules for one term)

```{r}
pool_pr_one_term <- function(robust, term) {
  m <- length(robust)
  
  Q <- sapply(robust, function(x) x$coef[term])
  U <- sapply(robust, function(x) x$vcov[term, term])
  
  Qbar <- mean(Q)
  Ubar <- mean(U)
  B <- var(Q)
  T <- Ubar + (1+1/m)*B
  se <- sqrt(T)
  
  z <- Qbar / se
  p <- 2 * pnorm(-abs(z)) #two-sided Wald p-value
  
  data.frame(
    term = term,
    PR = exp(Qbar),
    LCL = exp(Qbar - 1.96 * se),
    UCL = exp(Qbar + 1.96 * se),
    logPR = Qbar,
    SE = se,
    z = z,
    p_value = p,
    stringsAsFactors = FALSE
  )
}
```

Build summary table

```{r}
PR_summary_crude_mid <- do.call(rbind, lapply(terms_to_pool, function(term) {
  pool_pr_one_term(robust, term)
}))

PR_summary_crude_mid
```

MODEL 1

```{r}

covars <- covars1

```

```{r}
models <- lapply(1:20, function(i) {
  dat <- complete(vux_imputed, i)
  
  dat$overweight <- as.numeric(dat$overweight == "yes")
  
  dat <- subset(dat, !is.na(overweight))
  
  dat$tertile_mid <- factor(dat$tertile_mid, levels = c("T1", "T2", "T3"))
  
  form <- reformulate(
    termlabels = c("tertile_mid", covars),
    response = "overweight"
  )
  
  glm(form, family = poisson(link = "log"), data = dat)
})

```

Extract robust estimates

```{r}
robust <- lapply(models, function(m) {
  ct <- coeftest(m, vcov. = vcovHC(m, type = "HC0"))
  list(
    coef = coef(m),
    vcov = vcovHC(m, type = "HC0"),
    ct = ct
  )
})
```

Confirm coefficient names

```{r}
names(robust[[1]]$coef)
```

Chose the terms to pool

```{r}
terms_to_pool <- c("tertile_midT2", "tertile_midT3")
```

Build summary table

```{r}
PR_summary_model1_mid <- do.call(rbind, lapply(terms_to_pool, function(term) {
  pool_pr_one_term(robust, term)
}))

PR_summary_model1_mid
```

MODEL 2

```{r}

covars <- covars2

```

```{r}
models <- lapply(1:20, function(i) {
  dat <- complete(vux_imputed, i)
  
  dat$overweight <- as.numeric(dat$overweight == "yes")
  
  dat <- subset(dat, !is.na(overweight))
  
  dat$tertile_mid <- factor(dat$tertile_mid, levels = c("T1", "T2", "T3"))
  
  form <- reformulate(
    termlabels = c("tertile_mid", covars),
    response = "overweight"
  )
  
  glm(form, family = poisson(link = "log"), data = dat)
})

```

Extract robust estimates

```{r}
robust <- lapply(models, function(m) {
  ct <- coeftest(m, vcov. = vcovHC(m, type = "HC0"))
  list(
    coef = coef(m),
    vcov = vcovHC(m, type = "HC0"),
    ct = ct
  )
})
```

Confirm coefficient names

```{r}
names(robust[[1]]$coef)
```

Chose the terms to pool

```{r}
terms_to_pool <- c("tertile_midT2", "tertile_midT3")
```

Build summary table

```{r}
PR_summary_model2_mid <- do.call(rbind, lapply(terms_to_pool, function(term) {
  pool_pr_one_term(robust, term)
}))

PR_summary_model2_mid
```

MODEL 3

```{r}

covars <- covars3

```

```{r}
models <- lapply(1:20, function(i) {
  dat <- complete(vux_imputed, i)
  
  dat$overweight <- as.numeric(dat$overweight == "yes")
  
  dat <- subset(dat, !is.na(overweight))
  
  dat$tertile_mid <- factor(dat$tertile_mid, levels = c("T1", "T2", "T3"))
  
  form <- reformulate(
    termlabels = c("tertile_mid", covars),
    response = "overweight"
  )
  
  glm(form, family = poisson(link = "log"), data = dat)
})

```

Extract robust estimates

```{r}
robust <- lapply(models, function(m) {
  ct <- coeftest(m, vcov. = vcovHC(m, type = "HC0"))
  list(
    coef = coef(m),
    vcov = vcovHC(m, type = "HC0"),
    ct = ct
  )
})
```

Confirm coefficient names

```{r}
names(robust[[1]]$coef)
```

Chose the terms to pool

```{r}
terms_to_pool <- c("tertile_midT2", "tertile_midT3")
```

Build summary table

```{r}
PR_summary_model3_mid <- do.call(rbind, lapply(terms_to_pool, function(term) {
  pool_pr_one_term(robust, term)
}))

PR_summary_model3_mid
```

LOWER BOUND - MODEL 1

```{r}

covars <- covars1

```

```{r}
models <- lapply(1:20, function(i) {
  dat <- complete(vux_imputed, i)
  
  dat$overweight <- as.numeric(dat$overweight == "yes")
  
  dat <- subset(dat, !is.na(overweight))
  
  dat$tertile_low <- factor(dat$tertile_low, levels = c("T1", "T2", "T3"))
  
  form <- reformulate(
    termlabels = c("tertile_low", covars),
    response = "overweight"
  )
  
  glm(form, family = poisson(link = "log"), data = dat)
})

```

Extract robust estimates

```{r}
robust <- lapply(models, function(m) {
  ct <- coeftest(m, vcov. = vcovHC(m, type = "HC0"))
  list(
    coef = coef(m),
    vcov = vcovHC(m, type = "HC0"),
    ct = ct
  )
})
```

Confirm coefficient names

```{r}
names(robust[[1]]$coef)
```

Chose the terms to pool

```{r}
terms_to_pool <- c("tertile_lowT2", "tertile_lowT3")
```

Build summary table

```{r}
PR_summary_model1_low <- do.call(rbind, lapply(terms_to_pool, function(trm) {
  pool_pr_one_term(robust, trm)
}))

PR_summary_model1_low
```

UPPER BOUND - MODEL 1

```{r}

covars <- covars1

```

```{r}
models <- lapply(1:20, function(i) {
  dat <- complete(vux_imputed, i)
  
  dat$overweight <- as.numeric(dat$overweight == "yes")
  
  dat <- subset(dat, !is.na(overweight))
  
  dat$tertile_up <- factor(dat$tertile_up, levels = c("T1", "T2", "T3"))
  
  form <- reformulate(
    termlabels = c("tertile_up", covars),
    response = "overweight"
  )
  
  glm(form, family = poisson(link = "log"), data = dat)
})

```

Extract robust estimates

```{r}
robust <- lapply(models, function(m) {
  ct <- coeftest(m, vcov. = vcovHC(m, type = "HC0"))
  list(
    coef = coef(m),
    vcov = vcovHC(m, type = "HC0"),
    ct = ct
  )
})
```

Confirm coefficient names

```{r}
names(robust[[1]]$coef)
```

Chose the terms to pool

```{r}
terms_to_pool <- c("tertile_upT2", "tertile_upT3")
```

Build summary table

```{r}
PR_summary_model1_up <- do.call(rbind, lapply(terms_to_pool, function(trm) {
  pool_pr_one_term(robust, trm)
}))

PR_summary_model1_up
```

Create a descriptive table with n and % of overweight per tertile

1)  Subset to those non-missing overweight

```{r}
dat <- subset(vux, !is.na(overweight))
```

2)Create counts per tertile

```{r}
tab_n <- table(dat$tertile_mid, dat$overweight)
tab_n
```

3)  Extract number overweight == yes and calculate percentages

```{r}
n_yes <- tab_n[, "yes"]

pct_yes <- prop.table(tab_n, margin = 1) [, "yes"] * 100

```

4)  Combine into a summary table

```{r}
desc_table <- data.frame(
  tertile = rownames(tab_n),
  n_overweight = as.numeric(n_yes),
  percent_overweight = round(pct_yes, 1)
)

desc_table
```

Write to excel

```{r}
sheets <- list(
  desc_table = desc_table,
  PR_summary_crude_mid = PR_summary_crude_mid,
  PR_summary_model1_mid = PR_summary_model1_mid,
  PR_summary_model2_mid = PR_summary_model2_mid,
  PR_summary_model3_mid = PR_summary_model3_mid,
  PR_summary_model1_low = PR_summary_model1_low,
  PR_summary_model1_up = PR_summary_model1_up
)

write.xlsx(sheets, 
           file = file.path(dir_output, "PR_summaries_vux_sens.xlsx"), 
           overwrite = TRUE
          )
```

Sensitivity analysis of Poisson regression on pre-imputed dataset

1.  Prepare data for complete cases

```{r}
#Define covars

covars <- covars3
```

```{r}
dat <- vux
dat$overweight <- as.numeric(dat$overweight == "yes")

dat$tertile_mid <-  factor(dat$tertile_mid, levels = c("T1", "T2", "T3"))

vars_model <- c("overweight", "tertile_mid", covars)
dat_cc <- dat[complete.cases(dat[, vars_model]), ]
```

2.  Fit Poisson regression with log link

```{r}
form <- reformulate(
  termlabels = c("tertile_mid", covars),
  response = "overweight"
)

fit <- glm(
  form,
  family = poisson(link = "log"),
  data = dat_cc
)

```

3.  Obtain robust (sandwich) se

```{r}
robust_fit <- coeftest(fit, vcov. = vcovHC(fit, type = "HC0"))
robust_fit
```

4)  Convert to PR and CI

```{r}
est <- robust_fit[, "Estimate"]
se <- robust_fit[, "Std. Error"]

PR <- exp(est)
LCL <- exp(est - 1.96 * se)
UCL <- exp(est + 1.96 * se)

results <- data.frame(
  term = names(est),
  PR = PR,
  LCL = LCL,
  UCL = UCL,
  p_value = robust_fit[, "Pr(>|z|)"]
)

results
```
